### 2016-04-27 paper.tex
paper: switch to latex :(
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\documentclass[10pt,twocolumn]{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{graphicx,grffile}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[latin1]{inputenc}

% do not change these values
\baselineskip 12pt
\textheight 9in
\textwidth 6.5in
\oddsidemargin 0in
\topmargin 0in
\headheight 0in
\headsep 0in

%\makeatletter
%\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
%\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
%\makeatother
%% Scale images if necessary, so that they will not overflow the page
%% margins by default, and it is still possible to overwrite the defaults
%% using explicit options in \includegraphics[width, height, ...]{}
%\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
%

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\usepackage[font=footnotesize,labelfont=bf]{caption}

\begin{document}

\title{title}

\author{
%Noah Watkins, Neha Ojha\textsuperscript{*}, Carlos Maltzahn \\
%\small {\em University of California, Santa Cruz} \\
%\small {\{jayhawk,carlosm\}@cs.ucsc.edu} \textsuperscript{*}nojha@ucsc.edu \\ [2mm]
\small Submission Type: Research
}

\date{}
\maketitle

\begin{abstract}
abstract
\end{abstract}

\section{Introduction}

\begin{figure}[h]
  \centering
    \includegraphics[width=0.45\textwidth]{experiments/objclass-dev/output.png}
    \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/objclass-dev/visualize.ipynb}{source}]
Growth of officially supported, custom object interfaces in RADOS over 6
years. A \emph{method} is a specific object interface and a \emph{class} is a
logical grouping of methods.
}
\end{figure}

\section{Storage System Programmability}

When application goals are not met by a storage system the most common reaction
is to design a workaround. Workarounds roughly fall into one of two categories:
so called ``bolt-on'' services that introduce a 3rd party system (e.g. a
metadata service), or expanded application responsibility in the form of data
management (e.g. a new data layout).

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
Category & Specialization & Methods \\ \hline
\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                         & Exclusive & \\ \hline
\multirow{3}{*}{Logging} & Replica & 3 \\
                         & State & 4 \\
                         & Timestamped & 4 \\ \hline
Garbage Collection & Reference Counting & 4 \\ \hline
\multirow{4}{*}{Metadata Management} & RBD & 37 \\
 & RGW & 27 \\
 & User & 5 \\
 & Version & 5 \\ \hline
\end{tabular}
\caption{A variety of RADOS object storage classes exist that expose reusable interfaces to applications.}
\label{tab:objclasses}
\end{table}

\section{Consolidated Section}

Up until this point we have introduced Ceph, CORFU, the general problems
associated with programmability, described ZLog as being an instantiation of
the CORFU interface on top of Ceph.

Template for the CORFU object class interface. Maybe a simple state machine.

Describe indexing and protocol enforcement.
Differences between built-in interfaces and custom cls interfaces.

\begin{figure}
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.2cm,semithick]
%\tikzstyle{every state}=[fill=red,draw=none,text=white]

  \node[initial left,state] (A)              {$EG$};
  \node[state]         (B) [below right of=A] {$CP$};
  \node[state]         (D) [right of=B]       {$IO$};
  \node[state]         (C) [above right of=A] {$GC$};
  \node[state]         (E) [right of=C]       {$US$};
  \node[state]         (F) [below right of=E] {$Fi$};

  \path (A) edge        node [left] {R,W,F} (B)
            edge        node {T}     (C)
        (B) edge        node {F}     (C)
            edge        node {R,W}   (D)
        (C) edge        node {F,T}   (E)
        (D) edge        node {W}     (E)
            edge        node {R}     (F)
        (E) edge        node {F,T,W} (F);
\end{tikzpicture}
\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
and update state ($US$) access metadata. The I/O performs a log entry read or
write, and garbage collection ($GC$) marks entries for reclamation.}
\end{figure}

\section{ZLog Physical Design}

In this section we examine the design space for mapping the CORFU protocol onto
the RADOS storage system. Recall from Section X the options for rados,
and from section Y for corfu options. The cross product forms the design space.
Want to winnow this down.

\subsection{Storage Interface Selection}

\begin{table}
\begin{tabular}{ | l | l | l | l | l |}
\hline
Map & I/O & Entry Size & Addressing & Metadata \\ \hline
\multirow{2}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
                     & BS  & Flex     & Ceph/VFS  & KV/BS \\ \hline
\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
                     & WR  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
                     & AP  & Flex     & KV/BS     & KV/BS \\
\hline
\end{tabular}
\caption{The high-level design space of mapping CORFU log entry storage onto
the RADOS object storage system.}
\label{t:init-ds}
\end{table}

Constructing a complete implementation for each configuration would take
a lot of time, and may be time wasted if some configurations can be shown
to never be a viable option.

To winnow down the design space we ran a set of benchmarks---one for each
configuration--- that stressed the raw write performance using the built-in,
optimized interfaces on an unmodified version of Ceph. We explicitly avoided
simulating the costs associated with indexing or protocol enforcement in order
to form a clean baseline. Log entries may vary in size depending on the
application, but per-entry metadata is constant.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
interfaces in various configurations. The best performance is achieved using
the byte stream interface and a N:1 mapping strategy.
}
\end{figure}

The first thing to notice is that the configurations using a 1:1 mapping
perform poorly relative to the other configurations. This is unfortunate
because a 1:1 mapping simplifies data management by avoiding the need to
multiplex entries onto a smaller set of objects. However, even if performance
was acceptable, notice that throughput dramatically drops in the tail when the
number of objects reaches a threshold forcing the object cache to spill to
disk. Note that in a large cluster this mode may take a long time reach,
but we avoid it due to the inherent risk.

The KV/N:1 is a convenient configuration because the key-value interface
can be used to easily address log entries and manage index data. However, the
performance of KV/N:1 can vary widely depending on the size of the data entry.
As we will see next, even though the KV/N:1 configuration simplifies data
management, small entry performance falls behind that of N:1 configurations
that store data in the stream interface.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
Throughput of 1K random reads to a single OSD using the standard I/O
interfaces in various configurations. The best performance is achieved using
the byte stream interface and a N:1 mapping strategy.
}
\end{figure}

Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
for each of the points in the design space defined in Table \ref{t:init-ds}.
The results reveal poor relative performance using both the key-value storage
interface, as well as either of the 1:1 strategies---which incur overhead by
creating a new object for each log entry. Both of the N:1 strategies using the
bytestream I/O interface outperform all of the other approaches by over 2x
throughput, but otherwise have nearly identical performance. As a first
approximation these results show that when optimizing for log append throughput
the design space should be refined by limiting solutions to architectures based
on a N:1 mapping strategy using the bytestream I/O interface.

To examine the read performance of the points in the design space we generated
a dataset using each of the write workloads, and then issued random reads
across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
throughput for each of the design strategies. The results show that an N:1
mapping using bytestream interface performs best. Interestingly, the 1:1
mapping strategy using the bytestream interface exhibits good read performance
in comparison to writes using the same strategy.

The configurations based on bytestream storage exhibit the best performance.
The *write* variation depends on client-side mapping to object offset and
indirectly on the VFS index layer within the OSD, but this method is
incompatible with non-fixed size entries. Alternatively, the *append* variation
supports both fixed and dynamic size entries, but leaves the indexing and
protocl enforcment challenges unsolved.

In conclusion, both *write* and *append* variation of bytestream configurations
are the best choice, with an edge given to *append* if it can support dynamic
sized entries without sacrificing performance.

\subsection{Metadata Management}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
  \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
}
\end{figure}

Many variations of design here. What is the easiest way to 
winnow down the design space.

NEXT GRAPH A

This is an easy interface to simulate.
- Check Epoch (omap, stream header)
- this graph also has the APPEND raw through for comparison
- adds NEW OSD OP TO AVOID CLS OVERHEAD DISCUSSION FOR PAPER
- 3 runs in this graph, no need to write interface
- short runs to verify and then do all long runs togther with next graph

Both protocol enforcement and logical address translation
represent a larger design space involving index design. However
since in either case all state must be stored in the omap and
in the stream header, testing the overhead of epoch enforcement
provides a strict upper bound on performance.

- Protocol Enforcement
- Logical Translation (only append case)

Memory Cache

NEXT GRAPH B (or combined with A)
Use in-memory cache to implement the epoch check and show that we can match
the raw performance even with this added functionality.
This will show the content from graph A for comparison. Look
we match the raw performance.

NEXT GRAPH B or C (probably just two).. bar chart?
 Final consolidation of average throughputs of all the different
methods
 THIS INCLUDES THE FULL IMPL OF ZLOG APPEND using the in-memory cache

We solve the slow down of the epoch by storing the value in memory. Here is
the new performance.  We are able to with the epoch check match the raw
performance.

We know that we cannot use kv/bs for logicla address translation or protocol
enforcement without a performance hit so there is no point in trying. instead,
we generalize this memory cache hting used for epoch checks as a
new interface and use this interface to implement the full zlog
stack and show the performance.

Overall corfu is relatively simple abtraction, yet requires significant
work in making performant. Further, the decision may change over time.

We will probably have room in the paper to discuss the design of the
index structure, like collapsing a bit map and taking advantage of
the implicit strides in the offsets, etc... but we can probably move
that to a later section to keep this presentation concise. We should
probably mention this aspect here, but overall that becomes a ortholognal
optimization.

%# Abstract Model

%# Sequencer Creation

\end{document}

### 2016-04-28 paper.tex
paper updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-In this section we examine the design space for mapping the CORFU protocol onto
-the RADOS storage system. Recall from Section X the options for rados,
-and from section Y for corfu options. The cross product forms the design space.
-Want to winnow this down.
+In this section we examine the design space for implementing the CORFU
+protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
+design space. A description of each design parameter follows:
+
+{\bf Mapping strategy.} The method by which a log entry---identified by its
+logical position---is addressed within RADOS is referred to as the mapping
+strategy. A 1:1 strategy stores each log entry in a RADOS object with a
+distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
+positions across a set of objects (e.g.  round-robin). While we do consider
+both strategies in this paper, a 1:1 strategy is attractive because it allows
+a design in which clients can directly address log positions by constructing
+the correct object name.
+
+{\bf Storage interface.} entry data may be small or large. its primary
+storage location is important. kv, bs. we further sub-divide bs into
+write and append. each strategy is compatible with fixed size entries
+except for the write interface.
+
+{\bf Logical addressing.}
 
 \subsection{Storage Interface Selection}
+
+Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
+for each of the points in the design space defined in Table \ref{t:init-ds}.
+The results reveal poor relative performance using both the key-value storage
+interface, as well as either of the 1:1 strategies---which incur overhead by
+creating a new object for each log entry. Both of the N:1 strategies using the
+bytestream I/O interface outperform all of the other approaches by over 2x
+throughput, but otherwise have nearly identical performance.
+
+To examine the read performance of the points in the design space we generated
+a dataset using each of the write workloads, and then issued random reads
+across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
+throughput for each of the design strategies. The results show that an N:1
+mapping using bytestream interface performs best. Interestingly, the 1:1
+mapping strategy using the bytestream interface exhibits good read performance
+in comparison to writes using the same strategy.
 
@@
 
-Constructing a complete implementation for each configuration would take
-a lot of time, and may be time wasted if some configurations can be shown
-to never be a viable option.
-
-To winnow down the design space we ran a set of benchmarks---one for each
-configuration--- that stressed the raw write performance using the built-in,
-optimized interfaces on an unmodified version of Ceph. We explicitly avoided
-simulating the costs associated with indexing or protocol enforcement in order
-to form a clean baseline. Log entries may vary in size depending on the
-application, but per-entry metadata is constant.
-
 \begin{figure}[h]
   \centering
-  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.png}
+  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
   \caption{
@@
 \end{figure}
-
-The first thing to notice is that the configurations using a 1:1 mapping
-perform poorly relative to the other configurations. This is unfortunate
-because a 1:1 mapping simplifies data management by avoiding the need to
-multiplex entries onto a smaller set of objects. However, even if performance
-was acceptable, notice that throughput dramatically drops in the tail when the
-number of objects reaches a threshold forcing the object cache to spill to
-disk. Note that in a large cluster this mode may take a long time reach,
-but we avoid it due to the inherent risk.
-
-The KV/N:1 is a convenient configuration because the key-value interface
-can be used to easily address log entries and manage index data. However, the
-performance of KV/N:1 can vary widely depending on the size of the data entry.
-As we will see next, even though the KV/N:1 configuration simplifies data
-management, small entry performance falls behind that of N:1 configurations
-that store data in the stream interface.
 
@@
 
-Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
-for each of the points in the design space defined in Table \ref{t:init-ds}.
-The results reveal poor relative performance using both the key-value storage
-interface, as well as either of the 1:1 strategies---which incur overhead by
-creating a new object for each log entry. Both of the N:1 strategies using the
-bytestream I/O interface outperform all of the other approaches by over 2x
-throughput, but otherwise have nearly identical performance. As a first
-approximation these results show that when optimizing for log append throughput
-the design space should be refined by limiting solutions to architectures based
-on a N:1 mapping strategy using the bytestream I/O interface.
-
-To examine the read performance of the points in the design space we generated
-a dataset using each of the write workloads, and then issued random reads
-across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
-throughput for each of the design strategies. The results show that an N:1
-mapping using bytestream interface performs best. Interestingly, the 1:1
-mapping strategy using the bytestream interface exhibits good read performance
-in comparison to writes using the same strategy.
-
-The configurations based on bytestream storage exhibit the best performance.
-The *write* variation depends on client-side mapping to object offset and
-indirectly on the VFS index layer within the OSD, but this method is
-incompatible with non-fixed size entries. Alternatively, the *append* variation
-supports both fixed and dynamic size entries, but leaves the indexing and
-protocl enforcment challenges unsolved.
-
-In conclusion, both *write* and *append* variation of bytestream configurations
-are the best choice, with an edge given to *append* if it can support dynamic
-sized entries without sacrificing performance.
+As a first approximation these results show that when optimizing for log
+append and random read throughput the design space should be refined by
+limiting solutions to architectures based on a N:1 mapping strategy using the
+bytestream I/O interface. But what these microbenchmarks explicitly omit are
+the overheads associated with metadata management such as enforcing write-once
+semantics or using an index to map a log position to a physical offset within
+an object. In the next section we
 

### 2016-05-13 paper.tex
paper: fix build
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
             colorlinks=true,
-            citecolor=blue
+            citecolor=blue,
             urlcolor=blue,
@@
 
+Cloud providers are increasingly reliant upon extensions to distributed
+storage system such as Ceph as is shown in Figure~\ref{fig:objclass-dev}
+by a marked increase in production Ceph extensions. The construction of
+extensions is not limited to core developers; the community has been receptive
+to contributions with the recent inclusion by CERN developers of an extension
+for performing limited numeric operations on object data~\cite{cls_numops}.
+
 \begin{figure}[h]
@@
 }
+\label{fig:objclass-dev}
 \end{figure}
@@
 
+\bibliography{paper}
+\bibliographystyle{plain}
+
 \end{document}

### 2016-05-13 paper.tex
paper: notes
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 storage system such as Ceph as is shown in Figure~\ref{fig:objclass-dev}
-by a marked increase in production Ceph extensions. The construction of
-extensions is not limited to core developers; the community has been receptive
-to contributions with the recent inclusion by CERN developers of an extension
-for performing limited numeric operations on object data~\cite{cls_numops}.
+by a marked increase in production Ceph extensions.
+
+Figure~\ref{fig:objclass-dev} also shows the number of lines of C++
+code written to create these extensions.
+
+The construction of extensions is not limited to core developers; the
+community has been receptive to contributions with the recent inclusion by
+CERN developers of an extension for performing limited numeric operations on
+object data~\cite{cls_numops}, and a mechanism for dynamically defining
+extensions using Lua~\cite{cls_lua}.
+
+Operating systems and containers are similar. AUFS, dockerfile
+
+\section{Recurring Problem}
+
+Describe CORFU briefly, and then show the physical design challenge. The
+performance decision changes depending on the mix of hardware and software
+versions and configurations.
 

### 2016-05-13 paper.tex
paper: notes
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-04-28 paper.tex
@@
 
-\begin{figure}[h]
+Cloud providers are increasingly reliant upon application-specific extensions
+to distributed storage system such as Ceph as is shown in
+Figure~\ref{fig:objclass-dev} by a marked increase in Ceph extensions used in
+production.
+
+The construction of extensions is not limited to core developers; the
+development community has been receptive to contributions with the recent
+inclusion by CERN developers of an extension for performing limited numeric
+operations on object data~\cite{cls_numops}.
+
+This trend is not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
+KV-Drives, and Rhea, are all recent examples of storage systems embracing the
+power of including application-specific semantics. While Ceph has not yet
+reached the point of directly exposing these features to users, the recent
+inclusion of a mechanism for dynamically defining extensions using
+Lua~\cite{cls_lua} suggests that aspects of this may soon appear.
+
+If the growth illustrated in Figure~\ref{fig:objclass-dev} continues, the
+amount of low-level C++ is likely to grow very larage. Unfortunately, much of
+this code is written assuming a performance profile defined by a combination
+of the hardware and software versions available at the time of development. As
+Ceph development continues at a fast pace, the cost of adapting to changing
+assumptions regarding performance will become significant.
+
+% footnote: we could probably look through the cls_* changelog and find
+% examples of changes based on changes to semantics.
+
+% Virtualization. Operating systems and containers are similar. AUFS, dockerfile
+
+In order to avoid this we need declarative magic hats.
+
+\section{Recurring Problem}
+
+Describe CORFU briefly, and then show the physical design challenge. The
+performance decision changes depending on the mix of hardware and software
+versions and configurations.
+
+\section{Storage System Programmability}
+
+\begin{figure}[ht]
   \centering
@@
 }
+\label{fig:objclass-dev}
 \end{figure}
-
-\section{Storage System Programmability}
 
@@
 management (e.g. a new data layout).
-
-\begin{table}
-\begin{tabular}{|l|l|l|}
-\hline
-Category & Specialization & Methods \\ \hline
-\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
-                         & Exclusive & \\ \hline
-\multirow{3}{*}{Logging} & Replica & 3 \\
-                         & State & 4 \\
-                         & Timestamped & 4 \\ \hline
-Garbage Collection & Reference Counting & 4 \\ \hline
-\multirow{4}{*}{Metadata Management} & RBD & 37 \\
- & RGW & 27 \\
- & User & 5 \\
- & Version & 5 \\ \hline
-\end{tabular}
-\caption{A variety of RADOS object storage classes exist that expose reusable interfaces to applications.}
-\label{tab:objclasses}
-\end{table}
 
@@
 Differences between built-in interfaces and custom cls interfaces.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.2cm,semithick]
-%\tikzstyle{every state}=[fill=red,draw=none,text=white]
-
-  \node[initial left,state] (A)              {$EG$};
-  \node[state]         (B) [below right of=A] {$CP$};
-  \node[state]         (D) [right of=B]       {$IO$};
-  \node[state]         (C) [above right of=A] {$GC$};
-  \node[state]         (E) [right of=C]       {$US$};
-  \node[state]         (F) [below right of=E] {$Fi$};
-
-  \path (A) edge        node [left] {R,W,F} (B)
-            edge        node {T}     (C)
-        (B) edge        node {F}     (C)
-            edge        node {R,W}   (D)
-        (C) edge        node {F,T}   (E)
-        (D) edge        node {W}     (E)
-            edge        node {R}     (F)
-        (E) edge        node {F,T,W} (F);
-\end{tikzpicture}
-\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
-trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
-and update state ($US$) access metadata. The I/O performs a log entry read or
-write, and garbage collection ($GC$) marks entries for reclamation.}
-\end{figure}
 
@@
 in comparison to writes using the same strategy.
+
+\begin{table}
+\begin{tabular}{|l|l|l|}
+\hline
+Category & Specialization & Methods \\ \hline
+\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
+                         & Exclusive & \\ \hline
+\multirow{3}{*}{Logging} & Replica & 3 \\
+                         & State & 4 \\
+                         & Timestamped & 4 \\ \hline
+Garbage Collection & Reference Counting & 4 \\ \hline
+\multirow{4}{*}{Metadata Management} & RBD & 37 \\
+ & RGW & 27 \\
+ & User & 5 \\
+ & Version & 5 \\ \hline
+\end{tabular}
+\caption{A variety of RADOS object storage classes exist that expose reusable interfaces to applications.}
+\label{tab:objclasses}
+\end{table}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.2cm,semithick]
+%\tikzstyle{every state}=[fill=red,draw=none,text=white]
+
+  \node[initial left,state] (A)              {$EG$};
+  \node[state]         (B) [below right of=A] {$CP$};
+  \node[state]         (D) [right of=B]       {$IO$};
+  \node[state]         (C) [above right of=A] {$GC$};
+  \node[state]         (E) [right of=C]       {$US$};
+  \node[state]         (F) [below right of=E] {$Fi$};
+
+  \path (A) edge        node [left] {R,W,F} (B)
+            edge        node {T}     (C)
+        (B) edge        node {F}     (C)
+            edge        node {R,W}   (D)
+        (C) edge        node {F,T}   (E)
+        (D) edge        node {W}     (E)
+            edge        node {R}     (F)
+        (E) edge        node {F,T,W} (F);
+\end{tikzpicture}
+\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
+trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
+and update state ($US$) access metadata. The I/O performs a log entry read or
+write, and garbage collection ($GC$) marks entries for reclamation.}
+\end{figure}
 

### 2016-05-13 paper.tex
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-This trend is not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
+This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
 KV-Drives, and Rhea, are all recent examples of storage systems embracing the
@@
 If the growth illustrated in Figure~\ref{fig:objclass-dev} continues, the
-amount of low-level C++ is likely to grow very larage. Unfortunately, much of
+amount of low-level C++ will grow large. Unfortunately, much of
 this code is written assuming a performance profile defined by a combination
@@
 
-In order to avoid this we need declarative magic hats.
+Previous work in active storage has focused on the utility of moving
+computation closer to data in an effort to reduce data movement and take
+advantage of cpu and i/o parallelism, and gave little to no attention to
+security or performance isolation concerns. In contrast, the interfaces being
+deployed today focus on data management, indexing, and physical format.
+
+We propose the introduction of a declarative language for building object
+based interfaces that allows a storage system to meet the needs of an
+application throughout the development process without requiring rewrites.
 
 \section{Recurring Problem}
+
+Lets motivate here. Showing the need for programmability and some 
+declarative interface.
+
+Describe corfu and object class interfaces.
 
@@
 versions and configurations.
+
+Use state diagram to show how this works for corfu.
+
+then show physical design crazy ness as we go to different
+versions and/or different hardware.
+
+\section{Programming Model}
+
+Here we create a programming model for the objects
+
+Show how the corfu interface is defined in this model.
+
+\section{Other Interfaces}
+
+\section{Evaluation}
 
@@
 Map & I/O & Entry Size & Addressing & Metadata \\ \hline
-\multirow{2}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
+\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
                      & BS  & Flex     & Ceph/VFS  & KV/BS \\ \hline

### 2016-05-13 paper.tex
paper: working on outline
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \usepackage{multirow}
+\usepackage{subcaption}
 \usepackage{graphicx,grffile}
@@
 
-\title{title}
+\title{outline!}
 
@@
 \begin{abstract}
-abstract
 \end{abstract}
@@
 
-We propose the introduction of a declarative language for building object
-based interfaces that allows a storage system to meet the needs of an
-application throughout the development process without requiring rewrites.
-
-\section{Recurring Problem}
-
-Lets motivate here. Showing the need for programmability and some 
-declarative interface.
-
-Describe corfu and object class interfaces.
-
-Describe CORFU briefly, and then show the physical design challenge. The
-performance decision changes depending on the mix of hardware and software
-versions and configurations.
-
-Use state diagram to show how this works for corfu.
-
-then show physical design crazy ness as we go to different
-versions and/or different hardware.
-
-\section{Programming Model}
-
-Here we create a programming model for the objects
-
-Show how the corfu interface is defined in this model.
-
-\section{Other Interfaces}
-
-\section{Evaluation}
-
-\section{Storage System Programmability}
-
 \begin{figure}[ht]
   \centering
-    \includegraphics[width=0.45\textwidth]{experiments/objclass-dev/output.png}
+    \includegraphics[width=0.48\textwidth]{experiments/objclass-dev/output.png}
     \caption{
@@
 
-When application goals are not met by a storage system the most common reaction
-is to design a workaround. Workarounds roughly fall into one of two categories:
-so called ``bolt-on'' services that introduce a 3rd party system (e.g. a
-metadata service), or expanded application responsibility in the form of data
-management (e.g. a new data layout).
+We propose the introduction of a declarative language for building object
+based interfaces that allows a storage system to meet the needs of an
+application throughout the development process without requiring rewrites.
 
-\section{Consolidated Section}
+\section{Background}
 
-Up until this point we have introduced Ceph, CORFU, the general problems
-associated with programmability, described ZLog as being an instantiation of
-the CORFU interface on top of Ceph.
+In this section we briefly describe Ceph and object-based interface
+development. We will use CORFU as a motivating example of an interface, and
+then show the challenges with the current low-level mechanisms for defining
+interfaces.
 
-Template for the CORFU object class interface. Maybe a simple state machine.
+\subsection{Domain-specific Object Interfaces}
 
-Describe indexing and protocol enforcement.
-Differences between built-in interfaces and custom cls interfaces.
-
-\section{ZLog Physical Design}
-
-In this section we examine the design space for implementing the CORFU
-protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
-design space. A description of each design parameter follows:
-
-{\bf Mapping strategy.} The method by which a log entry---identified by its
-logical position---is addressed within RADOS is referred to as the mapping
-strategy. A 1:1 strategy stores each log entry in a RADOS object with a
-distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
-positions across a set of objects (e.g.  round-robin). While we do consider
-both strategies in this paper, a 1:1 strategy is attractive because it allows
-a design in which clients can directly address log positions by constructing
-the correct object name.
-
-{\bf Storage interface.} entry data may be small or large. its primary
-storage location is important. kv, bs. we further sub-divide bs into
-write and append. each strategy is compatible with fixed size entries
-except for the write interface.
-
-{\bf Logical addressing.}
-
-\subsection{Storage Interface Selection}
-
-Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
-for each of the points in the design space defined in Table \ref{t:init-ds}.
-The results reveal poor relative performance using both the key-value storage
-interface, as well as either of the 1:1 strategies---which incur overhead by
-creating a new object for each log entry. Both of the N:1 strategies using the
-bytestream I/O interface outperform all of the other approaches by over 2x
-throughput, but otherwise have nearly identical performance.
-
-To examine the read performance of the points in the design space we generated
-a dataset using each of the write workloads, and then issued random reads
-across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
-throughput for each of the design strategies. The results show that an N:1
-mapping using bytestream interface performs best. Interestingly, the 1:1
-mapping strategy using the bytestream interface exhibits good read performance
-in comparison to writes using the same strategy.
+Ceph allows the construction of domain specific interfaces. There are many
+different object interfaces in production today. An example of the spectrum of
+interface types is shown in Table~\ref{tab:objclass-cats}. Object classes are
+developed as a transactional composition of native storage interfaces.
 
@@
                          & Timestamped & 4 \\ \hline
-Garbage Collection & Reference Counting & 4 \\ \hline
-\multirow{4}{*}{Metadata Management} & RBD & 37 \\
+GC & Reference Counting & 4 \\ \hline
+\multirow{4}{*}{Metadata} & RBD & 37 \\
  & RGW & 27 \\
@@
 \end{tabular}
-\caption{A variety of RADOS object storage classes exist that expose reusable interfaces to applications.}
-\label{tab:objclasses}
+\caption{A variety of RADOS object storage classes exist that expose reusable
+interfaces to applications.}
+\label{tab:objclass-cats}
 \end{table}
+
+\subsection{Motivating Example: CORFU Log}
+
+Short description of CORFU system~\cite{balakrishnan:nsdi12}. It's a log!
+While CORFU relies on custom interfaces built into flash devices, an
+implementation in Ceph relies upon custom device interfaces to enforce its
+protocol.
+
+The state-machine shown in Figure~\ref{fig:corfu-sm} shows the composition of
+actions for each component of the CORFU interface. For example, each
+operations begins with an \emph{epoch guard}, and a \emph{write} operation
+will consult an index and perform a bulk I/O operation.  The primary concern
+when mapping an interface onto Ceph is deciding how each native interface is
+used in the interface composition.
 
@@
 write, and garbage collection ($GC$) marks entries for reclamation.}
+\label{fig:corfu-sm}
 \end{figure}
+
+\subsection{Physical Design}
+
+There is a finite set of strategies for mapping the CORFU interfaces onto the
+current set of native I/O interfaces used to construct object interfaces.
+Table~\ref{tab:pd-map} shows the design space for mapping the CORFU interfaces
+onto Ceph. In order to select the best mapping we performed a parameter sweet
+over the design space.
+
+\begin{figure*}[!ht]
+    \centering
+    \begin{subfigure}{.6\columnwidth}
+        \includegraphics[width=\columnwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
+        \caption{a}
+        \label{fig:vanilla-io-diff}
+    \end{subfigure}\hfill
+    \begin{subfigure}{.6\columnwidth}
+        \includegraphics[width=\columnwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
+        \caption{b}
+    \end{subfigure}\hfill
+    \begin{subfigure}{.6\columnwidth}
+        \includegraphics[width=\columnwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+        \caption{c}
+    \end{subfigure}\hfill
+    \caption{Shown here are the graphs and such that demonstrate that the same
+        physical design choices are not the same between differing version of
+    Ceph even on the same hardware.}
+\end{figure*}
+
+Figure~\ref{fig:vanilla-io-diff} shows the result of a parameter sweep which
+clearly shows that an N-1 mapping based on the bytestream interface provides
+superiour performance. However, the other two graphs show the same interfaces
+running on an old version of Ceph that show the same decision would not have
+been the optimal choice {\bf \emph{note: for the outline these are not the
+real graphs we'll be using}}.
 
@@
 the RADOS object storage system.}
-\label{t:init-ds}
+\label{tab:pd-map}
 \end{table}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
-  \caption{
-[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
-Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
-interfaces in various configurations. The best performance is achieved using
-the byte stream interface and a N:1 mapping strategy.
-}
-\end{figure}
+\section{Programming Model}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
-  \caption{
-[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
-Throughput of 1K random reads to a single OSD using the standard I/O
-interfaces in various configurations. The best performance is achieved using
-the byte stream interface and a N:1 mapping strategy.
-}
-\end{figure}
+In this section we are going to be describing our way of creating interfaces.
 
-As a first approximation these results show that when optimizing for log
-append and random read throughput the design space should be refined by
-limiting solutions to architectures based on a N:1 mapping strategy using the
-bytestream I/O interface. But what these microbenchmarks explicitly omit are
-the overheads associated with metadata management such as enforcing write-once
-semantics or using an index to map a log position to a physical offset within
-an object. In the next section we
+\section{Other Interfaces}
 
-\subsection{Metadata Management}
+Here we show the derivation of two other interfaces that are in production in
+Ceph today to demonstrate the generality of our interface.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-  \caption{
-[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
-}
-\end{figure}
+\section{Evaluation}
 
-Many variations of design here. What is the easiest way to 
-winnow down the design space.
+\section{Related Work}
 
-NEXT GRAPH A
+\section{Conclusion}
 
-This is an easy interface to simulate.
-- Check Epoch (omap, stream header)
-- this graph also has the APPEND raw through for comparison
-- adds NEW OSD OP TO AVOID CLS OVERHEAD DISCUSSION FOR PAPER
-- 3 runs in this graph, no need to write interface
-- short runs to verify and then do all long runs togther with next graph
+%In this section we examine the design space for implementing the CORFU
+%protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
+%design space. A description of each design parameter follows:
+%
+%{\bf Mapping strategy.} The method by which a log entry---identified by its
+%logical position---is addressed within RADOS is referred to as the mapping
+%strategy. A 1:1 strategy stores each log entry in a RADOS object with a
+%distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
+%positions across a set of objects (e.g.  round-robin). While we do consider
+%both strategies in this paper, a 1:1 strategy is attractive because it allows
+%a design in which clients can directly address log positions by constructing
+%the correct object name.
+%
+%{\bf Storage interface.} entry data may be small or large. its primary
+%storage location is important. kv, bs. we further sub-divide bs into
+%write and append. each strategy is compatible with fixed size entries
+%except for the write interface.
+%
+%{\bf Logical addressing.}
 
-Both protocol enforcement and logical address translation
-represent a larger design space involving index design. However
-since in either case all state must be stored in the omap and
-in the stream header, testing the overhead of epoch enforcement
-provides a strict upper bound on performance.
+%\subsection{Storage Interface Selection}
+%
+%Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
+%for each of the points in the design space defined in Table \ref{t:init-ds}.
+%The results reveal poor relative performance using both the key-value storage
+%interface, as well as either of the 1:1 strategies---which incur overhead by
+%creating a new object for each log entry. Both of the N:1 strategies using the
+%bytestream I/O interface outperform all of the other approaches by over 2x
+%throughput, but otherwise have nearly identical performance.
+%
+%To examine the read performance of the points in the design space we generated
+%a dataset using each of the write workloads, and then issued random reads
+%across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
+%throughput for each of the design strategies. The results show that an N:1
+%mapping using bytestream interface performs best. Interestingly, the 1:1
+%mapping strategy using the bytestream interface exhibits good read performance
+%in comparison to writes using the same strategy.
 
-- Protocol Enforcement
-- Logical Translation (only append case)
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
+%  \caption{
+%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
+%Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
+%interfaces in various configurations. The best performance is achieved using
+%the byte stream interface and a N:1 mapping strategy.
+%}
+%\end{figure}
+%
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
+%  \caption{
+%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
+%Throughput of 1K random reads to a single OSD using the standard I/O
+%interfaces in various configurations. The best performance is achieved using
+%the byte stream interface and a N:1 mapping strategy.
+%}
+%\end{figure}
+%
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+%  \caption{
+%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
+%}
+%\end{figure}
 
-Memory Cache
-
-NEXT GRAPH B (or combined with A)
-Use in-memory cache to implement the epoch check and show that we can match
-the raw performance even with this added functionality.
-This will show the content from graph A for comparison. Look
-we match the raw performance.
-
-NEXT GRAPH B or C (probably just two).. bar chart?
- Final consolidation of average throughputs of all the different
-methods
- THIS INCLUDES THE FULL IMPL OF ZLOG APPEND using the in-memory cache
-
-We solve the slow down of the epoch by storing the value in memory. Here is
-the new performance.  We are able to with the epoch check match the raw
-performance.
-
-We know that we cannot use kv/bs for logicla address translation or protocol
-enforcement without a performance hit so there is no point in trying. instead,
-we generalize this memory cache hting used for epoch checks as a
-new interface and use this interface to implement the full zlog
-stack and show the performance.
-
-Overall corfu is relatively simple abtraction, yet requires significant
-work in making performant. Further, the decision may change over time.
-
-We will probably have room in the paper to discuss the design of the
-index structure, like collapsing a bit map and taking advantage of
-the implicit strides in the offsets, etc... but we can probably move
-that to a later section to keep this presentation concise. We should
-probably mention this aspect here, but overall that becomes a ortholognal
-optimization.
-
-%# Abstract Model
-
-%# Sequencer Creation
+%As a first approximation these results show that when optimizing for log
+%append and random read throughput the design space should be refined by
+%limiting solutions to architectures based on a N:1 mapping strategy using the
+%bytestream I/O interface. But what these microbenchmarks explicitly omit are
+%the overheads associated with metadata management such as enforcing write-once
+%semantics or using an index to map a log position to a physical offset within
+%an object. In the next section we
 

### 2016-05-15 experiments/librados-sweep/run.sh
expr: add firefly short test
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -3,14 +3,15 @@
 reset=soft
 data_dev=sdc
 
-name=librados_sweep
-runtime=7200
+name=vanilla_firefly
+runtime=1800
 pg_nums="128"
 stripe_widths="32"
 queue_depths="32"
 entry_sizes="1024"
 pool=zlog
 rest=120
+ceph_version=firefly
 
 # workloads
 wl_11="map_11 bytestream_11"

### 2016-05-15 paper.tex
paper: corfu desc
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{Motivating Example: CORFU Log}
+\subsection{Motivating Example: CORFU}
+
+The primary motivating example we will use in this paper is the CORFU
+distributed shared-log designed to provide high-performance serialization
+across a set of flash storage devices~\cite{balakrishnan:nsdi12}. The
+shared-log is a powerful abstraction useful when building distributed systems
+and applications, but common implementations such as Paxos or Raft funnel I/O
+through a single node limiting the throughput of log
+operations~\cite{lamport:tocs89}. The CORFU protocol addresses this limitation
+by de-coupling log entry storage from log metadata, making use of a
+centeralized, volatile, in-memory \emph{sequencer service} that assigns log
+positions to clients that are appending to the log. Since the sequencer is
+centeralized serialization is trivial, and the use of non-durable state allows
+the sequencer service to operate at very high rates.
+
+While a full description of the CORFU protocol is beyond the scope of this
+paper, it is important to highlight the usage of application-specific flash
+I/O interfaces in CORFU that are used to guarantee correctness during failure.
+Each flash device in CORFU exposes a 64-bit write-once address space
+consisting of the primary interfaces \emph{write(pos, data)} and
+\emph{read(pos)}, as well as \emph{fill(pos)} and \emph{trim(pos)} that
+invalidate and reclaim log entries, respectively. All I/O operations initiated
+by clients are tagged with an \emph{epoch} value, and flash devices are
+expected to reject client requests requests tagged with an old epoch value. To
+facilitate recovery or handle system reconfiguration in CORFU, the storage
+devices are also required to support a \emph{seal(epoch)} command that marks
+the latest epoch and returns the maximum position written to that device. The
+seal interface is used following the failure of a sequencer to calculate the
+tail of the log that the sequencer should use to repopulate its in-memory
+state.
+
+why ceph and corfu?
 

### 2016-05-15 experiments/librados-sweep/run.sh
expr: add firefly long
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -4,7 +4,7 @@ reset=soft
 data_dev=sdc
 
 name=vanilla_firefly
-runtime=1800
+runtime=3600
 pg_nums="128"
 stripe_widths="32"
 queue_depths="32"

### 2016-05-15 paper.tex

@@
 
-why ceph and corfu?
+\paragraph*{Programmable Storage and CORFU}
+Two aspects of CORFU make its design attractive in the context of
+the Ceph storage system. First, CORFU assumes a cluster of flash devices
+because log-centric systems tend to have a larger percentage of random reads
+making it difficult to achieve high-performance with spinning disks. However,
+the speed of the underlying storage does not affect correctness. Thus, in a
+software-defined storage system such as Ceph a single implementation can
+transparently take advantage of any software or hardware upgrades, and make
+use of data management features such as tiering in RADOS, allowing users to
+freely choose between media types such as SSD, spinning disks, or emerging
+NVRAM technologies.
 
-Short description of CORFU system~\cite{balakrishnan:nsdi12}. It's a log!
-While CORFU relies on custom interfaces built into flash devices, an
-implementation in Ceph relies upon custom device interfaces to enforce its
-protocol.
+The second property of CORFU relevant in the context of Ceph is the dependency
+CORFU places on the storage device interface.  As was previously discussed, in
+order to enforce constraints defined within the CORFU protocol, storage
+devices are expected to expose a domain-specific interface specially designed
+for the protocol. While the authors of the CORFU paper describe both a
+host-based and FPGA-based solution, RADOS directly supports the concept of
+custom storage interfaces to its objects that can be used to achieve the
+transactional semantics necessary to implement the CORFU storage interfaces.
+This feature is called RADOS object classes, and is one of the most powerful
+features offered by RADOS for designing storage applications.
 
+\paragraph*{A CORFU Object Interface}
 The state-machine shown in Figure~\ref{fig:corfu-sm} shows the composition of

### 2016-05-16 paper.tex
paper: intro and diagram updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \usepackage{multirow}
+\usepackage{tabularx}
 \usepackage{subcaption}
@@
 
-Cloud providers are increasingly reliant upon application-specific extensions
-to distributed storage system such as Ceph as is shown in
-Figure~\ref{fig:objclass-dev} by a marked increase in Ceph extensions used in
-production.
+Storage systems are increasingly providing features that take advantage of
+application-specific knowledge to achieve optimizations and provide unique
+services. However, this trend is leading to the creation of a large amount of
+low-level software extensions that will be difficult to maintain as system
+software and hardware evolve.
 
-The construction of extensions is not limited to core developers; the
-development community has been receptive to contributions with the recent
-inclusion by CERN developers of an extension for performing limited numeric
-operations on object data~\cite{cls_numops}.
+The widely deployed Ceph distributed storage system is an example of a storage
+system that supports application-specific extensions in the form of custom I/O
+interfaces to objects within the underlying RADOS object storage
+system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly reliant
+upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by a marked
+increase in the number of object operations that are packaged are part of the
+Ceph distribution.  In addition to the growth of the quantity of operations,
+Figure~\ref{fig:objclass-dev} also depicts the amount of low-level C++ written
+to implement the set object operations. Unfortunately, this code is written
+assuming a performance profile defined by the combination of the hardware and
+software versions available at the time of development. As Ceph continues to
+evolve at a fast pace, the cost of adapting application-specific codes to
+changing assumptions regarding performance may become significant.
 
@@
 KV-Drives, and Rhea, are all recent examples of storage systems embracing the
-power of including application-specific semantics. While Ceph has not yet
-reached the point of directly exposing these features to users, the recent
-inclusion of a mechanism for dynamically defining extensions using
-Lua~\cite{cls_lua} suggests that aspects of this may soon appear.
-
-If the growth illustrated in Figure~\ref{fig:objclass-dev} continues, the
-amount of low-level C++ will grow large. Unfortunately, much of
-this code is written assuming a performance profile defined by a combination
-of the hardware and software versions available at the time of development. As
-Ceph development continues at a fast pace, the cost of adapting to changing
-assumptions regarding performance will become significant.
+power of including application-specific semantics.
 
@@
 
+And the construction of object operations is not limited to core Ceph
+developers; the development community has been receptive to contributions with
+the recent inclusion by CERN developers of an extension for performing limited
+numeric operations on object data~\cite{cls_numops}, and all 
+While Ceph has not yet
+reached the point of directly exposing these features to users, the recent
+inclusion of a mechanism for dynamically defining extensions using
+Lua~\cite{cls_lua} suggests that aspects of this may soon appear.
+
 Ceph allows the construction of domain specific interfaces. There are many
@@
 \begin{table}
-\begin{tabular}{|l|l|l|}
+\centering
+\begin{tabularx}{\columnwidth}{|X|l|l|l|}
 \hline
@@
                          & Timestamped & 4 \\ \hline
-GC & Reference Counting & 4 \\ \hline
+Garbage Collection & Ref. Counting & 4 \\ \hline
 \multirow{4}{*}{Metadata} & RBD & 37 \\
@@
  & Version & 5 \\ \hline
-\end{tabular}
+\end{tabularx}
 \caption{A variety of RADOS object storage classes exist that expose reusable
@@
     \begin{subfigure}{.6\columnwidth}
-        \includegraphics[width=\columnwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
+        \includegraphics[width=\columnwidth]{experiments/librados-sweep/jewel_ssd.png}
         \caption{a}
@@
     \begin{subfigure}{.6\columnwidth}
-        \includegraphics[width=\columnwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
+        \includegraphics[width=\columnwidth]{experiments/librados-sweep/firefly_ssd.png}
         \caption{b}

### 2016-05-16 experiments/librados-sweep/run.sh
expr: add disk/ssd firefly
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,9 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
+data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 
-name=vanilla_firefly
+name=vanilla_firefly_disk_ssd
 runtime=3600
 pg_nums="128"
 stripe_widths="32"

### 2016-05-16 experiments/librados-sweep/run.sh
expr: add disk expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -2,10 +2,9 @@
 
 reset=soft
 data_dev=sdb
-journal_dev=sdc
-noop_dev=sdc
+#noop_dev=yes
 
-name=vanilla_firefly_disk_ssd
+name=vanilla_firefly_disk
 runtime=3600
 pg_nums="128"
 stripe_widths="32"

### 2016-05-16 experiments/multi-node/run.sh
expr: local vs. remote sequencers - appends and default workloads
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -4,4 +4,4 @@ set -ex
 
 ANSIBLE_LOG_PATH="logs/ceph-zeqr.log" \
 ansible-playbook --forks 50 --skip-tags "with_pkg" \
-ceph.yml workloads/unit-tests.yml
+ceph.yml workloads/bench-ng.yml

### 2016-05-16 paper.tex
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \begin{abstract}
+To meet the needs of a diverse and growing set of cloud-based applications,
+modern distributed storage frameworks expose a variety of composable
+subsystems as building blocks.  This approach gives infrastructure programmers
+significant flexibility in implementing application-specific semantics while
+reusing trusted components.  Unfortunately, in current storage systems the
+composition of subsystems is a low-level task that couples (and hence
+obscures) a variety of orthogonal concerns, including functional correctness
+and performance.  Building an application by wiring together a collection of
+components typically requires thousands of lines of carefully-written C++
+code, an effort that must be repeated whenever device or subsystem
+characteristics change.
+
+In this paper, we propose a declarative approach to subservice composition
+that allows programmers to focus on the high-level functional properties that
+are required by applications.  Choosing an implementation that is consistent
+with the declarative functional specification then can be posed as a search
+problem over the space of parameters such as block sizes, storage interfaces
+(e.g. key/value or block storage) and concurrency control mechanisms.  We
+present experimental evaluation of our prototype, (etc etc)
 \end{abstract}
@@
 
+% whats the problem
 Storage systems are increasingly providing features that take advantage of
 application-specific knowledge to achieve optimizations and provide unique
-services. However, this trend is leading to the creation of a large amount of
-low-level software extensions that will be difficult to maintain as system
-software and hardware evolve.
+services. However, this trend is leading to the creation of a large number of
+software extensions that will be difficult to maintain as system software and
+hardware continue to evolve.
 
+% why is it interesting
 The widely deployed Ceph distributed storage system is an example of a storage
 system that supports application-specific extensions in the form of custom I/O
-interfaces to objects within the underlying RADOS object storage
+interfaces to objects managed by the underlying RADOS object storage
 system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly reliant
 upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by a marked
-increase in the number of object operations that are packaged are part of the
-Ceph distribution.  In addition to the growth of the quantity of operations,
-Figure~\ref{fig:objclass-dev} also depicts the amount of low-level C++ written
-to implement the set object operations. Unfortunately, this code is written
-assuming a performance profile defined by the combination of the hardware and
-software versions available at the time of development. As Ceph continues to
-evolve at a fast pace, the cost of adapting application-specific codes to
-changing assumptions regarding performance may become significant.
+increase in the number of object operations that are packaged as part of the
+Ceph distribution and widely used by internal Ceph subystems and by
+applications such as OpenStack Swift and Cinder~\cite{openstack}. In addition
+to the growth in the quantity of operations, Figure~\ref{fig:objclass-dev}
+also depicts the amount of low-level C++ written to implement these
+operations. Unfortunately, this code is written assuming a performance profile
+defined by the combination of the hardware and software versions available at
+the time of development. As Ceph continues to evolve at a fast pace, the cost
+of adapting application-specific codes to changing assumptions regarding
+performance may become significant.
 
-This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
-KV-Drives, and Rhea, are all recent examples of storage systems embracing the
-power of including application-specific semantics.
+%The construction of object operations is not limited to core Ceph developers;
+%the development community has been receptive to contributions with the recent
+%inclusion by CERN developers of an extension for performing limited numeric
+%operations on object data~\cite{cls_numops}. And while Ceph has not yet
+%reached the point of directly exposing these features to non-administrative
+%users, the recent inclusion of a mechanism for dynamically defining extensions
+%using Lua~\cite{cls_lua} suggests that aspects of this feature may soon appear.
 
-% footnote: we could probably look through the cls_* changelog and find
-% examples of changes based on changes to semantics.
+% a challenge is mapping application semantics onto these objects
 
-% Virtualization. Operating systems and containers are similar. AUFS, dockerfile
+%Proprietary storage systems are prevalent, but have fixed interfaces that are
+%developed in response to market demand. Alternative storage interfaces are
+%prototyped in research systems.
+%For the first time Ceph is providing a high-performance production quality
+%storage system developed as open-source.
 
-Previous work in active storage has focused on the utility of moving
-computation closer to data in an effort to reduce data movement and take
-advantage of cpu and i/o parallelism, and gave little to no attention to
-security or performance isolation concerns. In contrast, the interfaces being
-deployed today focus on data management, indexing, and physical format.
+% why is it hard / why do naive approaches fail
+%the naive approach here is the use of generic interfaces. one may
+%think that key-value is relatively slower than bytestreams, but by how much.
+%is this a principal we can bet the house on.
+%just building these simple interfaces don't work that well
+%Systems provide low-level native interfaces such as key-value or bytestream interfaces.
+%These abstract interfaces are useful, but they are not the only way to map
+%interfaces onto storage. For instance, a specialized key-value db could be serialized into
+%the bytestream interface.
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
@@
 Growth of officially supported, custom object interfaces in RADOS over 6
-years. A \emph{method} is a specific object interface and a \emph{class} is a
-logical grouping of methods.
-}
+years. An \emph{operation} is a function executed in the context of an object,
+and operations are grouped into different \emph{categories}
+corresponding to applications or utilities, such as reference counting}
 \label{fig:objclass-dev}
@@
 
-We propose the introduction of a declarative language for building object
-based interfaces that allows a storage system to meet the needs of an
-application throughout the development process without requiring rewrites.
+% why hasn't it been solved before / why have previous approaches failed
+%As hardware becomes faster the differences in storage interface performance
+%will continue shift relative to one another.
+%Existing systems fall into the same trap.
+%This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
+%KV-Drives, and Rhea, are all recent examples of storage systems embracing the
+%power of including application-specific semantics.
+%Previous work in active storage has focused on the utility of moving
+%computation closer to data in an effort to reduce data movement and take
+%advantage of cpu and i/o parallelism, and gave little to no attention to
+%security or performance isolation concerns. In contrast, the interfaces being
+%deployed today focus on data management, indexing, and physical format.
+%Database systems solve this problem with a declarative query language, compiler, and
+%optimizer. So should we.
+%
+%We propose the introduction of a declarative language for building object
+%based interfaces that allows a storage system to meet the needs of an
+%application throughout the development process without requiring rewrites.
 
@@
 
-In this section we briefly describe Ceph and object-based interface
-development. We will use CORFU as a motivating example of an interface, and
-then show the challenges with the current low-level mechanisms for defining
-interfaces.
+In this section we highlight the salient components of Ceph, especially its
+\emph{object class} feature that offers users the ability to load and execute
+application-specific codes. We provide a description of our motivating
+example, a high-performance distributed shared-log built upon Ceph that makes
+extensive use of the object class facility, and then we present the challenges
+that application developers face when using this extensibility feature offered
+by the storage system.
 
-\subsection{Domain-specific Object Interfaces}
+\subsection{Ceph and Storage Programmability}
+\label{sec:objclass}
 
-And the construction of object operations is not limited to core Ceph
-developers; the development community has been receptive to contributions with
-the recent inclusion by CERN developers of an extension for performing limited
-numeric operations on object data~\cite{cls_numops}, and all 
-While Ceph has not yet
-reached the point of directly exposing these features to users, the recent
-inclusion of a mechanism for dynamically defining extensions using
-Lua~\cite{cls_lua} suggests that aspects of this may soon appear.
+Figure~\ref{fig:ceph} illustrates the collection of components commonly
+referred to as Ceph. At the bottom, a cluster of 10s--10,000s \emph{object
+storage devices} compose the distributed object storage system called RADOS.
+Widely deployed applications such as the S3/Swift--compliant RADOS Gateway
+(RGW), RADOS Block Device (RBD), and the POSIX Ceph File System are built upon
+the \emph{librados} client layer that presents a fault-tolerant always-on view
+of the RADOS cluster.
 
-Ceph allows the construction of domain specific interfaces. There are many
-different object interfaces in production today. An example of the spectrum of
-interface types is shown in Table~\ref{tab:objclass-cats}. Object classes are
-developed as a transactional composition of native storage interfaces.
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{.48\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{figures/ceph}
+      \caption{Ceph}
+      \label{fig:ceph}
+  \end{subfigure}\quad
+  \begin{subfigure}[b]{.40\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{figures/osd}
+      \caption{OSD}
+      \label{fig:osd}
+  \end{subfigure}
+  \caption{Ceph and OSD... gotta cram more stuff in here...}
+\end{figure}
 
-\begin{table}
+The object storage device (OSD), illustrated in Figure~\ref{fig:osd}, is the
+building block of the RADOS cluster and is responsible for managing and
+providing access to a set of
+named objects. The configuration of an OSD is flexible, and commonly contains
+a mix of commodity hardware such as HDD and SSD bulk storage, a multi-core
+CPU, GBs of RAM, and one or two 10 Gb Ethernet links. Clients access object
+data managed by an OSD by invoking object operations exposed by the OSD
+such as reading or writing bytes, as well as more complex operations like taking
+snapshots or composing one or more native operations into compound procedures
+that execute in a transactional context.
+
+The native object operations in RADOS roughly fall into two categories based
+on the type of data being accessed: key-value items, or bulk bytestream data.
+The key-value interface operates as a dedicated database associated with each
+object, and the bytestream interface supports random byte-level access.  At a
+low-level each of these abstract I/O interfaces map to hardware storage
+devices through a pluggable object backend storage service. For instance,
+LevelDB or RocksDB may be used to store key-value data, while the
+\emph{FileStore} implementation maps the bytestream interface onto a local
+POSIX file system~\cite{leveldb,rocksdb}. Several backend implementations
+exist for storing data in targets such as a Kinetic Drive or in NVMe devices,
+among others.
+
+%Independent of the hardware or backends chosen, Ceph allows all
+%object operations to be combined to create compound operations such as storing
+%key-value metadata associated with a binary blob, allowing applications to
+%build object interfaces that work even as Ceph software and hardware evolve.
+
+\paragraph*{Object Classes}
+While Ceph provides a wide variety of native object operations, it also
+includes a facility referred to as \emph{object classes} that allow developers
+to create application-specific object operations in the form of C++ shared
+libraries dynamically loaded into the OSD process at runtime.  Object classes
+can be used to implement basic data management tasks such as indexing
+metadata, or used to perform complex operations such as data transformations
+or filtering. Table~\ref{tab:objclass-cats} summarizes the range of object
+classes maintained in the upstream Ceph project which support internal Ceph
+subsystems as well as applications and services that run on top of Ceph.
+
+\begin{table}[ht]
 \centering
@@
 
-\subsection{Motivating Example: CORFU}
+A critical step in the development of application-specific object interfaces
+is deciding how to best make use of the native object interfaces. For instance
+if an application stores an image in an object, it may also extract and store
+EXIF metadata as key-value pairs in the object key-value database.  However,
+depending on the application needs it may be sufficient or offer a performance
+advantage to store this metadata as a header within the bytestream. In the remainder
+of this section we will explore the challenges associated with these design questions.
+
+\subsection{Motivating Application: CORFU}
 
@@
 operations~\cite{lamport:tocs89}. The CORFU protocol addresses this limitation
-by de-coupling log entry storage from log metadata, making use of a
+by de-coupling log entry storage from log metadata management, making use of a
 centeralized, volatile, in-memory \emph{sequencer service} that assigns log
@@
 centeralized serialization is trivial, and the use of non-durable state allows
-the sequencer service to operate at very high rates.
+the sequencer service to operate at very high rates. The CORFU system has been
+used to demonstrate a number of interesting services such as transactional
+key-value and metadata services, replicated state machines, and an elastic
+cloud-based database management system~\cite{balakrishnan:sosp13,bernstein:cidr11}.
 
-While a full description of the CORFU protocol is beyond the scope of this
-paper, it is important to highlight the usage of application-specific flash
-I/O interfaces in CORFU that are used to guarantee correctness during failure.
-Each flash device in CORFU exposes a 64-bit write-once address space
-consisting of the primary interfaces \emph{write(pos, data)} and
-\emph{read(pos)}, as well as \emph{fill(pos)} and \emph{trim(pos)} that
-invalidate and reclaim log entries, respectively. All I/O operations initiated
-by clients are tagged with an \emph{epoch} value, and flash devices are
-expected to reject client requests requests tagged with an old epoch value. To
-facilitate recovery or handle system reconfiguration in CORFU, the storage
-devices are also required to support a \emph{seal(epoch)} command that marks
-the latest epoch and returns the maximum position written to that device. The
-seal interface is used following the failure of a sequencer to calculate the
-tail of the log that the sequencer should use to repopulate its in-memory
-state.
+Two aspects of CORFU make its design attractive in the context of the Ceph
+storage system. First, CORFU assumes a cluster of flash devices because
+log-centric systems tend to have a larger percentage of random reads making it
+difficult to achieve high-performance with spinning disks. However, the speed
+of the underlying storage does not affect correctness. Thus, in a
+software-defined storage system such as Ceph a single implementation can
+transparently take advantage of any software or hardware upgrades, and make use
+of existing and future data management features such as tiering in RADOS,
+allowing users to freely choose between media types such as SSD, spinning
+disks, or emerging NVRAM technologies.
 
-\paragraph*{Programmable Storage and CORFU}
-Two aspects of CORFU make its design attractive in the context of
-the Ceph storage system. First, CORFU assumes a cluster of flash devices
-because log-centric systems tend to have a larger percentage of random reads
-making it difficult to achieve high-performance with spinning disks. However,
-the speed of the underlying storage does not affect correctness. Thus, in a
-software-defined storage system such as Ceph a single implementation can
-transparently take advantage of any software or hardware upgrades, and make
-use of data management features such as tiering in RADOS, allowing users to
-freely choose between media types such as SSD, spinning disks, or emerging
-NVRAM technologies.
+\paragraph*{CORFU and Storage Programmability}
+The second property of CORFU relevant in the context of Ceph is the dependency
+CORFU places on custom storage device interfaces used to guarantee
+serialization during failure and reconfiguration. Each flash device in a CORFU
+cluster exposes a 64-bit write-once address space consisting of the primary I/O
+interfaces \emph{write(pos, data)} and \emph{read(pos)} for accessing log
+entries, as well as \emph{fill(pos)} and \emph{trim(pos)} that invalidate and
+reclaim log entries, respectively. All I/O operations in CORFU initiated by
+clients are tagged with an \emph{epoch} value, and flash devices are expected
+to reject client requests that contian an old epoch value. To facilitate
+recovery or handle system reconfiguration in CORFU, the storage devices are
+also required to support a \emph{seal(epoch)} command that marks the latest
+epoch and returns the maximum position written to that device. The seal
+interface is used following the failure of a sequencer to calculate the tail of
+the log that the sequencer should use to repopulate its in-memory state.
 
-The second property of CORFU relevant in the context of Ceph is the dependency
-CORFU places on the storage device interface.  As was previously discussed, in
-order to enforce constraints defined within the CORFU protocol, storage
-devices are expected to expose a domain-specific interface specially designed
-for the protocol. While the authors of the CORFU paper describe both a
-host-based and FPGA-based solution, RADOS directly supports the concept of
-custom storage interfaces to its objects that can be used to achieve the
-transactional semantics necessary to implement the CORFU storage interfaces.
-This feature is called RADOS object classes, and is one of the most powerful
-features offered by RADOS for designing storage applications.
+While the authors of the CORFU paper describe prototype device interfaces using
+both host-based and FPGA-based implementations, RADOS \emph{directly} supports
+the creation of logical storage devices through its object class feature
+described previously in Section~\ref{sec:objclass}. Thus, by using
+software-based object interfaces offered by RADOS flash devices in CORFU can be
+replaced by software-defined storage offering significant flexibility and a
+simplified design.
 
-\paragraph*{A CORFU Object Interface}
+The implementation of a custom object class that satisifies the needs of an
+application such as CORFU is often straightforward. However, as described in
+Section~\ref{sec:objclass} there are a variety of native object I/O interfaces
+available, and it is not always immediately clear how best to utilize these
+interfaces. A common approach is to first define the set of requirements for
+the desired interface, construct a design space parameterized on the set of
+native interfaces, and winnow this space down through a process of elimination.
+
+\paragraph*{Towards a CORFU Object Interface}
 The state-machine shown in Figure~\ref{fig:corfu-sm} shows the composition of
-actions for each component of the CORFU interface. For example, each
-operations begins with an \emph{epoch guard}, and a \emph{write} operation
-will consult an index and perform a bulk I/O operation.  The primary concern
-when mapping an interface onto Ceph is deciding how each native interface is
-used in the interface composition.
+actions for each component of the CORFU interface. For instance, all operations
+begin by applying an \emph{epoch guard} that ensures the request is tagged with
+an up-to-date epoch value. The \emph{read} (R) and \emph{write} (W) operations
+both proceed by (1) examining metadata associated with the target position, (2)
+performing I/O to read or write the log entry, and in the case of a write, (3)
+updates metadata for the target log position.
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 \end{figure}
+
+The primary concern of an application developer when implementing an object
+interface in Ceph is deciding how native interfaces are composed into a
+compound operation. These types of decisions are commonly referred to as
+physical design, and can affect performance and application flexibilty.
+For instance, one valid design option is to store each log position in
+an object with a name using a one-to-one mapping with the log entry position.
+However, as we will see in the next section this choice of a physical design
+results in poor performance in most circumstances.
 

### 2016-05-17 paper.tex
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 However, as we will see in the next section this choice of a physical design
-results in poor performance in most circumstances.
+can result in poor performance compared to other designs.
 
 \subsection{Physical Design}
+
+We categorize the challenges of physical design into log entry management, or
+metadata management.
+
+Our strategy is simply to construct the design space, and winnow it down using
+a set of targetted experiments.
+
+We'll start by forming a baseline performance profile without any protocol
+overhead.
+
+All designs store a log entry. We can benchmark the system without any
+overhead associated with indexing or metadata management.
+
+\paragraph*{Baseline Performance}
+
+Figure~\ref{fig:pd-map} illustrates the two primary dimensions of the physical
+design space. The first dimension concerns the strategy by which a log entry
+is addressed within RADOS. In a one-to-one strategy each log entry is stored
+in a distinct object with a name based on the log entry position.  This is an
+attractive option because it is trivial for a client to address any log entry,
+and an object interface does not require any complexity associated with
+multiplexing multiple entries---addressing relies on native indexing provided
+by the OSDs themselves. In contrast to a 1:1 strategy, an N:1 strategy
+\emph{stripes} log entries across a smaller set of objects, introducing added
+complexity into both the client and object interface implementations. The
+second design dimension selects the primary storage interface for log
+entries, namely the key-value or bytestream interfaces, described in
+Section~\ref{sec:objclass}.
+
+These two dimensions are represented by the first two columns of
+Table~\ref{tab:pd-map} which describes the entire design space, where
+\emph{KV} corresponds to the key-value interface, and the bytestream interface
+is represented by \emph{AP} and \emph{EX} for an append or explicit i/o
+straetgy.
+
+. Note that we
+decompose the bytestream I/O interface into two sub-strategies: I/O using
+explicit offsets (EX), and appends (AP).
+
+CORFU device interface requirements in Ceph as an object interface. So far we
+have
+
+append is more attractive, all things being equal, as it supports flexible
+mapping size.
 
@@
 \multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
-                     & BS  & Flex     & Ceph/VFS  & KV/BS \\ \hline
+                     & AP  & Flex     & Ceph/VFS  & KV/BS \\ \hline
 \multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
-                     & WR  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
+                     & EX  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
                      & AP  & Flex     & KV/BS     & KV/BS \\

### 2016-05-17 paper.tex
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-16 paper.tex
@@
 However, as we will see in the next section this choice of a physical design
-results in poor performance in most circumstances.
+can result in poor performance compared to other designs.
 
@@
 
-There is a finite set of strategies for mapping the CORFU interfaces onto the
-current set of native I/O interfaces used to construct object interfaces.
-Table~\ref{tab:pd-map} shows the design space for mapping the CORFU interfaces
-onto Ceph. In order to select the best mapping we performed a parameter sweet
-over the design space.
+As we have seen, Ceph provides a rich storage API and places few restrictions
+on the structure of applications. So how should we go about implementing CORFU
+on Ceph? Our strategy is straightforward: first we will define the entire
+design space, and we will then winnow down this space using a set of targetted
+benchmarks. The design space can be divided into three challenges:
+selecting a strategy for log entry addressing, choosing a native I/O interface
+for storing log entry content, and implementing efficient metadata management.
 
-\begin{figure*}[!ht]
+\begin{enumerate}
+    \item {\bf Entry addressing.} We refer to the method by which a client
+        locates a log entry in Ceph as entry addressing, and we consider two
+        strategies. In a one-to-one (1:1) strategy each log entry is stored in
+        a distinct object with a name derived from the associated log entry
+        position. This is an attractive option because it is trivial for
+        clients to locate a log entry given its position.  In contrast, an N:1
+        strategy \emph{stripes} log entires across a smaller set of objects,
+        but adds complexity to both the client and the object interface which
+        must multiplex a set of entries.
+
+    \item {\bf Log entry storage.} Clients read and write binary data
+        associated with each log entry, and these entries can be stored in the
+        bytestream or in the key-value database associated with an object.
+        Retrieval of log entry payloads should perform well for both large
+        (e.g. database checkpoint) and small log entries.
+
+    \item {\bf Metadata management.} The CORFU protocol defines the storage
+        interface semantics, such as enforcing up-to-date epoch values and a
+        write-once address space. The object interface constructed in Ceph must
+        implement these semantics in software by storing metadata (e.g. the
+        current epoch) and validating requests against this metadata (e.g. has
+        the target position been written?). A key-value store is a natural
+        location for this type of data, but metadata management adds overhead
+        to each request and must be carefully designed.
+\end{enumerate}
+
+In the remainder of this section we will explore the full design space defined
+by the cross product of these design challenges to arrive at a final design.
+In order to optimize this process we will iteratively winnow down the space
+through a series of targetted experiments.
+
+\paragraph*{Baseline Entry Storage Performance}
+We begin the process of exploring the design space by exploiting the fact that
+metadata management is a complexity and performance overhead that a full
+implementation must incur beyond the costs of storing log entry data.
+Therefore we first explore the costs of the design space by restricting the
+space to log entry addressing and log entry storage.  These two dimensions are
+represented by the first two columns of Table~\ref{tab:pd-map} which describes
+the entire design space. In the I/O column \emph{KV} corresponds to the
+key-value interface, and the bytestream interface is represented by \emph{AP}
+for an append strategy, and \emph{EX} for a strategy that writes to an
+explicit bytestream offset.
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=\columnwidth]{experiments/librados-sweep/jewel_ssd.png}
+%\caption{1hr, 1k, write}
+%\label{fig:vanilla_wr_jewel}
+%\end{figure}
+
+Figure~\ref{fig:vanilla_wr_jewel} shows the expected performance of 1K log
+appends without metadata management overhead using each of the five strategies
+defined in Table~\ref{tab:pd-map}. The first thing to notice in this figure is
+that both one-to-one strategies have relatively poor performance. The large
+period of reduced throughput for \emph{11wr} corresponds to the OSD splitting
+file system directories in order to maintain a maximum directory size, and
+will occur in \emph{11kv} although the threshold number of objects is not
+reached due to reduced overall throughput of the \emph{11kv} strategy.
+
+The second lesson that we can learn from Figure~\ref{fig:vanilla_wr_jewel} is
+that even when using an N:1 addressing strategy, the key-value interface
+imposes a large overhead. This is unfortuante because the key-value interface
+can provide a direct solution to addressing log entries within an object.
+Instead, what we find is that an N:1 addressing strategy that stores log
+entries in the bytestream using either object appends or writes to explicit
+offsets outperform all other strategies by a factor of over 2x.
+
+The apparent performance tie between the strategies of appending to an object
+and writing to explicit offsets can be broken by considering the flexibility
+offered by each approach. The third column \emph{entry size} shown in
+Table~\ref{tab:pd-map} lists if a particular strategy supports storage of
+entries with dynamic sizes, or if entries must be restricted to a fixed size.
+Notably the strategy that stores log entries at explicit object offsets is
+limited in this regard because each log entry is effectively pre-mapped into
+the storage system. This leaves the clear winner: an N:1 strategy that appends
+log entries to objects provides flexibilty and the best write performance in
+this particular configuration.
+
+\begin{figure}[t]
   \centering
-    \begin{subfigure}{.6\columnwidth}
-        \includegraphics[width=\columnwidth]{experiments/librados-sweep/jewel_ssd.png}
-        \caption{a}
-        \label{fig:vanilla-io-diff}
-    \end{subfigure}\hfill
-    \begin{subfigure}{.6\columnwidth}
-        \includegraphics[width=\columnwidth]{experiments/librados-sweep/firefly_ssd.png}
-        \caption{b}
-    \end{subfigure}\hfill
-    \begin{subfigure}{.6\columnwidth}
-        \includegraphics[width=\columnwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-        \caption{c}
-    \end{subfigure}\hfill
-    \caption{Shown here are the graphs and such that demonstrate that the same
-        physical design choices are not the same between differing version of
-    Ceph even on the same hardware.}
-\end{figure*}
+  \begin{subfigure}[b]{.47\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
+      \caption{1hr,1k,wr}
+      \label{fig:osdx}
+  \end{subfigure}\quad
+  \begin{subfigure}[b]{.47\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+      \caption{1hr,1k,rd}
+      \label{fig:cephx}
+  \end{subfigure}
+  \caption{Ceph and OSD... gotta cram more stuff in here...}
+\end{figure}
 
-Figure~\ref{fig:vanilla-io-diff} shows the result of a parameter sweep which
-clearly shows that an N-1 mapping based on the bytestream interface provides
-superiour performance. However, the other two graphs show the same interfaces
-running on an old version of Ceph that show the same decision would not have
-been the optimal choice {\bf \emph{note: for the outline these are not the
-real graphs we'll be using}}.
+%\begin{figure}
+%\centering
+%\includegraphics[width=\columnwidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+%\caption{1hr, 1k, read}
+%\label{fig:vanilla_rd_jewel}
+%\end{figure}
+
+This result is corroborated by considering the read performance for each
+strategy as well. Figure~\ref{fig:vanilla_rd_jewel} shows the expected random
+read performance from a log containing 1K entries in which an addressing
+strategy that stores log entries in the bytestream has the best performance.
 

### 2016-05-18 paper.tex
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 % why is it interesting
-The widely deployed Ceph distributed storage system is an example of a storage
-system that supports application-specific extensions in the form of custom I/O
-interfaces to objects managed by the underlying RADOS object storage
-system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly reliant
-upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by a marked
-increase in the number of object operations that are packaged as part of the
-Ceph distribution and widely used by internal Ceph subystems and by
-applications such as OpenStack Swift and Cinder~\cite{openstack}. In addition
-to the growth in the quantity of operations, Figure~\ref{fig:objclass-dev}
-also depicts the amount of low-level C++ written to implement these
-operations. Unfortunately, this code is written assuming a performance profile
-defined by the combination of the hardware and software versions available at
-the time of development. As Ceph continues to evolve at a fast pace, the cost
-of adapting application-specific codes to changing assumptions regarding
-performance may become significant.
-
-%The construction of object operations is not limited to core Ceph developers;
-%the development community has been receptive to contributions with the recent
-%inclusion by CERN developers of an extension for performing limited numeric
-%operations on object data~\cite{cls_numops}. And while Ceph has not yet
-%reached the point of directly exposing these features to non-administrative
-%users, the recent inclusion of a mechanism for dynamically defining extensions
-%using Lua~\cite{cls_lua} suggests that aspects of this feature may soon appear.
-
-% a challenge is mapping application semantics onto these objects
-
-%Proprietary storage systems are prevalent, but have fixed interfaces that are
-%developed in response to market demand. Alternative storage interfaces are
-%prototyped in research systems.
-%For the first time Ceph is providing a high-performance production quality
-%storage system developed as open-source.
-
-% why is it hard / why do naive approaches fail
-%the naive approach here is the use of generic interfaces. one may
-%think that key-value is relatively slower than bytestreams, but by how much.
-%is this a principal we can bet the house on.
-%just building these simple interfaces don't work that well
-%Systems provide low-level native interfaces such as key-value or bytestream interfaces.
-%These abstract interfaces are useful, but they are not the only way to map
-%interfaces onto storage. For instance, a specialized key-value db could be serialized into
-%the bytestream interface.
+The standardization of the POSIX file I/O interface has been a major success,
+allowing application developers to avoid vendor lock-in. However, large-scale
+storage systems have been dominated by proprietary products, preventing
+exploration of alternative interfaces and complicating future migration paths,
+eliminating the benefits of commodity systems. But the recent availability of
+high-performance open-source storage systems is changing this because these
+systems are modifiable, enabling interface change, and reducing the risks of
+lock-in. The widely deployed Ceph distributed storage system is an example of
+a storage system that supports application-specific extensions in the form of
+custom I/O interfaces to objects managed by the underlying RADOS object
+storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
+reliant upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by
+a marked increase in the number of object operations that are packaged as part
+of the Ceph distribution and widely used by internal Ceph subsystems and by
+applications such as OpenStack Swift and Cinder~\cite{openstack}.
 
@@
 
+%%% why is it hard / why do naive approaches fail
+In addition to the growth in the quantity of operations in use throughout Ceph
+installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
+low-level C++ written to implement these operations. Unfortunately, this code
+is written assuming a performance profile defined by the combination of the
+hardware and software versions available at the time of development.  While
+the bulk of these interfaces are created by core Ceph developers with a
+complete view of the performance model, this may be changing as the
+development community has been receptive to outside contributions with the
+recent inclusion by CERN developers of an extension for performing limited
+numeric operations on object data~\cite{cls_numops}. And while Ceph has not
+yet reached the point of directly exposing these features to
+non-administrative users, the recent inclusion of a mechanism for dynamically
+defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
+feature may soon appear. What is needed is support for creating storage
+interfaces using a method that allows transparent optimization as the system,
+application, and supporting environment evolve.
+
 % why hasn't it been solved before / why have previous approaches failed
-%As hardware becomes faster the differences in storage interface performance
-%will continue shift relative to one another.
-%Existing systems fall into the same trap.
+Previous work related to storage interface design has largely been in the
+context of standardization efforts and active storage. While the later has
+been specifically concerned with the creation of application-specific
+interfaces and demonstrating short-term benefits, the standardization seeks to
+develop a last set of interfaces. In neither case have we seen efforts focused
+on extensible interfaces with portability and longevity a primary concern.
+
 %This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
@@
 %power of including application-specific semantics.
-%Previous work in active storage has focused on the utility of moving
-%computation closer to data in an effort to reduce data movement and take
-%advantage of cpu and i/o parallelism, and gave little to no attention to
-%security or performance isolation concerns. In contrast, the interfaces being
-%deployed today focus on data management, indexing, and physical format.
-%Database systems solve this problem with a declarative query language, compiler, and
-%optimizer. So should we.
-%
-%We propose the introduction of a declarative language for building object
-%based interfaces that allows a storage system to meet the needs of an
-%application throughout the development process without requiring rewrites.
+
+We propose the introduction of a declarative language for building object
+based interfaces that allows a storage system to meet the needs of an
+application throughout the development process without requiring rewrites.
 
@@
 CPU, GBs of RAM, and one or two 10 Gb Ethernet links. Clients access object
-data managed by an OSD by invoking object operations exposed by the OSD
+data managed by an OSD by invoking native object operations exposed by the OSD
 such as reading or writing bytes, as well as more complex operations like taking
@@
 The key-value interface operates as a dedicated database associated with each
-object, and the bytestream interface supports random byte-level access.  At a
+object, and the bytestream interface supports random byte-level access similar
+to a file.  At a
 low-level each of these abstract I/O interfaces map to hardware storage
@@
 POSIX file system~\cite{leveldb,rocksdb}. Several backend implementations
-exist for storing data in targets such as a Kinetic Drive or in NVMe devices,
+exist for storing data in targets such as the Kinetic Drive or in NVMe devices,
 among others.
@@
 and applications, but common implementations such as Paxos or Raft funnel I/O
-through a single node limiting the throughput of log
-operations~\cite{lamport:tocs89}. The CORFU protocol addresses this limitation
+through a single node limiting total throughput~\cite{lamport:tocs89}. The CORFU protocol addresses this limitation
 by de-coupling log entry storage from log metadata management, making use of a
-centeralized, volatile, in-memory \emph{sequencer service} that assigns log
+centralized, volatile, in-memory \emph{sequencer service} that assigns
 positions to clients that are appending to the log. Since the sequencer is
-centeralized serialization is trivial, and the use of non-durable state allows
+centralized serialization is trivial, and the use of non-durable state allows
 the sequencer service to operate at very high rates. The CORFU system has been
@@
 transparently take advantage of any software or hardware upgrades, and make use
-of existing and future data management features such as tiering in RADOS,
+of existing and future data management features such as tiering,
 allowing users to freely choose between media types such as SSD, spinning
-disks, or emerging NVRAM technologies.
+disks for archival storage, or emerging NVRAM technologies.
 
@@
 clients are tagged with an \emph{epoch} value, and flash devices are expected
-to reject client requests that contian an old epoch value. To facilitate
+to reject client requests that contain an old epoch value. To facilitate
 recovery or handle system reconfiguration in CORFU, the storage devices are
-also required to support a \emph{seal(epoch)} command that marks the latest
+also required to support a \emph{seal(epoch)} command that stores the latest
 epoch and returns the maximum position written to that device. The seal
@@
 
-While the authors of the CORFU paper describe prototype device interfaces using
-both host-based and FPGA-based implementations, RADOS \emph{directly} supports
+While the authors of the CORFU paper describe prototype device interfaces
+implemented as both host-based and FPGA-based solutions, RADOS \emph{directly} supports
 the creation of logical storage devices through its object class feature
@@
 
-The implementation of a custom object class that satisifies the needs of an
+The implementation of a custom object class that satisfies the needs of an
 application such as CORFU is often straightforward. However, as described in
@@
 available, and it is not always immediately clear how best to utilize these
-interfaces. A common approach is to first define the set of requirements for
-the desired interface, construct a design space parameterized on the set of
-native interfaces, and winnow this space down through a process of elimination.
+interfaces.
 
 \paragraph*{Towards a CORFU Object Interface}
-The state-machine shown in Figure~\ref{fig:corfu-sm} shows the composition of
+The state-machine diagram in Figure~\ref{fig:corfu-sm} shows the composition of
 actions for each component of the CORFU interface. For instance, all operations
@@
 \caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
-trim ($T$) CORFU operaitons. The states epoch guard ($EG$), check position ($CP$),
+trim ($T$) CORFU operations. The states epoch guard ($EG$), check position ($CP$),
 and update state ($US$) access metadata. The I/O performs a log entry read or
@@
 compound operation. These types of decisions are commonly referred to as
-physical design, and can affect performance and application flexibilty.
-For instance, one valid design option is to store each log position in
-an object with a name using a one-to-one mapping with the log entry position.
-However, as we will see in the next section this choice of a physical design
-can result in poor performance compared to other designs.
+physical design, and can affect performance and application flexibility.  For
+instance, one valid design option is to store each log position in an object
+with a name using a one-to-one mapping with the log entry position.  This
+would simplify the design of the \emph{write} interface because a small amount
+of metadata stored as a header in the object could describe the state of the
+log entry.  However, as we will see in the next section this choice of a
+physical design can result in poor performance compared to other designs.
 
@@
 on Ceph? Our strategy is straightforward: first we will define the entire
-design space, and we will then winnow down this space using a set of targetted
+design space, and we will then winnow down this space using a set of targeted
 benchmarks. The design space can be divided into three challenges:
@@
         strategy \emph{stripes} log entires across a smaller set of objects,
-        but adds complexity to both the client and the object interface which
+        but this adds complexity to both the client and the object interface which
         must multiplex a set of entries.
@@
 by the cross product of these design challenges to arrive at a final design.
-In order to optimize this process we will iteratively winnow down the space
-through a series of targetted experiments.
 
@@
 implementation must incur beyond the costs of storing log entry data.
-Therefore we first explore the costs of the design space by restricting the
+Therefore we first explore the design space by restricting the
 space to log entry addressing and log entry storage.  These two dimensions are
@@
 for an append strategy, and \emph{EX} for a strategy that writes to an
-explicit bytestream offset.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\columnwidth]{experiments/librados-sweep/jewel_ssd.png}
-%\caption{1hr, 1k, write}
-%\label{fig:vanilla_wr_jewel}
-%\end{figure}
+explicit bytestream offset (both described shortly).
 
@@
 will occur in \emph{11kv} although the threshold number of objects is not
-reached due to reduced overall throughput of the \emph{11kv} strategy.
+reached in this example due to reduced overall throughput of the \emph{11kv} strategy.
 
@@
 that even when using an N:1 addressing strategy, the key-value interface
-imposes a large overhead. This is unfortuante because the key-value interface
+imposes a large overhead. This is unfortunate because the key-value interface
 can provide a direct solution to addressing log entries within an object.
@@
 and writing to explicit offsets can be broken by considering the flexibility
-offered by each approach. The third column \emph{entry size} shown in
-Table~\ref{tab:pd-map} lists if a particular strategy supports storage of
+offered by each approach. The third column \emph{entry size} in
+Table~\ref{tab:pd-map} shows if a particular strategy supports storage of
 entries with dynamic sizes, or if entries must be restricted to a fixed size.
@@
 the storage system. This leaves the clear winner: an N:1 strategy that appends
-log entries to objects provides flexibilty and the best write performance in
-this particular configuration.
+log entries to objects provides flexibility and the best write performance in
+this particular configuration.  This result is corroborated by considering the
+read performance for each strategy as well. Figure~\ref{fig:vanilla_rd_jewel}
+shows the expected random read performance from a log containing 1K entries in
+which an addressing strategy that stores log entries in the bytestream has the
+best performance.
 
@@
       \caption{1hr,1k,wr}
-      \label{fig:osdx}
+      \label{fig:vanilla_wr_jewel}
   \end{subfigure}\quad
@@
       \caption{1hr,1k,rd}
-      \label{fig:cephx}
+      \label{fig:vanilla_rd_jewel}
   \end{subfigure}
@@
 \end{figure}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\columnwidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-%\caption{1hr, 1k, read}
-%\label{fig:vanilla_rd_jewel}
-%\end{figure}
-
-This result is corroborated by considering the read performance for each
-strategy as well. Figure~\ref{fig:vanilla_rd_jewel} shows the expected random
-read performance from a log containing 1K entries in which an addressing
-strategy that stores log entries in the bytestream has the best performance.
 
@@
 \end{table}
+
+\paragraph*{Metadata Management}
+
+Adds in indexing and that sort of overheads.
+
+\paragraph*{Hardware and Software}
+
+Different hardware and different Ceph versions.
 

### 2016-05-18 paper.tex
paper: update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 on Ceph? Our strategy is straightforward: first we will define the entire
-design space, and we will then winnow down this space using a set of targeted
+design space, and then winnow down this space using a set of targeted
 benchmarks. The design space can be divided into three challenges:
@@
 Therefore we first explore the design space by restricting the
-space to log entry addressing and log entry storage.  These two dimensions are
+space to log entry addressing and log entry storage (the \emph{I/O} action
+shown in Figure~\ref{fig:corfu-sm}).  These two dimensions are
 represented by the first two columns of Table~\ref{tab:pd-map} which describes
@@
 explicit bytestream offset (both described shortly).
+
+\begin{table}[h]
+\begin{tabular}{ | l | l | l | l | l |}
+\hline
+Map & I/O & Entry Size & Addressing & Metadata \\ \hline
+\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
+                     & AP  & Flex     & Ceph/VFS  & KV/BS \\ \hline
+\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
+                     & EX  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
+                     & AP  & Flex     & KV/BS     & KV/BS \\
+\hline
+\end{tabular}
+\caption{The high-level design space of mapping CORFU log entry storage onto
+the RADOS object storage system.}
+\label{tab:pd-map}
+\end{table}
 
@@
 will occur in \emph{11kv} although the threshold number of objects is not
-reached in this example due to reduced overall throughput of the \emph{11kv} strategy.
+reached in this example due to reduced overall throughput of the \emph{11kv}
+strategy.
 
@@
       \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
-      \caption{1hr,1k,wr}
+      \caption{1hr,1k,wr,jewel}
       \label{fig:vanilla_wr_jewel}
-  \end{subfigure}\quad
+  \end{subfigure}
   \begin{subfigure}[b]{.47\linewidth}
@@
       \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-      \caption{1hr,1k,rd}
+      \caption{1hr,1k,rd,jewel}
       \label{fig:vanilla_rd_jewel}
   \end{subfigure}
-  \caption{Ceph and OSD... gotta cram more stuff in here...}
+  \begin{subfigure}[b]{.47\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+      \caption{1hr,1k,cls\_wr,jewel}
+      \label{fig:cls_wr_jewel}
+  \end{subfigure}
+  \begin{subfigure}[b]{.47\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/firefly_ssd.png}
+      \caption{1hr,1k,wr,firefly}
+      \label{fig:vanilla_wr_firefly}
+  \end{subfigure}
+  \caption{placeholder.}
 \end{figure}
-
-\begin{table}
-\begin{tabular}{ | l | l | l | l | l |}
-\hline
-Map & I/O & Entry Size & Addressing & Metadata \\ \hline
-\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
-                     & AP  & Flex     & Ceph/VFS  & KV/BS \\ \hline
-\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
-                     & EX  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
-                     & AP  & Flex     & KV/BS     & KV/BS \\
-\hline
-\end{tabular}
-\caption{The high-level design space of mapping CORFU log entry storage onto
-the RADOS object storage system.}
-\label{tab:pd-map}
-\end{table}
 
@@
 
-Adds in indexing and that sort of overheads.
+The previous results show that storing log entries in the bytestream has the
+potential to provide the best overall performance, but by design, those
+results do not contain the real-world overheads introduced by metadata
+management such as validating that requests are tagged with an up-to-date
+epoch value. Having only focused on entry I/O costs, in this section we
+consider the overhead of the remaining actions shown in
+Figure~\ref{fig:corfu-sm}.
+
+\begin{enumerate}
+    \item {\bf Epoch guard.} Each client request must be validated against the
+        current epoch. This singleton value in infrequently updated, but the
+        cost of accessing the value must be incurred for every request.
+
+    \item {\bf Per-entry metadata.} Unlike the singleton epoch value, metadata
+        associated with each log position must be read, and optionally updated
+        for every request. For instance the \emph{fill} operation must ensure
+        that it is not applied to a log position that has already been
+        written. When an operation that changes a log position is successful
+        (i.e. \emph{write}, \emph{fill}, \emph{trim}) the per-entry metadata
+        must also be updated to reflect the change. While the size of the
+        per-entry metadata is small, it is accessed for every operation and a
+        single object may manage a large number of entries.
+\end{enumerate}
+
+The design space for managing metadata can be large depending on the
+application. In many cases this space is significantly reduced when the
+key-value interface is used because it handles a large portion of common data
+management challenges such as indexing. On the other hand, we have seen that
+the bytestream can be used to achieve much higher performance when used in
+lieu of the key-value database, but leaves the design space wide open.
+
+We consider two high-level design prototypes in this paper based on both the
+key-value and bytestream interfaces. The summary of performance of these
+designs is shown in Figure~\ref{fig:cls_wr_jewel} where we begin by
+highlighting two baseline throughputs labeled \emph{librados} and
+\emph{cls\_null} that correspond to the baseline append throughput described
+in the previous section, and the append throughput achieved with \emph{object
+class} overhead included. We will discuss the performance of our two prototype
+designs relative to the base cost of using \emph{object class} facility.
+
+The first design is based on the key-value interface. In this design the epoch
+value is stored under a fixed key, and the metadata associated with each log
+entry is stored under a key derived from the log position. The expected throughput of
+this design is shown in Figure~\ref{fig:cls_wr_jewel} and labeled as
+\emph{cls\_full\_omap}. This result shows us that even for very small values
+the cost of managing data in the key-value store introduces significant
+overhead. In contrast, the unstructured bytestream interface imposes little
+to no restriction on how it is used.
+
+An obvious choice for encoding the singleton epoch value in the bytestream is
+to position it at a fixed offset such as in a header. A method for indexing
+entry metadata is far less obvious. Solutions span a wide design spectrum and
+include approaches such as encoding a search tree into the bytestream, or
+exploiting the regularity log addressing and storing a dense array that can be
+efficiently indexed. While there is clearly a wide variety of approaches, we
+have chosen to examine the overhead incurred by an hypothetical best-case
+design in which per-entry metadata can be accessed with a single I/O. The I/O
+overhead of the test includes (1) read epoch value from a fixed header
+location, (2) read the per-entry metadata, and in the case of an update (3)
+write the updated per-entry metadata. The expected throughput from such a
+design is shown in Figure~\ref{fig:cls_wr_jewel} labeled as
+\emph{cls\_full\_hdr} and performs with roughly a 6\% overhead.
 

### 2016-05-19 paper.tex
paper: update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
paper: updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-Different hardware and different Ceph versions.
+In this section we have demonstrated that a design based on an N:1 addressing
+scheme that stores both log entries and metadata in the bytestream can provide
+the best overall performance, and it does so by a large margin. However, this
+process of design we have outlined may not yield such clear cut results when
+applied in other contexts that may differ by network and hardware capabilities
+as well as configuration options and software versions.
+
+To illustrate this we ran the same benchmark that measures expected append
+throughput shown in Figure~\ref{fig:vanilla_wr_jewel} on identical hardware
+but with a version of long-term support version of Ceph released two years
+ago. The result of this benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly}
+in which the bytestream interface outperforms the key-value interface by 12\%,
+compared to over 100\% in the newer versino of Ceph. Qualatatively, given the
+reduced overall throughput achieved in the older version, some developers may
+find that incurring the overhead of using the key-value interface is an easy
+decision given the reduced complexity of the design space for metadata
+management.
 

### 2016-05-19 paper.tex
paper: update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 throughput shown in Figure~\ref{fig:vanilla_wr_jewel} on identical hardware
-but with a version of long-term support version of Ceph released two years
+but we used a long-term support version of Ceph released two years
 ago. The result of this benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly}
@@
 decision given the reduced complexity of the design space for metadata
-management.
+management. If this choice had been made, then a future system upgrade could
+drop a significant number of iops on the floor.
 

### 2016-05-19 experiments/multi-node/run.sh
expr: add cluster microbenchmarks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -2,6 +2,14 @@
 
 set -ex
 
-ANSIBLE_LOG_PATH="logs/ceph-zeqr.log" \
+#for i in 1 2 3 4 5 6 7 8 9 10; do
+#for i in 2; do
+#  ANSIBLE_LOG_PATH="logs/radosbench-${i}.log" \
+#  ansible-playbook --forks 50 --skip-tags "with_pkg" \
+#  -i inventory/cluster-${i}node -e inventory_name=cluster-${i}node \
+#  ceph.yml workloads/radosbench.yml
+#done
+
+ANSIBLE_LOG_PATH="logs/radosbench-${i}.log" \
 ansible-playbook --forks 50 --skip-tags "with_pkg" \
 ceph.yml workloads/bench-ng.yml

### 2016-05-20 experiments/basic-cls-rand-read/run.sh
expr: add new expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,28 +1,4 @@
 #!/bin/bash
 
-reset=soft
-data_dev=sdc
-
-name=basic_cls_rand_read
-runtime=3600
-pg_nums="128"
-stripe_widths="128"
-queue_depths="32"
-entry_sizes="4096"
-pool=zlog
-rest=120
-read_runtime=3600
-reads=yes
-
-# workloads
-wl_11="map_11 bytestream_11"
-wl_n1="map_n1 bytestream_n1_write"
-workloads="$wl_11 $wl_n1"
-
-# i/o interfaces
-map_n1_if="vanilla"
-bytestream_n1_write_if="vanilla"
-bytestream_n1_append_if="vanilla"
-
-. $ZLOG_ROOT/docker/physical-design/runner.sh
-run_pd
+./run1.sh
+./run2.sh

### 2016-05-20 experiments/basic-cls-rand-read/run1.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+reset=soft
+data_dev=nvme0n1
+journal_dev=nvme1n1
+noop_dev="nvme0n1 --noop nvme1n1"
+
+name=librados_read_data_jrnl_nvme
+runtime=1800
+pg_nums="128"
+stripe_widths="128"
+queue_depths="32"
+entry_sizes="1024"
+pool=zlog
+rest=120
+read_runtime=1800
+reads=yes
+
+# workloads
+wl_11="map_11 bytestream_11"
+wl_n1="map_n1 bytestream_n1_write"
+workloads="$wl_11 $wl_n1"
+
+# i/o interfaces
+map_n1_if="vanilla"
+bytestream_n1_write_if="vanilla"
+bytestream_n1_append_if="vanilla"
+
+. $ZLOG_ROOT/docker/physical-design/runner.sh
+run_pd

### 2016-05-20 experiments/basic-cls-rand-read/run2.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+reset=soft
+data_dev=nvme0n1
+noop_dev="nvme0n1"
+
+name=librados_read_data_nvme
+runtime=1800
+pg_nums="128"
+stripe_widths="128"
+queue_depths="32"
+entry_sizes="1024"
+pool=zlog
+rest=120
+read_runtime=1800
+reads=yes
+
+# workloads
+wl_11="map_11 bytestream_11"
+wl_n1="map_n1 bytestream_n1_write"
+workloads="$wl_11 $wl_n1"
+
+# i/o interfaces
+map_n1_if="vanilla"
+bytestream_n1_write_if="vanilla"
+bytestream_n1_append_if="vanilla"
+
+. $ZLOG_ROOT/docker/physical-design/runner.sh
+run_pd

### 2016-05-20 paper.tex
paper: related work mic drop
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 hardware continue to evolve.
+
+%%% Anderson FAST 2016 IoT Time Series DB on Ceph
 
@@
 
-\subsection{Physical Design}
+\section{Physical Design}
 
@@
 
-\paragraph*{Baseline Entry Storage Performance}
+\subsection{Baseline Performance}
+
 We begin the process of exploring the design space by exploiting the fact that
@@
 
-\paragraph*{Metadata Management}
+\subsection{Metadata Management}
 
@@
 
-\paragraph*{Hardware and Software}
+\subsection{Hardware and Software}
 
@@
 
+\subsection{Optimization}
+
+explain that the state space is large, this is hard for humans.
+configuration, software versions, hardware versions, interface selections, and
+future features.
+
 \section{Programming Model}
 
-In this section we are going to be describing our way of creating interfaces.
+Describe big picture as if we are going to introduce something like bloom for
+describing everything.
+
+], Bloom is purely declarative: the syntax of a program contains the full
+spec- ification of its semantics, and there is no need for the programmer to
+understand or reason about the behavior of the evaluation engine. Bloom is
+based on a formal temporal logic called Dedalu
+
+We use a subset of bloom as a basis for defining interface.
+
+Although we have not chosen to use the most powerful parts of the bloom (eg
+task coordination and consistency)
+language we have chosen to start with a powerful abstraction because we plan
+to take this to 11 (see future work).
 
@@
 
-\section{Conclusion}
+{\bf Active storage.}
+There is a wide variety of work related \emph{active storage} that seeks to
+increase performance and reduce data movement by exploiting remote storage
+resources. This concept has been applied in the context of on-disk
+controllers and object-based storage
+(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
+concerns have examined using managed lanuages, sandbox technologies, as well
+as restricting extension installation to
+vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.
+
+Techniques for remote storage processing have been applied in
+application-specific domains for database and file system acceleration, as
+well as in remote data filtering in cloud-based storage
+environments~\cite{uysal:hpca00,chiu:iccs03,lim:msst08,gkantsidis:nsdi13}
+Others have collected and used statistics to optimize the location of
+computation in general workloads as well as database systems
+~\cite{chen:cluster12,chen:icpp12,qiao:icde08}.
+
+Most closely related to our work are efforts that consider specific
+programming models in the context of active storage such as a stream-based
+model~\cite{acharya:asplos98}, a model that assisted in optmiizing balancing
+computation~\cite{wickremesinghe:hpdc022}, and optimization that took
+advantage of read-only data to reorder operations~\cite{huston:fast04}.
+While our goals are similar to previous work, we start with a declarative
+language that will enable us to apply optimization techniques beyond
+the limited domain-specific optimizations found in previous work.
+
+\section{Conclusion and Future Work}
+
+Include performance management topic in related work.
 

### 2016-05-21 experiments/basic-cls-rand-read/run.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -2,3 +2,4 @@
 
 ./run1.sh
 ./run2.sh
+./run3.sh

### 2016-05-21 experiments/basic-cls-rand-read/run1.sh

@@ -1,16 +1,15 @@
 #!/bin/bash
 
 reset=soft
-data_dev=nvme0n1
-journal_dev=nvme1n1
-noop_dev="nvme0n1 --noop nvme1n1"
+data_dev=sdc
+noop_dev=sdc
 
-name=librados_read_data_jrnl_nvme
+name=librados_jewel_ssd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
-entry_sizes="1024"
+entry_sizes="128"
 pool=zlog
 rest=120
 read_runtime=1800

### 2016-05-21 experiments/basic-cls-rand-read/run3.sh

@@ -4,12 +4,12 @@ reset=soft
 data_dev=sdc
 noop_dev=sdc
 
-name=librados_jewel_ssd_med
+name=librados_jewel_ssd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
-entry_sizes="1024"
+entry_sizes="4096"
 pool=zlog
 rest=120
 read_runtime=1800

### 2016-05-21 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,10 +1,9 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 
-name=librados_jewel_ssd_small
+name=librados_jewel_hdd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-21 experiments/basic-cls-rand-read/run2.sh

@@ -1,10 +1,9 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 
-name=librados_jewel_ssd_med
+name=librados_jewel_hdd_med
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-21 experiments/basic-cls-rand-read/run3.sh

@@ -1,10 +1,9 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 
-name=librados_jewel_ssd_large
+name=librados_jewel_hdd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-21 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -2,8 +2,10 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 
-name=librados_jewel_hdd_small
+name=librados_jewel_ssd_hdd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-21 experiments/basic-cls-rand-read/run2.sh

@@ -2,8 +2,10 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 
-name=librados_jewel_hdd_med
+name=librados_jewel_ssd_hdd_med
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-21 experiments/basic-cls-rand-read/run3.sh

@@ -2,8 +2,10 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 
-name=librados_jewel_hdd_large
+name=librados_jewel_ssd_hdd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run1.sh
expr: add experiment
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -2,10 +2,9 @@
 
 reset=soft
 data_dev=sdb
-journal_dev=sdc
-noop_dev=sdc
+ceph_version=firefly
 
-name=librados_jewel_ssd_hdd_small
+name=librados_firefly_hdd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run2.sh

@@ -2,10 +2,9 @@
 
 reset=soft
 data_dev=sdb
-journal_dev=sdc
-noop_dev=sdc
+ceph_version=firefly
 
-name=librados_jewel_ssd_hdd_med
+name=librados_firefly_hdd_med
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run3.sh

@@ -2,10 +2,9 @@
 
 reset=soft
 data_dev=sdb
-journal_dev=sdc
-noop_dev=sdc
+ceph_version=firefly
 
-name=librados_jewel_ssd_hdd_large
+name=librados_firefly_hdd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -2,9 +2,11 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_hdd_small
+name=librados_firefly_ssd_hdd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run2.sh

@@ -2,9 +2,11 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_hdd_med
+name=librados_firefly_ssd_hdd_med
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run3.sh

@@ -2,9 +2,11 @@
 
 reset=soft
 data_dev=sdb
+journal_dev=sdc
+noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_hdd_large
+name=librados_firefly_ssd_hdd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdb
-journal_dev=sdc
+data_dev=sdc
 noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_ssd_hdd_small
+name=librados_firefly_ssd_small
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run2.sh

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdb
-journal_dev=sdc
+data_dev=sdc
 noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_ssd_hdd_med
+name=librados_firefly_ssd_med
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-22 experiments/basic-cls-rand-read/run3.sh

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdb
-journal_dev=sdc
+data_dev=sdc
 noop_dev=sdc
 ceph_version=firefly
 
-name=librados_firefly_ssd_hdd_large
+name=librados_firefly_ssd_large
 runtime=1800
 pg_nums="128"
 stripe_widths="128"

### 2016-05-23 experiments/basic-cls-rand-read-bluestore/run1.sh
expr: add initial bluestore runs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+reset=soft
+data_dev=nvme0n1
+journal_dev=nvme1n1
+noop_dev="nvme0n1 --noop nvme1n1"
+osd_create="--bluestore"
+
+name=librados_read_data_jrnl_nvme_bluestore
+runtime=1800
+#runtime=500
+pg_nums="128"
+stripe_widths="128"
+queue_depths="32"
+entry_sizes="1024"
+pool=zlog
+rest=120
+#read_runtime=0
+read_runtime=1800
+reads=yes
+
+# workloads
+wl_11="map_11 bytestream_11"
+wl_n1="map_n1 bytestream_n1_write"
+#workloads="$wl_11 $wl_n1"
+workloads="$wl_n1"
+
+# i/o interfaces
+map_n1_if="vanilla"
+bytestream_n1_write_if="vanilla"
+bytestream_n1_append_if="vanilla"
+
+. $ZLOG_ROOT/docker/physical-design/runner.sh
+run_pd

### 2016-05-23 experiments/basic-cls-rand-read-bluestore/run2.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+reset=soft
+data_dev=nvme0n1
+noop_dev="nvme0n1"
+osd_create="--bluestore"
+
+name=librados_read_data_nvme_bluestore
+#runtime=1800
+runtime=500
+pg_nums="128"
+stripe_widths="128"
+queue_depths="32"
+entry_sizes="1024"
+pool=zlog
+rest=120
+#read_runtime=1800
+read_runtime=500
+reads=yes
+
+# workloads
+wl_11="map_11 bytestream_11"
+wl_n1="map_n1 bytestream_n1_write"
+#workloads="$wl_11 $wl_n1"
+workloads="$wl_n1"
+
+# i/o interfaces
+map_n1_if="vanilla"
+bytestream_n1_write_if="vanilla"
+bytestream_n1_append_if="vanilla"
+
+. $ZLOG_ROOT/docker/physical-design/runner.sh
+run_pd

### 2016-05-23 paper.tex
update paper
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \usepackage{tikz}
+\usepackage{listings}
+\usepackage{color}
 \usetikzlibrary{arrows,automata}
-\usepackage[latin1]{inputenc}
+%\usepackage[latin1]{inputenc}
+
+\lstset{frame=tb,
+  language=Ruby,
+  captionpos=b,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\scriptsize\ttfamily},
+  numbers=none,
+  %numberstyle=\tiny\color{gray},
+  %keywordstyle=\color{blue},
+  %commentstyle=\color{dkgreen},
+  %stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3
+}
 
@@
 
-\subsection{Hardware and Software}
+\subsection{Optimization}
 
@@
 process of design we have outlined may not yield such clear cut results when
-applied in other contexts that may differ by network and hardware capabilities
-as well as configuration options and software versions.
+applied in a different context.
 
@@
 
-\subsection{Optimization}
+In practice the real world state space is much larger than the one described
+in this section. Ceph alone has hundreds of configuration options, significant
+features becoming available with releases each year, and is designed to run on
+a wide variety of COTS hardware as well as new NVMe devices, all of which
+contribute to a large and dynamic state space. Application-level designs
+also contribute to an expanded design space. In the N:1 design of mapping
+CORFU onto Ceph the potential amount of I/O parallelism is controlled by 
+the value N, however if this value becomes too large there may be no benefit
+or performance could suffer from reduced data locality. Other aspects of
+application-level design are even more flexible. A wide variety of data
+structures could be serialized into the byte stream to handle metadata
+management tasks, but selecting a single external index that performs well
+across workloads and hardware may not be possible.
 
-explain that the state space is large, this is hard for humans.
-configuration, software versions, hardware versions, interface selections, and
-future features.
+Next we present a declarative programming model that seeks to solve the issue
+of selecting an optimal mapping between application requirements and object
+class implementation.
 
@@
 
-Describe big picture as if we are going to introduce something like bloom for
-describing everything.
+Ceph is turning into a target for building distrubted services, bringing
+distributed programming into the hands of many new people that may have
+worked on the periphery of ceph.
 
-], Bloom is purely declarative: the syntax of a program contains the full
-spec- ification of its semantics, and there is no need for the programmer to
-understand or reason about the behavior of the evaluation engine. Bloom is
-based on a formal temporal logic called Dedalu
+Getting consistency and coordination correct is a challenge And as we have seen
+optimization at a low-level given a simple set of semantics can yield a large
+state space from which to draw designs Difficulty and challenges of writing
+distributed software.  Programming distributed systems and applications.
 
-We use a subset of bloom as a basis for defining interface.
+We draw upon the recent work of Bloom a declarative lang thing
 
-Although we have not chosen to use the most powerful parts of the bloom (eg
-task coordination and consistency)
-language we have chosen to start with a powerful abstraction because we plan
-to take this to 11 (see future work).
+BLL (a.k.a. BRADOS) is a declarative language with relational (or tabular)
+semantics for RADOS objects, which corresponds to a subset of the BLOOM
+language (leaving out temporal logic). BLL borrows from BLOOM the table and
+interface collections, the instantaneous implication (‘<=’) operator, as well
+as all the collection operators.
 
-\section{Other Interfaces}
+BRADOS models the storage system state uniformly as a collection of relations.
+The composition of a collection of existing interfaces is then expressed as a
+collection of high-level \emph{queries} that describe how a stream of requests
+(API calls) are filtered, transformed and combined with existing state to
+define streams of outputs (API call returns as well as updates to system
+state).  Separating the relational semantics of such compositions from details
+of their physical implementations introduces a number of degrees of freedom,
+similar to the ``data independence’’ offered by database query languages.  The
+choice of access methods (for example, whether to use a bytestream interface or
+a key/value store), storage decide classes(e.g., whether to use HDDs or SDDs),
+physical layout details (e.g. a 1:1 or 1:N mapping) and execution plans (e.g.
+operator ordering) can be postponed to execution time.  The optimal choices for
+these physical design details are likely to change much more often than the
+logical specification, freeing the interface designer from the need to rewrite
+systems and device and interface characteristics change.
+
+\subsection{The CORFU Interface in Bloom}
+
+Listing~\ref{lab} shows code.
+
+\begin{lstlisting}[caption={Hello}, label=lab]
+class Corfu
+include Bud
+
+state do
+table :epoch, [:epoch]
+table :log, [:pos] => [:state, :data]
+
+scratch :valid_op, [:type, :pos, :epoch] => [:data]
+scratch :invalid_op, [:type, :pos, :epoch] => [:data]
+
+interface input, :op, [:type, :pos, :epoch] => [:data]
+interface output, :ret, [:type, :pos, :epoch] => [:retval]
+end
+\end{lstlisting}
+
+\subsection{Optimzer}
+
+\subsection{Other Interfaces}
 

### 2016-05-23 paper.tex
paper: update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\subsection{Optimzer}
+\subsection{Optimizer}
 
@@
 
-Here we show the derivation of two other interfaces that are in production in
-Ceph today to demonstrate the generality of our interface.
+The subset of the Bloom language we are using is powerful enough to express the
+CORFU language, as well as many of the object class interfaces that are already
+included in Ceph. We briefly sketch the Bloom imlementation of an object class
+that provides a locking service. This is an interesting example because Bloom
+implementations use mechanisms for introducing points of coordination in
+execution, which in the context of Ceph, may be written in Bloom itself.
+
+\begin{lstlisting}[caption={Hello}, label=lab]
+bloom :take_lock do
+end
+
+bloom :break_lock do
+end
+
+bloom :list_locks do
+end
+
+bloom :lock_info do
+end
+\end{lstlisting}
 
@@
 
-Include performance management topic in related work.
-
-%In this section we examine the design space for implementing the CORFU
-%protocol in the RADOS storage system. Table \ref{t:init-ds} shows the entire
-%design space. A description of each design parameter follows:
-%
-%{\bf Mapping strategy.} The method by which a log entry---identified by its
-%logical position---is addressed within RADOS is referred to as the mapping
-%strategy. A 1:1 strategy stores each log entry in a RADOS object with a
-%distinct name (e.g. ``mylog.pos443''), and an N:1 strategy stripes the log
-%positions across a set of objects (e.g.  round-robin). While we do consider
-%both strategies in this paper, a 1:1 strategy is attractive because it allows
-%a design in which clients can directly address log positions by constructing
-%the correct object name.
-%
-%{\bf Storage interface.} entry data may be small or large. its primary
-%storage location is important. kv, bs. we further sub-divide bs into
-%write and append. each strategy is compatible with fixed size entries
-%except for the write interface.
-%
-%{\bf Logical addressing.}
-
-%\subsection{Storage Interface Selection}
-%
-%Figure \ref{f:librados-sweep} shows the single-node write-only I/O throughput
-%for each of the points in the design space defined in Table \ref{t:init-ds}.
-%The results reveal poor relative performance using both the key-value storage
-%interface, as well as either of the 1:1 strategies---which incur overhead by
-%creating a new object for each log entry. Both of the N:1 strategies using the
-%bytestream I/O interface outperform all of the other approaches by over 2x
-%throughput, but otherwise have nearly identical performance.
-%
-%To examine the read performance of the points in the design space we generated
-%a dataset using each of the write workloads, and then issued random reads
-%across the entire dataset. Figure \ref{f:librados-rand-read} shows the read
-%throughput for each of the design strategies. The results show that an N:1
-%mapping using bytestream interface performs best. Interestingly, the 1:1
-%mapping strategy using the bytestream interface exhibits good read performance
-%in comparison to writes using the same strategy.
-
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=0.45\textwidth]{experiments/librados-sweep/output.soft.reset.2hr.png}
-%  \caption{
-%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/librados-sweep/visualize.ipynb}{source}]
-%Throughput (IOPS) of 1K writes to a single OSD using the standard I/O
-%interfaces in various configurations. The best performance is achieved using
-%the byte stream interface and a N:1 mapping strategy.
-%}
-%\end{figure}
-%
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-rand-read/output.read.60min.png}
-%  \caption{
-%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-rand-read/visualize.ipynb}{source}]
-%Throughput of 1K random reads to a single OSD using the standard I/O
-%interfaces in various configurations. The best performance is achieved using
-%the byte stream interface and a N:1 mapping strategy.
-%}
-%\end{figure}
-%
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=0.45\textwidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-%  \caption{
-%[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/basic-cls-overhead/visualize.ipynb}{source}]
-%}
-%\end{figure}
-
-%As a first approximation these results show that when optimizing for log
-%append and random read throughput the design space should be refined by
-%limiting solutions to architectures based on a N:1 mapping strategy using the
-%bytestream I/O interface. But what these microbenchmarks explicitly omit are
-%the overheads associated with metadata management such as enforcing write-once
-%semantics or using an index to map a log position to a physical offset within
-%an object. In the next section we
+Include performance management topic in future work.
 

### 2016-05-23 paper.tex
paper: adjust org
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\section{Background}
+\section{Background and Motivation}
 
@@
 example, a high-performance distributed shared-log built upon Ceph that makes
-extensive use of the object class facility, and then we present the challenges
-that application developers face when using this extensibility feature offered
-by the storage system.
+extensive use of the object class facility.
 
@@
 
-\paragraph*{Towards a CORFU Object Interface}
-The state-machine diagram in Figure~\ref{fig:corfu-sm} shows the composition of
-actions for each component of the CORFU interface. For instance, all operations
-begin by applying an \emph{epoch guard} that ensures the request is tagged with
-an up-to-date epoch value. The \emph{read} (R) and \emph{write} (W) operations
-both proceed by (1) examining metadata associated with the target position, (2)
-performing I/O to read or write the log entry, and in the case of a write, (3)
-updates metadata for the target log position.
+\section{Physical Design}
+
+As we have seen, Ceph provides a rich storage API and places few restrictions
+on the structure of applications. Thus a primary concern when implementing an
+object interface in Ceph is deciding how native interfaces are composed into
+compound operations in order to implement the semantics of the target
+interface. These types of decisions are commonly referred to as
+physical design, and can affect performance and application flexibility.
+As we will see the design space that developers must operate in is often
+large, and its dynamic nature can lead to design decisions that become
+obsolete and non-optimal.
 
@@
 
-The primary concern of an application developer when implementing an object
-interface in Ceph is deciding how native interfaces are composed into a
-compound operation. These types of decisions are commonly referred to as
-physical design, and can affect performance and application flexibility.  For
-instance, one valid design option is to store each log position in an object
-with a name using a one-to-one mapping with the log entry position.  This
-would simplify the design of the \emph{write} interface because a small amount
-of metadata stored as a header in the object could describe the state of the
-log entry.  However, as we will see in the next section this choice of a
-physical design can result in poor performance compared to other designs.
+To understand the developer process in the context of CORFU we have included the
+state-machine diagram in Figure~\ref{fig:corfu-sm} showing the composition of
+actions for each component of the CORFU interface which must be mapped onto
+Ceph object classes. For instance, all operations begin by applying an
+\emph{epoch guard} that ensures the request is tagged with an up-to-date epoch
+value. The \emph{read} (R) and \emph{write} (W) operations both proceed by (1)
+examining metadata associated with the target position, (2) performing I/O to
+read or write the log entry, and in the case of a write, (3) updates metadata
+for the target log position.
 
-\section{Physical Design}
+As an example, one valid design option is to store each log position in an
+object with a name using a one-to-one mapping with the log entry position.
+This would simplify the design of the \emph{write} interface because a small
+amount of metadata stored as a header in the object could describe the state
+of the log entry.  However, as we will see this choice of a physical design
+can result in poor performance compared to other designs. In the remainder of
+this section we will define the entire design space and use a set of targetted
+benchmarks to arrive at a final design. Finally we will show that these design
+decisions can lead to non-optimal decisions and suggest an automated approach
+is desirable.
 
-As we have seen, Ceph provides a rich storage API and places few restrictions
-on the structure of applications. So how should we go about implementing CORFU
-on Ceph? Our strategy is straightforward: first we will define the entire
-design space, and then winnow down this space using a set of targeted
-benchmarks. The design space can be divided into three challenges:
-selecting a strategy for log entry addressing, choosing a native I/O interface
-for storing log entry content, and implementing efficient metadata management.
+\subsection{Challenges}
+
+The design space can be divided into three challenges: selecting a strategy
+for log entry addressing, choosing a native I/O interface for storing log
+entry content, and implementing efficient metadata management.
 

### 2016-05-23 paper.tex

@@
 \usepackage{tabularx}
+\usepackage{caption}
 \usepackage{subcaption}
@@
 
-\subsection{Optimization}
+Application-level designs also contribute to an expanded design space.  In the
+N:1 design of mapping CORFU onto Ceph the potential amount of I/O parallelism
+is controlled by the value N, however if this value becomes too large there may
+be no benefit or performance could suffer from reduced data locality. Other
+aspects of application-level design are even more flexible. A wide variety of
+data structures could be serialized into the byte stream to handle metadata
+management tasks, but selecting a single external index that performs well
+across workloads and hardware may not be possible.
+
+\subsection{Cluster-Specific Optimizations}
 
@@
 process of design we have outlined may not yield such clear cut results when
-applied in a different context.
+applied in a different context. To illustrate the complexity and breadth of
+this design space, we describe the software, system tunables, and hardware
+parameters that affect the performance of ZLog.
 
-To illustrate this we ran the same benchmark that measures expected append
-throughput shown in Figure~\ref{fig:vanilla_wr_jewel} on identical hardware
-but we used a long-term support version of Ceph released two years
-ago. The result of this benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly}
-in which the bytestream interface outperforms the key-value interface by 12\%,
-compared to over 100\% in the newer versino of Ceph. Qualatatively, given the
-reduced overall throughput achieved in the older version, some developers may
-find that incurring the overhead of using the key-value interface is an easy
-decision given the reduced complexity of the design space for metadata
-management. If this choice had been made, then a future system upgrade could
-drop a significant number of iops on the floor.
+\subsubsection{Software Parameters}
 
-In practice the real world state space is much larger than the one described
-in this section. Ceph alone has hundreds of configuration options, significant
-features becoming available with releases each year, and is designed to run on
-a wide variety of COTS hardware as well as new NVMe devices, all of which
-contribute to a large and dynamic state space. Application-level designs
-also contribute to an expanded design space. In the N:1 design of mapping
-CORFU onto Ceph the potential amount of I/O parallelism is controlled by 
-the value N, however if this value becomes too large there may be no benefit
-or performance could suffer from reduced data locality. Other aspects of
-application-level design are even more flexible. A wide variety of data
-structures could be serialized into the byte stream to handle metadata
-management tasks, but selecting a single external index that performs well
-across workloads and hardware may not be possible.
+Ceph releases stable versions every year (Oct/Nov) and LTS versions every 3-4
+months~\cite{website:ceph-releases}. The head of the master branch moves
+quickly because there are over 400 contributors and an active mailing list.
+Over the calendar year, there were between  70 and 260 commits per
+week~\cite{website:ceph-github}. These releases contain performance fixes and
+new features that could directly ZLog.
+
+\textbf{Ceph Versions}: to show the affect that this has on ZLog's design, we
+ran the same benchmark that measures expected append throughput shown in
+Figure~\ref{fig:vanilla_wr_jewel} on identical hardware but instead use a
+long-term support version of Ceph released two years ago. The result of this
+benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly} in which the
+bytestream interface outperforms the key-value interface by 12\%, compared to
+over 100\% in the newer version of Ceph. Qualitatively, given the reduced
+overall throughput achieved in the older version, some developers may find that
+incurring the overhead of using the key-value interface is an easy decision
+given the reduced complexity of the design space for metadata management. If
+this choice had been made, then a future system upgrade could drop a
+significant number of iops on the floor.
+
+\textbf{Ceph Features}: to show the affect of cutting-edge Ceph features, we
+now show results for BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
+replacement for the FileStore file system in the OSD (traditionally XFS).
+FileStore has performance problems with transactions and enumerations; namely
+the journal needed to assure atomicity incurs double writes and the file system
+metadata model makes object listings slow, respectively. BlueStore stores data
+directly on a block device and the metadata in RocksDB, which is provided by a
+minimalistic, non-POSIX C++ filesystem. This model adheres to the overall
+software defined storage strategy of Ceph because it gives the administrator
+the flexiblity to store the 3 componenets of BlueStore (e.g., data, RocksDB
+database, and RocksDB write-ahead log) on any partition on any device in the
+OSD.
+
+\subsubsection{System Tunables}
+
+The most recent version of Ceph (v10.2.0-1281-g1f03205) has 994 tunable
+parameters\footnote{This number comes from \texttt{src/common/config\_opts.h}
+with debug options filtered out.}, where 195 of them pertain to the OSD itself
+and 95 of them focus on the OSD back end file system (i.e. its
+\texttt{filestore}). Ceph also has tunables for the subystems it uses, like
+LevelDB (10 tunables), RocksDB (5 tunables), its own key-value stores (5
+tunables), its object cache (6 tunables), its journals (24 tunables), and its
+other optional filestores like BlueStore (49 tunables).
+
+This many domain-specific tunables makes it almost impossible to come up with
+the best set of tunables, although auto-tuning like the work done
+in~\cite{behzad:sc2013-autotuning} could go a long way. Regardless of the
+technique that we use, it is clear the the number of tunables increases the
+physical design parameters to an unwieldy stae space size.
+
+\subsubsection{Hardware Parameters}
+
+Ceph is designed to run on a wide variety of commodity hardware as well as new
+NVMe devices, all of which their own set of characteristics and tunables (e.g.,
+the IO operation scheduler type). In our experiments, we tested SSD, HDDs, NVMe
+devices, and
+
+\subsection{Discussion}
+
+This physical design process, of first thinking about the design trade-offs and
+then doing parameters sweeps, is insufficient when general-purpose systems. Had
+we built ZLog from the ground up (like CORFU) the story might be different but
+to deal wtih the large state space we need a way to automatically and
+autonomically decide which parameters to choose.
 

### 2016-05-24 paper.tex
paper updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
   showstringspaces=false,
-  columns=flexible,
+  %columns=flexible,
   basicstyle={\scriptsize\ttfamily},
@@
   %stringstyle=\color{mauve},
-  breaklines=true,
+  %breaklines=true,
   breakatwhitespace=true,
@@
 
-\begin{lstlisting}[caption={Hello}, label=lab]
-class Corfu
-include Bud
-
-state do
-table :epoch, [:epoch]
-table :log, [:pos] => [:state, :data]
-
-scratch :valid_op, [:type, :pos, :epoch] => [:data]
-scratch :invalid_op, [:type, :pos, :epoch] => [:data]
-
-interface input, :op, [:type, :pos, :epoch] => [:data]
-interface output, :ret, [:type, :pos, :epoch] => [:retval]
+\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
+bloom :epoch_guard do
+  invalid <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
+  ret <= invalid{|o| [o.type, o.pos, o.epoch, 'stale']}
+  valid <= op.notin(invalid)
 end

### 2016-05-24 paper.tex
paper: add some code descriptions
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-Listing~\ref{lab} shows code.
+\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
+state do
+  table :epoch, [:epoch]
+  table :log, [:pos] => [:state, :data]
+
+  interface input,  :op,   [:type, :pos, :epoch] => [:data]
+  interface output, :ret,  [:type, :pos, :epoch] => [:retval]
+
+  scratch :write_op, [:type, :pos, :epoch] => [:data]
+  scratch :read_op,  [:type, :pos, :epoch] => [:data]
+  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
+  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
+  scratch :seal_op,  [:type, :pos, :epoch] => [:data]
+
+  # op did or did not pass the epoch guard
+  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
+  scratch :invalid_op, [:type, :pos, :epoch] => [:data]
+
+  # op's position was or was not found in the log
+  scratch :found_op,    [:type, :pos, :epoch] => [:data]
+  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
+end
+
+bloom do
+  # op's position found in log
+  found_op <= (valid_op * log).lefts(pos => pos)
+  notfound_op <= valid_op.notin(found_op)
+
+  # demux on operation type
+  write_op <= valid_op { |o| o if o.type == 'write' }
+  read_op  <= valid_op { |o| o if o.type == 'read'  }
+  fill_op  <= valid_op { |o| o if o.type == 'fill'  }
+  trim_op  <= valid_op { |o| o if o.type == 'trim'  }
+  seal_op  <= valid_op { |o| o if o.type == 'seal'  }
+end
+\end{lstlisting}
+
+Listing~\ref{lst:eg} shows the epoch guard that is applied to all operations.
+The guard rejects requests that are tagged with old epoch values, ensuring that
+a client generating a request has an up-to-date view of the system.  First the
+\texttt{invalid\_op} collection is defined to include the current operation if
+its epoch value is no larger than the stored epoch value. Next the
+\texttt{valid\_op} collection is defined to be the inverse of
+\texttt{invalid\_op} and is a helper used to refine other operations later in
+the dataflow. Finally we handle the case for all operations tagged with an
+out-of-date epoch by merging the \texttt{invalid\_op} set into the output
+\texttt{ret} collection.
 
@@
 bloom :epoch_guard do
-  invalid <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
-  ret <= invalid{|o| [o.type, o.pos, o.epoch, 'stale']}
-  valid <= op.notin(invalid)
+  invalid_op <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
+  valid_op <= op.notin(invalid_op)
+  ret <= invalid_op{|o| [o.type, o.pos, o.epoch, 'stale']}
+end
+\end{lstlisting}
+
+The process of sealing an object requires installing a new epoch value and
+returning the current maximum position written. Listing~\ref{lst:seal} 
+implements the seal interface by first removing the current epoch value and
+replacing it with the epoch value contained in the operation. Next an aggregate
+is computed over the log to find the maximum position written, and this value
+is returned.
+
+\begin{lstlisting}[caption={Seal}, label=lst:seal]
+bloom :seal do
+  epoch <- (seal_op * epoch).rights
+  epoch <+ seal_op { |o| [o.epoch] }
+  temp :maxpos <= log.group([], max(pos))
+  ret <= (seal_op * maxpos).pairs do |o, m|
+    [o.type, nil, o.epoch, m.content]
+  end
+end
+\end{lstlisting}
+
+Trimming a log entry always succeeds. In Listing~\ref{lst:trim} The
+\texttt{<+-} operator simulatenously removes the log entry with the given 
+position and replaces it with an entry with its state set to \emph{trimmed}.
+
+\begin{lstlisting}[caption={Trim}, label=lst:trim]
+bloom :trim do
+  log <+- trim_op{|op| [op.pos, 'trimmed']}
+  ret <=  trim_op{|op| [op.type, op.pos, op.epoch, 'ok']}
+end
+\end{lstlisting}
+
+The write and fill interfaces are implemented similarily, and are show in
+Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
+write collection is created if the operation position is not found in the log.
+A valid write operation is then merged into log, otherwise a read only error is
+returend indicating that hte log position was already written to.
+The fill operation is identical except a the fill state is set on the log entry.
+
+\begin{lstlisting}[caption={Write}, label=lst:write]
+bloom :write do
+  temp :valid_write <= write_op.notin(found_op)
+  log <+ valid_write { |o| [o.pos, 'valid', o.data] }
+  ret <= valid_write { |o| [o.type, o.pos, o.epoch, 'ok'] }
+  ret <= write_op.notin(valid_write) { |o|
+    [o.type, o.pos, o.epoch, 'read-only']
+  }
+end
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Fill}, label=lst:fill]
+bloom :fill do
+  temp :valid_fill <= fill_op.notin(found_op)
+  log <+ valid_fill { |o| [o.pos, 'fill'] }
+  ret <= valid_fill { |o| [o.type, o.pos, o.epoch, 'ok'] }
+  ret <= fill_op.notin(valid_fill) { |o|
+    [o.type, o.pos, o.epoch, 'read-only']
+  }
+end
+\end{lstlisting}
+
+The read interface is shown in Listing~\ref{lst:read}, and structured similar
+to the write interface. First we create a collection containing a valid read
+operation that is in the log and does not have the filled or trimmed state set.
+The data is returend in the case of a valid read op, otherwise an error is
+returend through the output interface.
+
+\begin{lstlisting}[caption={Read}, label=lst:read]
+bloom :read do
+  temp :valid_read <= (read_op * log).pairs(pos => pos) { |o, l|
+    [o.type, o.pos, o.epoch, l.data] unless
+      ['filled', 'trimmed'].include?(l.state)
+  }
+  ret <= valid_read { |e| [e.type, e.pos, e.epoch, e.data] }
+  ret <= read_op.notin(valid_read, type => type) do |o|
+    [o.type, o.pos, o.epoch, 'invalid']
+  end
 end

### 2016-05-24 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -3,21 +3,22 @@
 reset=soft
 data_dev=sdc
 noop_dev=sdc
-ceph_version=firefly
+ceph_version=jewel
+bluestore=yes
 
-name=librados_firefly_ssd_small
-runtime=1800
+name=librados_jewel_bluestore_ssd_small
+runtime=900
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
 entry_sizes="128"
 pool=zlog
 rest=120
-read_runtime=1800
+read_runtime=900
 reads=yes
 
 # workloads
-wl_11="map_11 bytestream_11"
+#wl_11="map_11 bytestream_11"
 wl_n1="map_n1 bytestream_n1_write"
 workloads="$wl_11 $wl_n1"
 

### 2016-05-24 experiments/basic-cls-rand-read/run2.sh

@@ -3,21 +3,22 @@
 reset=soft
 data_dev=sdc
 noop_dev=sdc
-ceph_version=firefly
+ceph_version=jewel
+bluestore=yes
 
-name=librados_firefly_ssd_med
-runtime=1800
+name=librados_jewel_bluestore_ssd_med
+runtime=900
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
 entry_sizes="1024"
 pool=zlog
 rest=120
-read_runtime=1800
+read_runtime=900
 reads=yes
 
 # workloads
-wl_11="map_11 bytestream_11"
+#wl_11="map_11 bytestream_11"
 wl_n1="map_n1 bytestream_n1_write"
 workloads="$wl_11 $wl_n1"
 

### 2016-05-24 experiments/basic-cls-rand-read/run3.sh

@@ -3,21 +3,22 @@
 reset=soft
 data_dev=sdc
 noop_dev=sdc
-ceph_version=firefly
+ceph_version=jewel
+bluestore=yes
 
-name=librados_firefly_ssd_large
-runtime=1800
+name=librados_jewel_bluestore_ssd_large
+runtime=900
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
 entry_sizes="4096"
 pool=zlog
 rest=120
-read_runtime=1800
+read_runtime=900
 reads=yes
 
 # workloads
-wl_11="map_11 bytestream_11"
+#wl_11="map_11 bytestream_11"
 wl_n1="map_n1 bytestream_n1_write"
 workloads="$wl_11 $wl_n1"
 

### 2016-05-24 experiments/basic-cls-rand-read/run1.sh
expr: add expr
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 ceph_version=jewel
 bluestore=yes
 
-name=librados_jewel_bluestore_ssd_small
+name=librados_jewel_bluestore_hdd_small
 runtime=900
 pg_nums="128"
 stripe_widths="128"

### 2016-05-24 experiments/basic-cls-rand-read/run2.sh

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 ceph_version=jewel
 bluestore=yes
 
-name=librados_jewel_bluestore_ssd_med
+name=librados_jewel_bluestore_hdd_med
 runtime=900
 pg_nums="128"
 stripe_widths="128"

### 2016-05-24 experiments/basic-cls-rand-read/run3.sh

@@ -1,12 +1,11 @@
 #!/bin/bash
 
 reset=soft
-data_dev=sdc
-noop_dev=sdc
+data_dev=sdb
 ceph_version=jewel
 bluestore=yes
 
-name=librados_jewel_bluestore_ssd_large
+name=librados_jewel_bluestore_hdd_large
 runtime=900
 pg_nums="128"
 stripe_widths="128"

### 2016-05-24 experiments/basic-cls-rand-read-bluestore/run1.sh
expr: bluestore with NVMe for all entry sizes
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -7,23 +7,21 @@ noop_dev="nvme0n1 --noop nvme1n1"
 osd_create="--bluestore"
 
 name=librados_read_data_jrnl_nvme_bluestore
-runtime=1800
-#runtime=500
+runtime=3600
 pg_nums="128"
 stripe_widths="128"
 queue_depths="32"
-entry_sizes="1024"
+entry_sizes="128 1024 4096"
 pool=zlog
 rest=120
-#read_runtime=0
-read_runtime=1800
+read_runtime=3600
 reads=yes
 
 # workloads
 wl_11="map_11 bytestream_11"
 wl_n1="map_n1 bytestream_n1_write"
 #workloads="$wl_11 $wl_n1"
-workloads="$wl_n1"
+workloads="$wl_n1 $wl_11"
 
 # i/o interfaces
 map_n1_if="vanilla"

### 2016-05-24 paper.tex
paper: update fig
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
paper: udpate
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\title{outline!}
+\title{brados: declarative, programmable object storage}
 
 \author{
-%Noah Watkins, Neha Ojha\textsuperscript{*}, Carlos Maltzahn \\
-%\small {\em University of California, Santa Cruz} \\
+Noah Watkins, Michael Sevilla, Ivo Jimenez, \\Neha Ojha, Peter Alvaro, Carlos
+Maltzahn\\
+\small {\em University of California, Santa Cruz} \\
 %\small {\{jayhawk,carlosm\}@cs.ucsc.edu} \textsuperscript{*}nojha@ucsc.edu \\ [2mm]
@@
       \includegraphics[width=1.0\linewidth]{figures/ceph}
-      \caption{Ceph}
+      \caption{Ceph Cluster}
       \label{fig:ceph}
@@
       \includegraphics[width=1.0\linewidth]{figures/osd}
-      \caption{OSD}
+      \caption{OSD Arch}
       \label{fig:osd}
   \end{subfigure}
-  \caption{Ceph and OSD... gotta cram more stuff in here...}
+  \caption{The Ceph cluster is composed of object storage devices (OSD) that
+  support object transactions and a rich set of hardware and software
+configurations.}
 \end{figure}
@@
 
-\section{Programming Model}
+\section{The Brados Programming Model}
 
-Ceph is turning into a target for building distrubted services, bringing
-distributed programming into the hands of many new people that may have
-worked on the periphery of ceph.
+As Ceph continues to develop into a target for building distributed services
+and application, it is important to not restrict use to developers seasonsed
+in distributed programming and those with knowledge of intracacies of the Ceph
+performance model.  To this end we present Brados, a declarative language with
+relational semantics for defining the behavior of RADOS object interfaces.
 
-Getting consistency and coordination correct is a challenge And as we have seen
-optimization at a low-level given a simple set of semantics can yield a large
-state space from which to draw designs Difficulty and challenges of writing
-distributed software.  Programming distributed systems and applications.
+The Brados model corresponds to a subset of the Bloom language, and models the
+storage system state uniformly as a collection of
+relations~\cite{alvaro:cidr11}. The composition of a collection of existing
+interfaces is then expressed as a collection of high-level \emph{queries} that
+describe how a stream of requests (API calls) are filtered, transformed and
+combined with existing state to define streams of outputs (API call returns as
+well as updates to system state).  Separating the relational semantics of such
+compositions from details of their physical implementations introduces a
+number of degrees of freedom, similar to the ``data independence’’ offered by
+database query languages.  The choice of access methods (for example, whether
+to use a bytestream interface or a key/value store), storage decide
+classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
+1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
+postponed to execution time.  The optimal choices for these physical design
+details are likely to change much more often than the logical specification,
+freeing the interface designer from the need to rewrite systems and device and
+interface characteristics change.
 
-We draw upon the recent work of Bloom a declarative lang thing
+\subsection{Basics}
 
-BLL (a.k.a. BRADOS) is a declarative language with relational (or tabular)
-semantics for RADOS objects, which corresponds to a subset of the BLOOM
-language (leaving out temporal logic). BLL borrows from BLOOM the table and
-interface collections, the instantaneous implication (‘<=’) operator, as well
-as all the collection operators.
-
-BRADOS models the storage system state uniformly as a collection of relations.
-The composition of a collection of existing interfaces is then expressed as a
-collection of high-level \emph{queries} that describe how a stream of requests
-(API calls) are filtered, transformed and combined with existing state to
-define streams of outputs (API call returns as well as updates to system
-state).  Separating the relational semantics of such compositions from details
-of their physical implementations introduces a number of degrees of freedom,
-similar to the ``data independence’’ offered by database query languages.  The
-choice of access methods (for example, whether to use a bytestream interface or
-a key/value store), storage decide classes(e.g., whether to use HDDs or SDDs),
-physical layout details (e.g. a 1:1 or 1:N mapping) and execution plans (e.g.
-operator ordering) can be postponed to execution time.  The optimal choices for
-these physical design details are likely to change much more often than the
-logical specification, freeing the interface designer from the need to rewrite
-systems and device and interface characteristics change.
-
-\subsection{The CORFU Interface in Bloom}
+\subsection{The CORFU Log Interface}
 

### 2016-05-24 paper.tex
paper update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 yet reached the point of directly exposing these features to
-non-administrative users, the recent inclusion of a mechanism for dynamically
+non-administrative users, the under review inclusion of a mechanism for dynamically
 defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
@@
 % why hasn't it been solved before / why have previous approaches failed
+Unfortunately the size of the design space for even simple interfaces can
+be very large, and it can be difficult to choose a design that future proofs an
+implementation against future hardware and software upgrades.
 Previous work related to storage interface design has largely been in the
-context of standardization efforts and active storage. While the later has
-been specifically concerned with the creation of application-specific
-interfaces and demonstrating short-term benefits, the standardization seeks to
-develop a last set of interfaces. In neither case have we seen efforts focused
-on extensible interfaces with portability and longevity a primary concern.
+context of standardization efforts and active storage. While the former is
+primarily concerned with fixed interfaces, application-specific interfaces
+built using active storage techniques have not to the best of our knoweldge
+addressed portability concerne that arise from defining interfaces
+imperatively. We address this gap by using a declarative language we call
+\emph{brados} to define application-specific storage interfaces in
+object-based storage systems, decoupling interface definition from
+implementation and allowing interfaces to adapt to system changes without
+modification.
 
-%This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
-%KV-Drives, and Rhea, are all recent examples of storage systems embracing the
-%power of including application-specific semantics.
-
-We propose the introduction of a declarative language for building object
-based interfaces that allows a storage system to meet the needs of an
-application throughout the development process without requiring rewrites.
+To demonstrate the use of the \emph{brados} language we will show how a
+high-performance shared-log based on the CORFU protocol can be built, as well
+as existing real-world interfaces used in the Ceph storage system. First we
+discuss the Ceph storage system which our prototype is built upon and the 
+motivating system, CORFU. We then present the challenges that programmers face
+when navigating the design space during the process of building an
+application-storage interface. The \emph{brados} language is then described
+using the CORFU system and an existing Ceph-specific interface as examples.
+We conclude by discussing the optimization opportunities that we can realize
+using analysis of interfaces. I hope you enjoy the show.
 
@@
 
-\section{Physical Design}
+\section{Physical Design Space}
 
@@
 relational semantics for defining the behavior of RADOS object interfaces.
+Brados uses a simple declarative data model and can reduce the learning curve
+for new users, and allows existing developers to increase productivity by
+writing less code that is more portable.
 
-The Brados model corresponds to a subset of the Bloom language, and models the
+The Brados language corresponds to a subset of the Bloom language, and models the
 storage system state uniformly as a collection of

### 2016-05-25 paper.tex
Adds some refs on declarative languages

@@
 
+{\bf Declarative Specifications.}
+Declarative interfaces have been used for specifying distributed analytic 
+tasks~\cite{olston_pig_2008,thusoo_hive_2010}, information 
+extraction~\cite{shen_declarative_2007} and distributed 
+systems~\cite{conway_logic_2012,bauer_legion_2014}. The use of declarative 
+specifications has also been explored in other contexts such as cloud recovery 
+testing~\cite{gunawi_fate_2011}, bug reproduction~\cite{li_reprolite_2014} and 
+cloud resource orchestration~\cite{liu_declarative_2011}.
+
 \section{Conclusion and Future Work}

### 2016-05-25 paper.tex
paper: section 3 graph parm sweep
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-24 paper.tex
@@
 
-\title{brados: declarative, programmable object storage}
+\title{outline!}
 
 \author{
-Noah Watkins, Michael Sevilla, Ivo Jimenez, \\Neha Ojha, Peter Alvaro, Carlos
-Maltzahn\\
-\small {\em University of California, Santa Cruz} \\
+%Noah Watkins, Neha Ojha\textsuperscript{*}, Carlos Maltzahn \\
+%\small {\em University of California, Santa Cruz} \\
 %\small {\{jayhawk,carlosm\}@cs.ucsc.edu} \textsuperscript{*}nojha@ucsc.edu \\ [2mm]
@@
       \includegraphics[width=1.0\linewidth]{figures/ceph}
-      \caption{Ceph Cluster}
+      \caption{Ceph}
       \label{fig:ceph}
@@
       \includegraphics[width=1.0\linewidth]{figures/osd}
-      \caption{OSD Arch}
+      \caption{OSD}
       \label{fig:osd}
   \end{subfigure}
-  \caption{The Ceph cluster is composed of object storage devices (OSD) that
-  support object transactions and a rich set of hardware and software
-configurations.}
+  \caption{Ceph and OSD... gotta cram more stuff in here...}
 \end{figure}
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\centering
-  \begin{subfigure}[b]{.47\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
-      \caption{1hr,1k,wr,jewel}
-      \label{fig:vanilla_wr_jewel}
-  \end{subfigure}
-  \begin{subfigure}[b]{.47\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-      \caption{1hr,1k,rd,jewel}
-      \label{fig:vanilla_rd_jewel}
-  \end{subfigure}
-  \begin{subfigure}[b]{.47\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-      \caption{1hr,1k,cls\_wr,jewel}
-      \label{fig:cls_wr_jewel}
-  \end{subfigure}
-  \begin{subfigure}[b]{.47\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/firefly_ssd.png}
-      \caption{1hr,1k,wr,firefly}
-      \label{fig:vanilla_wr_firefly}
-  \end{subfigure}
-  \caption{placeholder.}
-\end{figure}
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel_v_firefly_v_es.png}
+	\caption{The cluster-specific optimizations depend heavily on the hardware and software of the storage system. This caption is quite terrible.}
+\end{figure*}
+%\begin{figure}[t]
+%  \centering
+%  \begin{subfigure}[b]{.47\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
+%      \caption{1hr,1k,wr,jewel}
+%      \label{fig:vanilla_wr_jewel}
+%  \end{subfigure}
+%  \begin{subfigure}[b]{.47\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+%      \caption{1hr,1k,rd,jewel}
+%      \label{fig:vanilla_rd_jewel}
+%  \end{subfigure}
+%  \begin{subfigure}[b]{.47\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+%      \caption{1hr,1k,cls\_wr,jewel}
+%      \label{fig:cls_wr_jewel}
+%  \end{subfigure}
+%  \begin{subfigure}[b]{.47\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/firefly_ssd.png}
+%      \caption{1hr,1k,wr,firefly}
+%      \label{fig:vanilla_wr_firefly}
+%  \end{subfigure}
+%  \caption{placeholder.}
+%\end{figure}
 
@@
 
-\section{The Brados Programming Model}
+\section{Programming Model}
 
-As Ceph continues to develop into a target for building distributed services
-and application, it is important to not restrict use to developers seasonsed
-in distributed programming and those with knowledge of intracacies of the Ceph
-performance model.  To this end we present Brados, a declarative language with
-relational semantics for defining the behavior of RADOS object interfaces.
+Ceph is turning into a target for building distrubted services, bringing
+distributed programming into the hands of many new people that may have
+worked on the periphery of ceph.
 
-The Brados model corresponds to a subset of the Bloom language, and models the
-storage system state uniformly as a collection of
-relations~\cite{alvaro:cidr11}. The composition of a collection of existing
-interfaces is then expressed as a collection of high-level \emph{queries} that
-describe how a stream of requests (API calls) are filtered, transformed and
-combined with existing state to define streams of outputs (API call returns as
-well as updates to system state).  Separating the relational semantics of such
-compositions from details of their physical implementations introduces a
-number of degrees of freedom, similar to the ``data independence’’ offered by
-database query languages.  The choice of access methods (for example, whether
-to use a bytestream interface or a key/value store), storage decide
-classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
-1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
-postponed to execution time.  The optimal choices for these physical design
-details are likely to change much more often than the logical specification,
-freeing the interface designer from the need to rewrite systems and device and
-interface characteristics change.
+Getting consistency and coordination correct is a challenge And as we have seen
+optimization at a low-level given a simple set of semantics can yield a large
+state space from which to draw designs Difficulty and challenges of writing
+distributed software.  Programming distributed systems and applications.
 
-\subsection{Basics}
+We draw upon the recent work of Bloom a declarative lang thing
 
-\subsection{The CORFU Log Interface}
+BLL (a.k.a. BRADOS) is a declarative language with relational (or tabular)
+semantics for RADOS objects, which corresponds to a subset of the BLOOM
+language (leaving out temporal logic). BLL borrows from BLOOM the table and
+interface collections, the instantaneous implication (‘<=’) operator, as well
+as all the collection operators.
+
+BRADOS models the storage system state uniformly as a collection of relations.
+The composition of a collection of existing interfaces is then expressed as a
+collection of high-level \emph{queries} that describe how a stream of requests
+(API calls) are filtered, transformed and combined with existing state to
+define streams of outputs (API call returns as well as updates to system
+state).  Separating the relational semantics of such compositions from details
+of their physical implementations introduces a number of degrees of freedom,
+similar to the ``data independence’’ offered by database query languages.  The
+choice of access methods (for example, whether to use a bytestream interface or
+a key/value store), storage decide classes(e.g., whether to use HDDs or SDDs),
+physical layout details (e.g. a 1:1 or 1:N mapping) and execution plans (e.g.
+operator ordering) can be postponed to execution time.  The optimal choices for
+these physical design details are likely to change much more often than the
+logical specification, freeing the interface designer from the need to rewrite
+systems and device and interface characteristics change.
+
+\subsection{The CORFU Interface in Bloom}
 

### 2016-05-25 paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
   language=Ruby,
-  captionpos=b,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  %columns=flexible,
-  basicstyle={\scriptsize\ttfamily},
-  numbers=none,
-  %numberstyle=\tiny\color{gray},
-  %keywordstyle=\color{blue},
-  %commentstyle=\color{dkgreen},
-  %stringstyle=\color{mauve},
-  %breaklines=true,
-  breakatwhitespace=true,
-  tabsize=3
+  basicstyle={\scriptsize\sffamily},
+  breaklines=true,
+  captionpos=b
 }
+
+%  aboveskip=3mm,
+%  belowskip=3mm,
+%  showstringspaces=false,
+%  %columns=flexible,
+%  basicstyle={\scriptsize\ttfamily},
+%  numbers=none,
+%  %numberstyle=\tiny\color{gray},
+%  %keywordstyle=\color{blue},
+%  %commentstyle=\color{dkgreen},
+%  %stringstyle=\color{mauve},
+%  %breaklines=true,
+%  breakatwhitespace=true,
+%  tabsize=3
+%}
 
@@
 bloom :trim do
-  log <+- trim_op{|op| [op.pos, 'trimmed']}
-  ret <=  trim_op{|op| [op.type, op.pos, op.epoch, 'ok']}
+  log <+- trim_op{|o| [o.pos, 'trimmed']}
+  ret <=  trim_op{|o| [o.type, o.pos, o.epoch, 'ok']}
 end
@@
   temp :valid_write <= write_op.notin(found_op)
-  log <+ valid_write { |o| [o.pos, 'valid', o.data] }
-  ret <= valid_write { |o| [o.type, o.pos, o.epoch, 'ok'] }
+  log <+ valid_write{|o| [o.pos, 'valid', o.data]}
+  ret <= valid_write{|o| [o.type, o.pos, o.epoch, 'ok']}
   ret <= write_op.notin(valid_write) { |o|
@@
 
-\subsection{Optimizer}
+\subsection{Optimization}
+
+Choosing implementation details
+  - the refcount example is good. small data kept in ceph object bytestream
+until it becomes large then use omap.
+
+  - could use the dataflow graph to do interesting things like understand how
+  to do prefecting of things like keys or data that are often only known at
+  execution time
 
@@
 
-The subset of the Bloom language we are using is powerful enough to express the
-CORFU language, as well as many of the object class interfaces that are already
-included in Ceph. We briefly sketch the Bloom imlementation of an object class
-that provides a locking service. This is an interesting example because Bloom
-implementations use mechanisms for introducing points of coordination in
-execution, which in the context of Ceph, may be written in Bloom itself.
+As we have seen the subset of Bloom used in Brados is powerful enough to
+express the semantics of the CORFU storage interface. We briefly sketch the
+implementation of an additional interface found in Ceph and used by the
+S3-compatible RADOS-Gateway service.
+
+Channel
+C++ code comparison
+
+The subset of the Bloom language we are using in Brados is powerful enough to
+express the CORFU language, as well as many of the object class interfaces that
+are already included in Ceph. We briefly sketch the Bloom imlementation of an
+object class that provides a locking service. This is an interesting example
+because Bloom implementations use mechanisms for introducing points of
+coordination in execution, which in the context of Ceph, may be written in
+Bloom itself.
 
 \begin{lstlisting}[caption={Hello}, label=lab]
-bloom :take_lock do
+class RefCount
+  include Bud
+
+  state do
+    table :refs, [:tag]
+    channel :remove_chn
+    interface input, :get_op, [:tag]
+    interface input, :put_op, [:tag]
+    interface output, :ret, [:status]
   end
 
-bloom :break_lock do
+  bloom :get do
+    refs <= get_op
+    ret  <= get_op{|op| ['ok']}
   end
 
-bloom :list_locks do
+  bloom :put do
+    temp :found <= (put_op * refs).matches
+    refs <- found
+    ret <= put_op{|op| ['ok']}
+
+    temp :remove <= found.notin(refs)	
+    remove_chn <~ remove
   end
-
-bloom :lock_info do
 end

### 2016-05-25 paper.tex
paper: add final graphs to sec3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-25 paper.tex
@@
 	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel_v_firefly_v_es.png}
-	\caption{The cluster-specific optimizations depend heavily on the hardware and software of the storage system. This caption is quite terrible.}
+	\caption{The cluster-specific optimizations impact performance... a lot. This caption is quite terrible.}
 \end{figure*}
+
+\begin{figure}[t]
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+      \caption{This graph shows that a good choice for writes is also good for reads.}
+\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+	\caption{1hr,1k,cls\_wr,jewel: this graph shows the overhead of class}
+	\label{fig:cls_wr_jewel}
+\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel-hdd-128b.png}
+	\caption{shows that with a HDD, it don't matter if you choose omap!!!!!}
+	\label{fig:cls_wr_jewel}
+\end{figure}
 %\begin{figure}[t]

### 2016-05-25 paper.tex
paper: fix sec3.2/3.3 links
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-25 paper.tex
@@
 
-Figure~\ref{fig:vanilla_wr_jewel} shows the expected performance of 1K log
-appends without metadata management overhead using each of the five strategies
-defined in Table~\ref{tab:pd-map}. The first thing to notice in this figure is
-that both one-to-one strategies have relatively poor performance. The large
-period of reduced throughput for \emph{11wr} corresponds to the OSD splitting
-file system directories in order to maintain a maximum directory size, and
-will occur in \emph{11kv} although the threshold number of objects is not
-reached in this example due to reduced overall throughput of the \emph{11kv}
-strategy.
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel_v_firefly_v_es.png}
+	\caption{The cluster-specific optimizations impact performance... a lot. This caption is quite terrible.
+	\label{fig:jewel_v_firefly_v_es}}
+\end{figure*}
 
-The second lesson that we can learn from Figure~\ref{fig:vanilla_wr_jewel} is
-that even when using an N:1 addressing strategy, the key-value interface
-imposes a large overhead. This is unfortunate because the key-value interface
-can provide a direct solution to addressing log entries within an object.
-Instead, what we find is that an N:1 addressing strategy that stores log
-entries in the bytestream using either object appends or writes to explicit
-offsets outperform all other strategies by a factor of over 2x.
+The top row of Figure~\ref{fig:jewel_v_firefly_v_es} shows the expected
+performance of different sized log appends without metadata management overhead
+using each of the five strategies defined in Table~\ref{tab:pd-map}.  Starting
+with Jewel and 4KB entry sizes, we see that both one-to-one strategies have
+relatively poor performance. Jewel with 1KB entry sizes sees a 2.5\(\times\)
+performance increase for the N:1 KV mapping. The large period of reduced
+throughput for corresponds to the OSD splitting file system directories in
+order to maintain a maximum directory size, and will occur in \emph{1:1, KV}
+although the threshold number of objects is not reached in this example due to
+the reduced overall throughput of the \emph{1:1, KV} strategy.
+
+The second lesson that we can learn from the Jewel row of
+Figure~\ref{fig:jewel_v_firefly_v_es} is that even when using an N:1 addressing
+strategy, the key-value interface imposes a large overhead. This is unfortunate
+because the key-value interface can provide a direct solution to addressing log
+entries within an object.  Instead, what we find is that an N:1 addressing
+strategy that stores log entries in the bytestream using either object appends
+or writes to explicit offsets outperform all other strategies by a factor of
+over 2\(\times\).
+
+\begin{figure}[t]
+      \centering
+      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+      \caption{The expected random read performance from a log with 1KB entries
+      performs best on the same strategy that the write performance performs
+      best on - the N:1 bystream append strategy.}
+      \label{fig:vanilla_rd_jewel}
+\end{figure}
 
@@
 
-\begin{figure*}[t]
-	\centering
-	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel_v_firefly_v_es.png}
-	\caption{The cluster-specific optimizations impact performance... a lot. This caption is quite terrible.}
-\end{figure*}
-
-\begin{figure}[t]
-      \centering
-      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-      \caption{This graph shows that a good choice for writes is also good for reads.}
-\end{figure}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-	\caption{1hr,1k,cls\_wr,jewel: this graph shows the overhead of class}
-	\label{fig:cls_wr_jewel}
-\end{figure}
 \begin{figure}[t]
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+	\caption{1hr,1k,cls\_wr,jewel: this graph shows the overhead of class}
+	\label{fig:cls_wr_jewel}
+\end{figure}
+
 We consider two high-level design prototypes in this paper based on both the
@@
 process of design we have outlined may not yield such clear cut results when
-applied in a different context. To illustrate the complexity and breadth of
-this design space, we describe the software, system tunables, and hardware
-parameters that affect the performance of ZLog.
+applied in a different context. To illustrate these contexts, we describe the 
+software, system tunables, and hardware parameters that affect the performance 
+of ZLog.
 
@@
 quickly because there are over 400 contributors and an active mailing list.
-Over the calendar year, there were between  70 and 260 commits per
-week~\cite{website:ceph-github}. These releases contain performance fixes and
-new features that could directly ZLog.
+Over the past calendar year, there were between 70 and 260 commits per
+week~\cite{website:ceph-github}.
 
-\textbf{Ceph Versions}: to show the affect that this has on ZLog's design, we
-ran the same benchmark that measures expected append throughput shown in
-Figure~\ref{fig:vanilla_wr_jewel} on identical hardware but instead use a
-long-term support version of Ceph released two years ago. The result of this
-benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly} in which the
+\textbf{Ceph Versions}: to show the evolution of Ceph's performance and
+behavior, we ran the same benchmark, which measures append throughput, with the
+same configurations, hardware, and tunables on 2 versions of Ceph: Jewel (April
+2016) - the newest stable release and Firefly (May 2014)- long term support
+version that introduced cache tiering and erasure coding.
+Figure~\ref{fig:jewel_v_firefly_v_es} shows the behavior and performance of the
+benchmark for 128b, 1KB, and 4KB objects.
+
+For all graphs,
+
+That figure demonstrates that the 
 bytestream interface outperforms the key-value interface by 12\%, compared to

### 2016-05-25 paper.tex
paper update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 yet reached the point of directly exposing these features to
-non-administrative users, the under review inclusion of a mechanism for dynamically
-defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
+non-administrative users, the inclusion of a mechanism for dynamically
+defining extensions using Lua~\cite{cls_lua} (currently pending review)
+suggests that aspects of this
 feature may soon appear. What is needed is support for creating storage
@@
 be very large, and it can be difficult to choose a design that future proofs an
-implementation against future hardware and software upgrades.
+implementation against hardware and software upgrades.
 Previous work related to storage interface design has largely been in the
@@
 built using active storage techniques have not to the best of our knoweldge
-addressed portability concerne that arise from defining interfaces
+addressed portability concerns that arise from defining interfaces
 imperatively. We address this gap by using a declarative language we call
@@
 To demonstrate the use of the \emph{brados} language we will show how a
-high-performance shared-log based on the CORFU protocol can be built, as well
+high-performance shared-log based on the CORFU protocol can be built in Ceph, as well
 as existing real-world interfaces used in the Ceph storage system. First we
@@
 application-storage interface. The \emph{brados} language is then described
-using the CORFU system and an existing Ceph-specific interface as examples.
+using the CORFU system and an existing Ceph-specific interface as motivating examples.
 We conclude by discussing the optimization opportunities that we can realize
-using analysis of interfaces. I hope you enjoy the show.
-
-\section{Background and Motivation}
-
-In this section we highlight the salient components of Ceph, especially its
-\emph{object class} feature that offers users the ability to load and execute
-application-specific codes. We provide a description of our motivating
-example, a high-performance distributed shared-log built upon Ceph that makes
-extensive use of the object class facility.
-
-\subsection{Ceph and Storage Programmability}
+using analysis of interfaces. We hope you enjoy the show.
+
+\section{Background}
+
+In this section we describe the salient components of Ceph that we use to
+construct application-specific extensions, and provide an overview of the
+CORFU distributed shared-log which is the primary motivating example used in
+this paper. The content of this section sets the stage for a discussion of the
+complexities developers face building application-specific storage interfaces.
+
+\subsection{Ceph Basics and Storage Interfaces}
 \label{sec:objclass}
@@
       \includegraphics[width=1.0\linewidth]{figures/osd}
-      \caption{OSD Arch}
+      \caption{OSD}
       \label{fig:osd}
@@
   \caption{The Ceph cluster is composed of object storage devices (OSD) that
-  support object transactions and a rich set of hardware and software
-configurations.}
+  provide access to objects on behalf of client requests. An OSD supports a
+    wide variety of hardware and software configurations, as well as the
+    transactional composition of object operations for defining
+    application-level interfaces.}
 \end{figure}
@@
 on the type of data being accessed: key-value items, or bulk bytestream data.
-The key-value interface operates as a dedicated database associated with each
+The key-value interface operates as an isolated database associated with each
 object, and the bytestream interface supports random byte-level access similar
-to a file.  At a
-low-level each of these abstract I/O interfaces map to hardware storage
-devices through a pluggable object backend storage service. For instance,
-LevelDB or RocksDB may be used to store key-value data, while the
-\emph{FileStore} implementation maps the bytestream interface onto a local
-POSIX file system~\cite{leveldb,rocksdb}. Several backend implementations
-exist for storing data in targets such as the Kinetic Drive or in NVMe devices,
-among others.
+to a standard file interface.  At a low-level each of these abstract I/O
+interfaces map to hardware storage devices through a pluggable object backend
+shown in Figure~\ref{fig:osd}. For instance, LevelDB or RocksDB may be used to
+store key-value data, while the \emph{FileStore} implementation maps the
+bytestream interface onto a local POSIX file system~\cite{leveldb,rocksdb} via
+XFS or other supported file systems. Several backend implementations exist for
+storing data in a range of targets such as HDD, SSD, as well as
+Ethernet-attached disks, and NVMe devices using optimized access libraries.
+%and in SMR disks using a specialized backend called BlueStore.
 

### 2016-05-25 paper.tex
paper: add dpic
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-25 paper.tex
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel-hdd-128b.png}
-	\caption{shows that with a HDD, it don't matter if you choose omap!!!!!}
-	\label{fig:cls_wr_jewel}
-\end{figure}
-%\begin{figure}[t]
-%  \centering
-%  \begin{subfigure}[b]{.47\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
-%      \caption{1hr,1k,wr,jewel}
-%      \label{fig:vanilla_wr_jewel}
-%  \end{subfigure}
-%  \begin{subfigure}[b]{.47\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-%      \caption{1hr,1k,rd,jewel}
-%      \label{fig:vanilla_rd_jewel}
-%  \end{subfigure}
-%  \begin{subfigure}[b]{.47\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
-%      \caption{1hr,1k,cls\_wr,jewel}
-%      \label{fig:cls_wr_jewel}
-%  \end{subfigure}
-%  \begin{subfigure}[b]{.47\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/firefly_ssd.png}
-%      \caption{1hr,1k,wr,firefly}
-%      \label{fig:vanilla_wr_firefly}
-%  \end{subfigure}
-%  \caption{placeholder.}
-%\end{figure}
-
 \subsection{Metadata Management}
@@
 applied in a different context. To illustrate these contexts, we describe the
-software, system tunables, and hardware parameters that affect the performance 
-of ZLog.
+software, system tunable, and hardware parameters that affect the performance
+of our CORFU implementation. The complexity and breadth of our parameter sweep,
+and the fact that our state space grew so large, motivates brados in section 4.
 
@@
 same configurations, hardware, and tunables on 2 versions of Ceph: Jewel (April
-2016) - the newest stable release and Firefly (May 2014)- long term support
-version that introduced cache tiering and erasure coding.
-Figure~\ref{fig:jewel_v_firefly_v_es} shows the behavior and performance of the
-benchmark for 128b, 1KB, and 4KB objects.
-
-For all graphs,
-
-That figure demonstrates that the 
-bytestream interface outperforms the key-value interface by 12\%, compared to
-over 100\% in the newer version of Ceph. Qualitatively, given the reduced
+2016) - the newest stable release and Firefly (May 2014) - a long term support
+version that introduced cache tiering and erasure coding. The Jewel row of 
+Figure~\ref{fig:jewel_v_firefly_v_es} shows that the size of the log entries
+has a clear impact on performance but the Firefly release shows no such trend.
+The fact that the appends per second are an order of magnitude slower on
+Firefly indicates that the the benchmark is bottlenecked by the Ceph software
+instead of the external log implementation.
+
+The bytestream interface outperforms the key-value interface by 12\%, compared
+to over 100\% in the newer version of Ceph. Qualitatively, given the reduced
 overall throughput achieved in the older version, some developers may find that
@@
 
+\textbf{Takeway}: choosing the best implementations is dependent on both the
+timing of the developemnt (Ceph Version) and the expertise of the administrator
+(Ceph Features). Different versions and features of Ceph may lead the
+administrator to choose a suboptimal implementation for the system's next
+upgrade. The software parameters must be accounted for and benchmarked when
+making design decisions.
+
 \subsubsection{System Tunables}
@@
 
+\textbf{Takeaway}: the number and complexity of Ceph's tunables make
+brute-force parameter selection impossible.
+
 \subsubsection{Hardware Parameters}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel-hdd-128b.png}
+	\caption{shows that with a HDD, it don't matter if you choose omap!!!!!}
+	\label{fig:jewel-hdd-128b}
+\end{figure}
 
@@
 the IO operation scheduler type). In our experiments, we tested SSD, HDDs, NVMe
-devices, and
+devices and discovered a wide range of behaviors and performance profiles.
+Figure~\ref{fig:jewel-hdd-128b}, which shows the write performance of 128 by
+log entries using Jewel and a single HDD, shows an especially tricky situation
+for a developer.
+
+In that figure, the performance is 10\(\times\) slower than its SSD counterpart
+but performance of the N:1 bystestream implementation is almost identical to
+the N:1 KV implementation (within 1\% mean throughput). Regardless of the true
+bottleneck, it is clear that choosing the N:1 KV implementation is the better
+choice because of the simple implementation and the minimal performance
+degradation.
+
+\textbf{Takeaway}: choosing the best implementations is dependent on hardware.
+Something as common as an upgrade from HDD to SSD may nullify the benefits of a
+certain implementation.
 

### 2016-05-25 paper.tex
cap update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \end{tabularx}
-\caption{A variety of RADOS object storage classes exist that expose reusable
-interfaces to applications.}
+\caption{A variety of RADOS object storage interfaces exist that expose reusable
+    interfaces to applications. The \emph{locking}, \emph{logging}, and
+    \emph{GC} categories are generic, while high-level services like \emph{RBD} and
+    \emph{RGW} implement application-specific metadata management interfaces.}
 \label{tab:objclass-cats}
@@
 EXIF metadata as key-value pairs in the object key-value database.  However,
-depending on the application needs it may be sufficient or offer a performance
+depending on the application requirements it may be sufficient or offer a performance
 advantage to store this metadata as a header within the bytestream. In the remainder

### 2016-05-25 paper.tex
addresing some of peters feedback
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 eliminating the benefits of commodity systems. But the recent availability of
-high-performance open-source storage systems is changing this because these
-systems are modifiable, enabling interface change, and reducing the risks of
+high-performance open-source storage systems is changing the way people
+approach storage because these
+systems are modifiable, enabling interface change and reducing the risks of
 lock-in. The widely deployed Ceph distributed storage system is an example of
@@
 storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
-reliant upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by
+reliant upon these extensions. Figure~\ref{fig:objclass-dev} shows
 a marked increase in the number of object operations that are packaged as part
@@
 Unfortunately the size of the design space for even simple interfaces can
-be very large, and it can be difficult to choose a design that future proofs an
+be very large, and it can be difficult to choose a design that
+``future-proofs'' an
 implementation against hardware and software upgrades.
@@
 Figure~\ref{fig:ceph} illustrates the collection of components commonly
-referred to as Ceph. At the bottom, a cluster of 10s--10,000s \emph{object
+referred to as Ceph. At the bottom, a cluster of 10s--10,000s of \emph{object
 storage devices} compose the distributed object storage system called RADOS.
@@
 \section{Physical Design Space}
+\label{sec:pd}
 
@@
 As Ceph continues to develop into a target for building distributed services
-and application, it is important to not restrict use to developers seasonsed
-in distributed programming and those with knowledge of intracacies of the Ceph
+and application, it is important to not restrict use to developers seasoned
+in distributed programming and those with knowledge of intricacies of the Ceph
 performance model. To this end we present Brados, a declarative language with
@@
 compositions from details of their physical implementations introduces a
-number of degrees of freedom, similar to the ``data independence’’ offered by
+number of degrees of freedom, similar to the ``data independence'' offered by
 database query languages.  The choice of access methods (for example, whether
@@
 
-  scratch :write_op, [:type, :pos, :epoch] => [:data]
-  scratch :read_op,  [:type, :pos, :epoch] => [:data]
-  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
-  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
-  scratch :seal_op,  [:type, :pos, :epoch] => [:data]
+  scratch :write_op, op.schema
+  scratch :read_op,  op.schema
+  scratch :trim_op,  op.schema
+  scratch :fill_op,  op.schema
+  scratch :seal_op,  op.schema
 
   # op did or did not pass the epoch guard
-  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
-  scratch :invalid_op, [:type, :pos, :epoch] => [:data]
+  scratch :valid_op,   op.schema
+  scratch :invalid_op, op.schema
 
   # op's position was or was not found in the log
-  scratch :found_op,    [:type, :pos, :epoch] => [:data]
-  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
+  scratch :found_op,    op.schema
+  scratch :notfound_op, op.schema
 end
@@
 A valid write operation is then merged into log, otherwise a read only error is
-returend indicating that hte log position was already written to.
+returend indicating that the log position was already written to.
 The fill operation is identical except a the fill state is set on the log entry.
@@
 
+Amazingly these few code snippets express the semantics of the entire storage
+device interface requirements in CORFU. For reference the ZLog implementation
+of the CORFU storage interface in C++ comprises nearly 700 lines of code and
+was designed to store log entries and metadata in the key-value interface,
+thus requiring a lengthy rewrite to realize the performance advantage
+available by using the bytestream interface, as was highlighted in
+Section~\ref{sec:pd}. Furthermore, the complexity introduced by using the
+bytestream interface would grow the amount of code written in the optimized
+version.
+
+But beyond the convience of writing less code, it is far easier for the
+programmer writing an interface such as CORFU to convinence herself of the
+correctness of the high-level details of the implementation without being
+distracted by issues related to physical design or the many other gotchas that
+one must deal with when writing low-level systems software.
+
 \subsection{Optimization}

### 2016-05-25 paper.tex
make sentence shrt
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 the bulk of these interfaces are created by core Ceph developers with a
-complete view of the performance model, this may be changing as the
+complete view of the performance model, this may be changing. The
 development community has been receptive to outside contributions with the

### 2016-05-25 paper.tex
final edit from pter
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 applications such as OpenStack Swift and Cinder~\cite{openstack}.
+Unfortunately, the code composing these interfaces
+is written assuming a performance profile defined by the combination of the
+hardware and software versions available at the time of development.
 
@@
 installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
-low-level C++ written to implement these operations. Unfortunately, this code
-is written assuming a performance profile defined by the combination of the
-hardware and software versions available at the time of development.  While
+low-level C++ written to implement these operations.
+While
 the bulk of these interfaces are created by core Ceph developers with a

### 2016-05-25 paper.tex
paper: cleanup graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-25 paper.tex
@@
 eliminating the benefits of commodity systems. But the recent availability of
-high-performance open-source storage systems is changing the way people
-approach storage because these
-systems are modifiable, enabling interface change and reducing the risks of
+high-performance open-source storage systems is changing this because these
+systems are modifiable, enabling interface change, and reducing the risks of
 lock-in. The widely deployed Ceph distributed storage system is an example of
@@
 storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
-reliant upon these extensions. Figure~\ref{fig:objclass-dev} shows
+reliant upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by
 a marked increase in the number of object operations that are packaged as part
@@
 Unfortunately the size of the design space for even simple interfaces can
-be very large, and it can be difficult to choose a design that
-``future-proofs'' an
+be very large, and it can be difficult to choose a design that future proofs an
 implementation against hardware and software upgrades.
@@
 Figure~\ref{fig:ceph} illustrates the collection of components commonly
-referred to as Ceph. At the bottom, a cluster of 10s--10,000s of \emph{object
+referred to as Ceph. At the bottom, a cluster of 10s--10,000s \emph{object
 storage devices} compose the distributed object storage system called RADOS.
@@
 \end{tabularx}
-\caption{A variety of RADOS object storage interfaces exist that expose reusable
-    interfaces to applications. The \emph{locking}, \emph{logging}, and
-    \emph{GC} categories are generic, while high-level services like \emph{RBD} and
-    \emph{RGW} implement application-specific metadata management interfaces.}
+\caption{A variety of RADOS object storage classes exist that expose reusable
+interfaces to applications.}
 \label{tab:objclass-cats}
@@
 EXIF metadata as key-value pairs in the object key-value database.  However,
-depending on the application requirements it may be sufficient or offer a performance
+depending on the application needs it may be sufficient or offer a performance
 advantage to store this metadata as a header within the bytestream. In the remainder
@@
 \section{Physical Design Space}
-\label{sec:pd}
 
@@
       \centering
-      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
+      \includegraphics[width=0.85\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
       \caption{The expected random read performance from a log with 1KB entries
@@
 	\centering
-	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
+	\includegraphics[width=0.85\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}	
 	\caption{1hr,1k,cls\_wr,jewel: this graph shows the overhead of class}
@@
 As Ceph continues to develop into a target for building distributed services
-and application, it is important to not restrict use to developers seasoned
-in distributed programming and those with knowledge of intricacies of the Ceph
+and application, it is important to not restrict use to developers seasonsed
+in distributed programming and those with knowledge of intracacies of the Ceph
 performance model. To this end we present Brados, a declarative language with
@@
 compositions from details of their physical implementations introduces a
-number of degrees of freedom, similar to the ``data independence'' offered by
+number of degrees of freedom, similar to the ``data independence’’ offered by
 database query languages.  The choice of access methods (for example, whether
@@
 
-  scratch :write_op, op.schema
-  scratch :read_op,  op.schema
-  scratch :trim_op,  op.schema
-  scratch :fill_op,  op.schema
-  scratch :seal_op,  op.schema
+  scratch :write_op, [:type, :pos, :epoch] => [:data]
+  scratch :read_op,  [:type, :pos, :epoch] => [:data]
+  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
+  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
+  scratch :seal_op,  [:type, :pos, :epoch] => [:data]
 
   # op did or did not pass the epoch guard
-  scratch :valid_op,   op.schema
-  scratch :invalid_op, op.schema
+  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
+  scratch :invalid_op, [:type, :pos, :epoch] => [:data]
 
   # op's position was or was not found in the log
-  scratch :found_op,    op.schema
-  scratch :notfound_op, op.schema
+  scratch :found_op,    [:type, :pos, :epoch] => [:data]
+  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
 end
@@
 A valid write operation is then merged into log, otherwise a read only error is
-returend indicating that the log position was already written to.
+returend indicating that hte log position was already written to.
 The fill operation is identical except a the fill state is set on the log entry.
@@
 
-Amazingly these few code snippets express the semantics of the entire storage
-device interface requirements in CORFU. For reference the ZLog implementation
-of the CORFU storage interface in C++ comprises nearly 700 lines of code and
-was designed to store log entries and metadata in the key-value interface,
-thus requiring a lengthy rewrite to realize the performance advantage
-available by using the bytestream interface, as was highlighted in
-Section~\ref{sec:pd}. Furthermore, the complexity introduced by using the
-bytestream interface would grow the amount of code written in the optimized
-version.
-
-But beyond the convience of writing less code, it is far easier for the
-programmer writing an interface such as CORFU to convinence herself of the
-correctness of the high-level details of the implementation without being
-distracted by issues related to physical design or the many other gotchas that
-one must deal with when writing low-level systems software.
-
 \subsection{Optimization}

### 2016-05-25 paper.tex
some edits
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
+\definecolor{mygray}{rgb}{0.5,0.5,0.5}
 \lstset{frame=tb,
   language=Ruby,
-  basicstyle={\scriptsize\sffamily},
+  basicstyle={\scriptsize\ttfamily},
   breaklines=true,
-  captionpos=b
+  captionpos=b,
+  numbers=left,
+  numbersep=5pt,
+  numberstyle=\tiny\color{mygray},
+  xleftmargin=2em,
+frame=l,
+framexleftmargin=1.5em
 }
@@
 %  basicstyle={\scriptsize\ttfamily},
-%  numbers=none,
 %  %numberstyle=\tiny\color{gray},
@@
 
-\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
+\begin{lstlisting}[caption={State Declaration}, label=lst:state]
 state do
@@
 
-  interface input,  :op,   [:type, :pos, :epoch] => [:data]
-  interface output, :ret,  [:type, :pos, :epoch] => [:retval]
+  interface input, :op,
+    [:type, :pos, :epoch] => [:data]
+
+  interface output, :ret,
+    [:type, :pos, :epoch] => [:retval]
 
@@
 end
-
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Setup}, label=lst:setup]
 bloom do
@@
   # demux on operation type
-  write_op <= valid_op { |o| o if o.type == 'write' }
-  read_op  <= valid_op { |o| o if o.type == 'read'  }
-  fill_op  <= valid_op { |o| o if o.type == 'fill'  }
-  trim_op  <= valid_op { |o| o if o.type == 'trim'  }
-  seal_op  <= valid_op { |o| o if o.type == 'seal'  }
+  write_op <= valid_op {|o| o if o.type == 'write'}
+  read_op  <= valid_op {|o| o if o.type == 'read'}
+  fill_op  <= valid_op {|o| o if o.type == 'fill'}
+  trim_op  <= valid_op {|o| o if o.type == 'trim'}
+  seal_op  <= valid_op {|o| o if o.type == 'seal'}
 end
@@
 bloom :epoch_guard do
-  invalid_op <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
+  invalid_op <= (op * epoch).pairs{|o,e|
+    o.epoch <= e.epoch}
   valid_op <= op.notin(invalid_op)
-  ret <= invalid_op{|o| [o.type, o.pos, o.epoch, 'stale']}
+  ret <= invalid_op{|o|
+    [o.type, o.pos, o.epoch, 'stale']}
 end
@@
   log <+- trim_op{|o| [o.pos, 'trimmed']}
-  ret <=  trim_op{|o| [o.type, o.pos, o.epoch, 'ok']}
+  ret <=  trim_op{|o|
+    [o.type, o.pos, o.epoch, 'ok']}
 end
@@
   temp :valid_write <= write_op.notin(found_op)
-  log <+ valid_write{|o| [o.pos, 'valid', o.data]}
-  ret <= valid_write{|o| [o.type, o.pos, o.epoch, 'ok']}
-  ret <= write_op.notin(valid_write) { |o|
-    [o.type, o.pos, o.epoch, 'read-only']
-  }
+  log <+ valid_write{ |o| [o.pos, 'valid', o.data]}
+  ret <= valid_write{ |o|
+    [o.type, o.pos, o.epoch, 'ok'] }
+  ret <= write_op.notin(valid_write) {|o|
+    [o.type, o.pos, o.epoch, 'read-only'] }
 end
@@
   log <+ valid_fill { |o| [o.pos, 'fill'] }
-  ret <= valid_fill { |o| [o.type, o.pos, o.epoch, 'ok'] }
+  ret <= valid_fill { |o|
+    [o.type, o.pos, o.epoch, 'ok'] }
   ret <= fill_op.notin(valid_fill) { |o|
-    [o.type, o.pos, o.epoch, 'read-only']
-  }
+    [o.type, o.pos, o.epoch, 'read-only'] }
 end
@@
 bloom :read do
-  temp :valid_read <= (read_op * log).pairs(pos => pos) { |o, l|
+  temp :ok_read <= (read_op * log).pairs(pos => pos) { |o, l|
     [o.type, o.pos, o.epoch, l.data] unless
-      ['filled', 'trimmed'].include?(l.state)
-  }
-  ret <= valid_read { |e| [e.type, e.pos, e.epoch, e.data] }
-  ret <= read_op.notin(valid_read, type => type) do |o|
+      ['filled', 'trimmed'].include?(l.state) }
+  ret <= ok_read { |e|
+    [e.type, e.pos, e.epoch, e.data] }
+  ret <= read_op.notin(ok_read, type=>type) do |o|
     [o.type, o.pos, o.epoch, 'invalid']
@@
 express the semantics of the CORFU storage interface. We briefly sketch the
-implementation of an additional interface found in Ceph and used by the
-S3-compatible RADOS-Gateway service.
-
-Channel
-C++ code comparison
-
-The subset of the Bloom language we are using in Brados is powerful enough to
-express the CORFU language, as well as many of the object class interfaces that
-are already included in Ceph. We briefly sketch the Bloom imlementation of an
-object class that provides a locking service. This is an interesting example
-because Bloom implementations use mechanisms for introducing points of
-coordination in execution, which in the context of Ceph, may be written in
-Bloom itself.
-
-\begin{lstlisting}[caption={Hello}, label=lab]
-class RefCount
-  include Bud
-
+implementation of an additional interface found in Ceph that used by the
+S3-compatible RADOS-Gateway service to perform object reference counting.
+
+The reference count implementation that we describe does not use a typical
+counter, but rather uses a tag-based interface in which each \emph{reference}
+to an object is identified by a application-level tag. Each tag represents
+one or more references allowing an idempotent property useful for application
+that may replay reference count operations out of order. Otherwise the
+semantics are straightforward: clients record references to an object, and
+when the number of references drop to zero the object is removed.
+
+The reference implementation in Ceph is 300 LOC written in C++. The version
+written in Brados is shown in Listing~\ref{lst:rc}.
+
+ Unlike the CORFU interface
+that took advantage of a mapping between backend storage and collections, the
+reference counting interface requires access to RADOS runtime facilities that
+allow objects to be removed. This is modeled in Brados using the Bloom
+\emph{channel} abstraction that is used to transmit facts between entities in
+the system; we effectively send a message to the runtime with our removal
+request.
+
+\begin{lstlisting}[caption={Reference counting interface}, label=lst:rc]
 state do
@@
   remove_chn <~ remove
-  end
 end
@@
 
-Include performance management topic in future work.
+Much work remains to be done.  While we demonstrated some of the immediate
+benefits of separating logical and physical specifications of component
+compositions, the ``dream’’ of data independence is to take the physical
+design question completely off the table for programmers.  We are currently
+exploring both cost-based optimizers and autotuners to automate the selection
+of a high performance physical design consistent with a given logical
+specification.
+
+The database literature on materialized view maintenance and selection can be
+brought to bear in the context of programmable storage as well.  Recall the
+sequencer logic of Corfu described in Section XX.  The Bloom specification
+(Figure YY) simply states that the current value of the sequencer is the
+result of evaluating an aggregate query over the log relation.  This means
+that the system may choose any strategy to maintain a sequence value that is
+consistent with this constraint.  The physical design decision to store and
+update a memory-resident sequence value rather re-computing the maximal value
+at each request is a view materialization decision that could similarly be
+made by a compile-time optimizer.
+
+The future of programmable storage is likely to witness a further
+decomposition of services to allow greater flexibility in customizing
+deployments.  More than likely, not all of these services will assume or
+provide strongly-consistent operations by default.  Instead, programmers will
+have the flexibility (and burden!) of trading off between performance (offered
+by weakly consistent subsystems) and programmability.  Languages such as
+Bloom---which provide static analyses to determine when loosely-ordered
+operations nevertheless produce consistent outcomes---will become increasingly
+relevant in this domain.
+
+%Include performance management topic in future work.
 

### 2016-05-25 paper.tex

--- 2016-05-25 paper.tex
@@
 Ceph is designed to run on a wide variety of commodity hardware as well as new
-NVMe devices, all of which their own set of characteristics and tunables (e.g.,
-the IO operation scheduler type). In our experiments, we tested SSD, HDDs, NVMe
-devices and discovered a wide range of behaviors and performance profiles.
-Figure~\ref{fig:jewel-hdd-128b}, which shows the write performance of 128 by
-log entries using Jewel and a single HDD, shows an especially tricky situation
-for a developer.
-
-In that figure, the performance is 10\(\times\) slower than its SSD counterpart
-but performance of the N:1 bystestream implementation is almost identical to
+NVMe devices. All these devices have their own set of characteristics and
+tunables (e.g., the IO operation scheduler type). In our experiments, we tested
+SSD, HDDs, NVMe devices and discovered a wide range of behaviors and
+performance profiles. As an example, Figure~\ref{fig:jewel-hdd-128b} shows the
+write performance of 128 byte log entries using Jewel and a single HDD.
+Performance is 10\(\times\) slower than its SSD counterpart in
+Figure~\ref{ig:jewel_v_firefly_v_es} (top row, third column) but the behavior
+and relative performance make this hardware configuration especially tricky.
+
+The behavior of the 1:1 implementations shows throughput drops lasting for
+minutes at a time -- this limits our focus to the N:1 implementations.
+
+ The performance of the N:1 bystestream implementation is almost identical to
 the N:1 KV implementation (within 1\% mean throughput). Regardless of the true

### 2016-05-25 paper.tex
paper: sec3 eloquence
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2016-05-25 paper.tex
@@
 
-\textbf{Takeway}: choosing the best implementations is dependent on both the
+\textbf{Takeaway}: choosing the best implementations is dependent on both the
 timing of the developemnt (Ceph Version) and the expertise of the administrator
@@
 The behavior of the 1:1 implementations shows throughput drops lasting for
-minutes at a time -- this limits our focus to the N:1 implementations.
-
- The performance of the N:1 bystestream implementation is almost identical to
-the N:1 KV implementation (within 1\% mean throughput). Regardless of the true
+minutes at a time -- this limits our focus to the N:1 implementations. The
+performance of the N:1 bystestream implementation is almost identical to the
+N:1 KV implementation (within 1\% mean throughput). Regardless of the true
 bottleneck, it is clear that choosing the N:1 KV implementation is the better
-choice because of the simple implementation and the minimal performance
-degradation.
+choice because of the resulting implementation should be less complex and there
+is minimal performance degradation.
 
@@
 
+What started out as a pragmatic approach to designing a complex service on Ceph
+quickly proved to be unwieldy. While finding the best solution for a specific
+set of configurations (i.e. software, tunable, and hardware parameters) can be
+done, finding a solution that is future-proof is still unsolved.
+
 This physical design process, of first thinking about the design trade-offs and
-then doing parameters sweeps, is insufficient when general-purpose systems. Had
-we built ZLog from the ground up (like CORFU) the story might be different but
-to deal wtih the large state space we need a way to automatically and
-autonomically decide which parameters to choose.
+then doing parameters sweeps, is insufficient for large general-purpose systems
+like Ceph. Had we built ZLog from the ground up (like CORFU) the story might be
+different but to deal with the large state space we need a way to automatically
+and autonomically decide which parameters to choose.
 

### 2016-05-25 paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 As Ceph continues to develop into a target for building distributed services
-and application, it is important to not restrict use to developers seasonsed
-in distributed programming and those with knowledge of intracacies of the Ceph
-performance model. To this end we present Brados, a declarative language with
+and applications, it is important to not restrict use to developers seasonsed
+in distributed programming and those with knowledge of the intracacies of Ceph
+and its performance model. To this end we present Brados, a declarative language with
 relational semantics for defining the behavior of RADOS object interfaces.
-Brados uses a simple declarative data model and can reduce the learning curve
+Brados uses a simple declarative data model, can reduce the learning curve
 for new users, and allows existing developers to increase productivity by
@@
 database query languages.  The choice of access methods (for example, whether
-to use a bytestream interface or a key/value store), storage decide
-classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
-1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
+to use a bytestream interface or a key/value store), storage device
+classes (e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
+1:1 or N:1 mapping) and execution plans (e.g. operator ordering) can be
 postponed to execution time.  The optimal choices for these physical design
@@
 \subsection{The CORFU Log Interface}
+
+We now show the implementation of CORFU using Brados and use this as a vehicle
+for describing additional aspects of the language of semantics.
+Listing~\ref{lst:state} shows the declaration of state for the CORFU
+interface. Lines 1 and 2 define the schema of the two persistent collections
+that hold the current epoch value, and the log contents. These collections are
+mapped onto storage within Ceph but abstract away the low-level interface
+(e.g. bytestream vs key-value). Lines 5-9 define the input and output
+interfaces. We use a generic schema for the input operation to simplify
+how rules are defined that apply to all operation types.
 

### 2016-05-25 paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 interfaces. We use a generic schema for the input operation to simplify
-how rules are defined that apply to all operation types.
+how rules are defined that apply to all operation types. Lines 11-15 define
+named collections for each operation type. The scratch type indicates that the
+data is not persistent, and only remains in the collection for a single 
+execution time step. The remaining scratch collections are defined to further
+subdivide the operations based on different properties which we'll describe next.
 
@@
 
-  scratch :write_op, [:type, :pos, :epoch] => [:data]
-  scratch :read_op,  [:type, :pos, :epoch] => [:data]
-  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
-  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
-  scratch :seal_op,  [:type, :pos, :epoch] => [:data]
+  scratch :write_op, op.schema
+  scratch :read_op,  op.schema
+  scratch :trim_op,  op.schema
+  scratch :fill_op,  op.schema
+  scratch :seal_op,  op.schema
 
   # op did or did not pass the epoch guard
-  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
-  scratch :invalid_op, [:type, :pos, :epoch] => [:data]
+  scratch :valid_op,   op.schema
+  scratch :invalid_op, op.schema
 
   # op's position was or was not found in the log
-  scratch :found_op,    [:type, :pos, :epoch] => [:data]
-  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
+  scratch :found_op,    op.schema
+  scratch :notfound_op, op.schema
 end
 \end{lstlisting}
+
+Initialization is performed in Listing~\ref{lst:setup} which acts as a demux
+for the operation type and properties.  Lines 3-7 show the epoch guard that is
+applied to all operations.  The guard rejects requests that are tagged with old
+epoch values, ensuring that a client generating a request has an up-to-date
+view of the system.  First the \texttt{invalid\_op} collection is defined to
+include the current operation if its epoch value is no larger than the stored
+epoch value. Next the \texttt{valid\_op} collection is defined to be the
+inverse of \texttt{invalid\_op} and is a helper used to refine other operations
+later in the dataflow. Finally we handle the case for all operations tagged
+with an out-of-date epoch by merging the \texttt{invalid\_op} set into the
+output \texttt{ret} collection.
+
+Lines 10 and 11 populate the \texttt{found\_op} and \texttt{notfound\_op}
+collections that allow operations behavior to be predicated on if the position
+associated with a request is found in the log. Finally the remaining lines in
+Listing~\ref{lst:setup} populate each of the specific operation collections.
 
@@
 bloom do
+  # epoch guard
+  invalid_op <= (op * epoch).pairs{|o,e|
+    o.epoch <= e.epoch}
+  valid_op <= op.notin(invalid_op)
+  ret <= invalid_op{|o|
+    [o.type, o.pos, o.epoch, 'stale']}
+
   # op's position found in log
@@
 
-Listing~\ref{lst:eg} shows the epoch guard that is applied to all operations.
-The guard rejects requests that are tagged with old epoch values, ensuring that
-a client generating a request has an up-to-date view of the system.  First the
-\texttt{invalid\_op} collection is defined to include the current operation if
-its epoch value is no larger than the stored epoch value. Next the
-\texttt{valid\_op} collection is defined to be the inverse of
-\texttt{invalid\_op} and is a helper used to refine other operations later in
-the dataflow. Finally we handle the case for all operations tagged with an
-out-of-date epoch by merging the \texttt{invalid\_op} set into the output
-\texttt{ret} collection.
-
-\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
-bloom :epoch_guard do
-  invalid_op <= (op * epoch).pairs{|o,e|
-    o.epoch <= e.epoch}
-  valid_op <= op.notin(invalid_op)
-  ret <= invalid_op{|o|
-    [o.type, o.pos, o.epoch, 'stale']}
-end
-\end{lstlisting}
-
 The process of sealing an object requires installing a new epoch value and
@@
 implements the seal interface by first removing the current epoch value and
-replacing it with the epoch value contained in the operation. Next an aggregate
-is computed over the log to find the maximum position written, and this value
-is returned.
+replacing it with the epoch value contained in the input operation. Next an
+aggregate is computed over the log to find the maximum position written, and
+this value is returned, typically to a client performing a reconfiguration of
+the system or following the failure of a sequencer.
 
@@
 
-Trimming a log entry always succeeds. In Listing~\ref{lst:trim} The
+Trimming a log entry always succeeds. In Listing~\ref{lst:trim} the
 \texttt{<+-} operator simulatenously removes the log entry with the given 
 position and replaces it with an entry with its state set to \emph{trimmed}.
+In practice the removal of a log entry may trigger garbage collection, but
+we model it here as an update for brevity.
 
@@
 
-The write and fill interfaces are implemented similarily, and are show in
-Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
-write collection is created if the operation position is not found in the log.
-A valid write operation is then merged into log, otherwise a read only error is
-returend indicating that hte log position was already written to.
-The fill operation is identical except a the fill state is set on the log entry.
-
-\begin{lstlisting}[caption={Write}, label=lst:write]
+The write and fill interfaces are implemented similarily, and are both shown in
+Listing~\ref{lst:write}.  A \texttt{valid\_write} collection is created if the
+operation position is not found in the log.  A \texttt{valid\_write} operation
+is then merged into log, otherwise a read only error is returend indicating
+that the log position was already written to.  The fill operation is identical
+except a the fill state is set on the log entry.
+
+\begin{lstlisting}[caption={Write and Fill}, label=lst:write]
 bloom :write do
@@
 end
-\end{lstlisting}
-
-\begin{lstlisting}[caption={Fill}, label=lst:fill]
+
 bloom :fill do
@@
 
-The read interface is shown in Listing~\ref{lst:read}, and structured similar
-to the write interface. First we create a collection containing a valid read
-operation that is in the log and does not have the filled or trimmed state set.
-The data is returend in the case of a valid read op, otherwise an error is
+Finally the read interface is shown in Listing~\ref{lst:read}, and structured
+in a similar way to the write and fill interfaces. First we create a collection
+containing a valid read operation (named \texttt{ok\_read}) that is in the log
+and does not have the filled or trimmed state set.  The data read from the log
+is returend in the case of a valid read operation, otherwise an error is
 returend through the output interface.
@@
 \end{lstlisting}
+
+Amazingly these few code snippets express the semantics of the entire storage
+device interface requirements in CORFU. For reference the ZLog implementation
+of the CORFU storage interface in C++ comprises nearly 700 lines of code and
+was designed to store log entries and metadata in the key-value interface,
+thus requiring a lengthy rewrite to realize the performance advantage
+available by using the bytestream interface, as was highlighted in
+Section~\ref{sec:pd}. Furthermore, the complexity introduced by using the
+bytestream interface would grow the amount of code written in the optimized
+version.
+
+But beyond the convience of writing less code, it is far easier for the
+programmer writing an interface such as CORFU to convinence herself of the
+correctness of the high-level details of the implementation without being
+distracted by issues related to physical design or the many other gotchas that
+one must deal with when writing low-level systems software.
 

### 2016-05-25 paper.tex

@@
 We conclude by discussing the optimization opportunities that we can realize
-using analysis of interfaces. We hope you enjoy the show.
+using analysis of interfaces.
 
@@
 
-\paragraph*{CORFU and Storage Programmability}
 The second property of CORFU relevant in the context of Ceph is the dependency
@@
 
+\paragraph*{Storage Programmability}
 While the authors of the CORFU paper describe prototype device interfaces
@@
 the creation of logical storage devices through its object class feature
-described previously in Section~\ref{sec:objclass}. Thus, by using
+described in Section~\ref{sec:objclass}. Thus, by using
 software-based object interfaces offered by RADOS flash devices in CORFU can be
-replaced by software-defined storage offering significant flexibility and a
+replaced by software-defined interfaces offering significant flexibility and a
 simplified design.
@@
 
-As we have seen, Ceph provides a rich storage API and places few restrictions
+As we have seen, Ceph provides a rich storage API and places few requirements
 on the structure of applications. Thus a primary concern when implementing an
@@
   \node[state]         (E) [right of=C]       {$US$};
-  \node[state]         (F) [below right of=E] {$Fi$};
+  \node[state]         (F) [below right of=E] {$Out$};
 
@@
 
-We begin the process of exploring the design space by exploiting the fact that
-metadata management is a complexity and performance overhead that a full
-implementation must incur beyond the costs of storing log entry data.
-Therefore we first explore the design space by restricting the
-space to log entry addressing and log entry storage (the \emph{I/O} action
-shown in Figure~\ref{fig:corfu-sm}).  These two dimensions are
-represented by the first two columns of Table~\ref{tab:pd-map} which describes
-the entire design space. In the I/O column \emph{KV} corresponds to the
-key-value interface, and the bytestream interface is represented by \emph{AP}
-for an append strategy, and \emph{EX} for a strategy that writes to an
-explicit bytestream offset (both described shortly).
+We begin the process of exploring the design space by focusing on log entry
+addressing and log entry storage (the \emph{I/O} action shown in
+Figure~\ref{fig:corfu-sm}).  These two dimensions are represented by the first
+two columns of Table~\ref{tab:pd-map} which describes the entire design space.
+In the I/O column \emph{KV} corresponds to the key-value interface, and the
+bytestream interface is represented by \emph{AP} for an append strategy, and
+\emph{EX} for a strategy that writes to an explicit bytestream offset (both
+described shortly).
 
@@
 relatively poor performance. Jewel with 1KB entry sizes sees a 2.5\(\times\)
-performance increase for the N:1 KV mapping. The large period of reduced
+performance increase for the \texttt{(N:1,KV)} mapping. The large period of reduced
 throughput for corresponds to the OSD splitting file system directories in
-order to maintain a maximum directory size, and will occur in \emph{1:1, KV}
+order to maintain a maximum directory size, and will occur in \texttt{(1:1, KV)}
 although the threshold number of objects is not reached in this example due to
-the reduced overall throughput of the \emph{1:1, KV} strategy.
+the reduced overall throughput of the \texttt{(1:1, KV)} strategy.
 
@@
       \caption{The expected random read performance from a log with 1KB entries
-      performs best on the same strategy that the write performance performs
-      best on - the N:1 bystream append strategy.}
+      performs best on the same strategy that the writes perform
+      best on - the \texttt{(N:1, BS)} append strategy.}
       \label{fig:vanilla_rd_jewel}
@@
 Performance is 10\(\times\) slower than its SSD counterpart in
-Figure~\ref{ig:jewel_v_firefly_v_es} (top row, third column) but the behavior
+Figure~\ref{fig:jewel_v_firefly_v_es} (top row, third column) but the behavior
 and relative performance make this hardware configuration especially tricky.
@@
 minutes at a time -- this limits our focus to the N:1 implementations. The
-performance of the N:1 bystestream implementation is almost identical to the
-N:1 KV implementation (within 1\% mean throughput). Regardless of the true
-bottleneck, it is clear that choosing the N:1 KV implementation is the better
-choice because of the resulting implementation should be less complex and there
-is minimal performance degradation.
+performance of \texttt{(N:1, BS)} implementation is almost identical to
+\texttt{(N:1, KV)} (within 1\% mean throughput). Regardless of the true
+bottleneck, it is clear that choosing \texttt{(N:1, KV)} is the better choice
+because of the resulting implementation should be less complex and there is
+minimal performance degradation.
 

### 2016-05-25 paper.tex
importnat
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 counter, but rather uses a tag-based interface in which each \emph{reference}
-to an object is identified by a application-level tag. Each tag represents
-one or more references allowing an idempotent property useful for application
+to an object is identified by a application-level tag. Each tag represents one
+or more references supporting an idempotent property useful for applications
 that may replay reference count operations out of order. Otherwise the
-semantics are straightforward: clients record references to an object, and
-when the number of references drop to zero the object is removed.
+semantics are straightforward: clients record references to an object by
+inserting a tag, and remove a reference by removing a tag. When the number of
+references (i.e., tags) drop to zero the object is garbage collected.
 
 The reference implementation in Ceph is 300 LOC written in C++. The version
-written in Brados is shown in Listing~\ref{lst:rc}.
-
- Unlike the CORFU interface
-that took advantage of a mapping between backend storage and collections, the
+written in Brados is shown in Listing~\ref{lst:rc}. Line 10 shows the
+implementation of requesting a reference which is an idempotent merge of the
+tag into the persistent \texttt{refs} collection.  Unlike the CORFU interface
+that took highlighted how collections are unified with persistent storage, the
 reference counting interface requires access to RADOS runtime facilities that
@@
 the system; we effectively send a message to the runtime with our removal
-request.
+request. The \texttt{remove\_chn} channel may be populated with messages and
+the RADOS runtime will processes these asynchronously.
 

### 2016-05-25 paper.tex

@@
         clients to locate a log entry given its position.  In contrast, an N:1
-        strategy \emph{stripes} log entires across a smaller set of objects,
+        strategy \emph{stripes} log entries across a smaller set of objects,
         but this adds complexity to both the client and the object interface which
@@
 designs is shown in Figure~\ref{fig:cls_wr_jewel} where we begin by
-highlighting two baseline throughputs labeled \emph{librados} and
-\emph{cls\_null} that correspond to the baseline append throughput described
+highlighting two baseline throughputs labelled \texttt{librados} and
+\texttt{write only} that correspond to the baseline append throughput described
 in the previous section, and the append throughput achieved with \emph{object
@@
 value is stored under a fixed key, and the metadata associated with each log
-entry is stored under a key derived from the log position. The expected throughput of
-this design is shown in Figure~\ref{fig:cls_wr_jewel} and labeled as
-\emph{cls\_full\_omap}. This result shows us that even for very small values
-the cost of managing data in the key-value store introduces significant
-overhead. In contrast, the unstructured bytestream interface imposes little
-to no restriction on how it is used.
+entry is stored under a key derived from the log position. The expected
+throughput of this design is shown in Figure~\ref{fig:cls_wr_jewel} and labeled
+as \texttt{KV}. This result shows us that even for very small values the cost
+of managing data in the key-value store introduces significant overhead. In
+contrast, the unstructured bytestream interface imposes little to no
+restriction on how it is used.
 
@@
 write the updated per-entry metadata. The expected throughput from such a
-design is shown in Figure~\ref{fig:cls_wr_jewel} labeled as
-\emph{cls\_full\_hdr} and performs with roughly a 6\% overhead.
+design is shown in Figure~\ref{fig:cls_wr_jewel} labeled as \texttt{BS} for
+bytestream and performs with roughly a 6\% overhead.
 
@@
 
-In this section we have demonstrated that a design based on an N:1 addressing
-scheme that stores both log entries and metadata in the bytestream can provide
-the best overall performance, and it does so by a large margin. However, this
+So far, we have demonstrated that a design based on an N:1 addressing scheme
+that stores both log entries and metadata in the bytestream can provide the
+best overall performance, and it does so by a large margin. However, this
 process of design we have outlined may not yield such clear cut results when
 applied in a different context. To illustrate these contexts, we describe the
-software, system tunable, and hardware parameters that affect the performance
-of our CORFU implementation. The complexity and breadth of our parameter sweep,
-and the fact that our state space grew so large, motivates brados in section 4.
+software parameters, system tunables, and hardware parameters that affect the
+performance of our CORFU implementation. The complexity and breadth of our
+parameter sweep, and the fact that our state space grew so large, motivates the
+language for declarative, programmable object storage described in section 4.
 
@@
 
-Ceph releases stable versions every year (Oct/Nov) and LTS versions every 3-4
-months~\cite{website:ceph-releases}. The head of the master branch moves
-quickly because there are over 400 contributors and an active mailing list.
-Over the past calendar year, there were between 70 and 260 commits per
-week~\cite{website:ceph-github}.
+Ceph releases stable versions every year (Oct/Nov) and long-term support (LTS)
+versions every 3-4 months~\cite{website:ceph-releases}. The head of the master
+branch moves quickly because there are over 400 contributors and an active
+mailing list.  Over the past calendar year, there were between 70 and 260
+commits per week~\cite{website:ceph-github}.
 
@@
 technique that we use, it is clear the the number of tunables increases the
-physical design parameters to an unwieldy stae space size.
-
-\textbf{Takeaway}: the number and complexity of Ceph's tunables make
-brute-force parameter selection impossible.
+physical design parameters to an unwieldy state space size.
+
+\textbf{Takeaway}: the number and complexity of Ceph's tunables makes
+brute-force parameter selection hard.
 

### 2016-05-25 paper.tex
claros update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 in distributed programming and those with knowledge of the intracacies of Ceph
-and its performance model. To this end we present Brados, a declarative language with
-relational semantics for defining the behavior of RADOS object interfaces.
-Brados uses a simple declarative data model, can reduce the learning curve
-for new users, and allows existing developers to increase productivity by
-writing less code that is more portable.
-
-The Brados language corresponds to a subset of the Bloom language, and models the
-storage system state uniformly as a collection of
+and its performance model. The hard-coded imperative method of writing object
+interfaces today restrict the optimizations that can be performed
+automatically or derived from static analysis.  To this end we present Brados,
+a declarative language with relational semantics for defining the behavior of
+RADOS object interfaces.  Brados uses a declarative data model, can reduce the
+learning curve for new users, and allows existing developers to increase
+productivity by writing less code that is more portable.
+
+The Brados language corresponds to a subset of the Bloom language which a
+declarative language for expressing distributed programs as an unordered set
+of rules. These rules fully specify program semantics and allow a programmer
+to ignore the details associated with how a program is evaluated. This level
+of abstraction is attractive for building storage interfaces whose portability
+and correctness is critical.
+
+\subsection{Basics}
+
+Brados models the storage system state uniformly as a collection of
 relations~\cite{alvaro:cidr11}. The composition of a collection of existing
@@
 database query languages.  The choice of access methods (for example, whether
-to use a bytestream interface or a key/value store), storage device
-classes (e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
-1:1 or N:1 mapping) and execution plans (e.g. operator ordering) can be
-postponed to execution time.  The optimal choices for these physical design
-details are likely to change much more often than the logical specification,
-freeing the interface designer from the need to rewrite systems and device and
-interface characteristics change.
-
-\subsection{Basics}
+to use a bytestream interface or a key/value store), storage device classes
+(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a 1:1 or
+N:1 mapping) and execution plans (e.g. operator ordering) can be postponed to
+execution time.  The optimal choices for these physical design details are
+likely to change much more often than the logical specification, freeing the
+interface designer from the need to rewrite systems and device and interface
+characteristics change.
 

### 2016-05-25 paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 declarative language for expressing distributed programs as an unordered set
-of rules. These rules fully specify program semantics and allow a programmer
+of rules~\cite{alvaro:cidr11}. These rules fully specify program semantics and allow a programmer
 to ignore the details associated with how a program is evaluated. This level
@@
 Brados models the storage system state uniformly as a collection of
-relations~\cite{alvaro:cidr11}. The composition of a collection of existing
+relations. The composition of a collection of existing
 interfaces is then expressed as a collection of high-level \emph{queries} that
@@
 Amazingly these few code snippets express the semantics of the entire storage
-device interface requirements in CORFU. For reference the ZLog implementation
-of the CORFU storage interface in C++ comprises nearly 700 lines of code and
+device interface requirements in CORFU. For reference our prototype
+implementation of CORFU in Ceph (called
+ZLog\footnote{https://github.com/noahdesu/zlog})
+is written in C++ and the storage interface component comprises nearly 700
+lines of code. This version
 was designed to store log entries and metadata in the key-value interface,

### 2016-05-25 paper.tex

@@
 \caption{The high-level design space of mapping CORFU log entry storage onto
-the RADOS object storage system.}
+the RADOS object storage system. }
 \label{tab:pd-map}
@@
 	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel_v_firefly_v_es.png}
-	\caption{The cluster-specific optimizations impact performance... a lot. This caption is quite terrible.
+	\caption{The relative performance difference between the different mapping strategies in Table~\ref{tab:pd-map} changes with the Ceph version.
+These discrepancies alter the implementation that a developer \emph{would have chosen} had they been developing on a different release.
+Here we vary the log entry size to show that there is a trend for Jewel (larger entry sizes decreases \texttt{(N:1)}'s relative performance) that is not present in Firefly.
 	\label{fig:jewel_v_firefly_v_es}}
@@
 throughput for corresponds to the OSD splitting file system directories in
-order to maintain a maximum directory size, and will occur in \texttt{(1:1, KV)}
+order to maintain a maximum directory size, and will occur in \texttt{(1:1,KV)}
 although the threshold number of objects is not reached in this example due to
-the reduced overall throughput of the \texttt{(1:1, KV)} strategy.
+the reduced overall throughput of the \texttt{(1:1,KV)} strategy.
 
@@
       \includegraphics[width=0.85\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
-      \caption{The expected random read performance from a log with 1KB entries
+      \caption{The expected random read performance for a log with 1KB entries
       performs best on the same strategy that the writes perform
-      best on - the \texttt{(N:1, BS)} append strategy.}
+      best on - \texttt{(N:1, BS)}.}
       \label{fig:vanilla_rd_jewel}
@@
 	\includegraphics[width=0.85\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}	
-	\caption{1hr,1k,cls\_wr,jewel: this graph shows the overhead of class}
+	\caption{The overhead of Ceph's object class interfaces and the cost of checking metadata affect the performance of the CORFU API on RADOS.
+The performance degradation of \texttt{KV} might not be enough to offset the simplicity of the implementation.}
 	\label{fig:cls_wr_jewel}
@@
 	\includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/jewel-hdd-128b.png}
-	\caption{shows that with a HDD, it don't matter if you choose omap!!!!!}
+	\caption{The performance of \texttt{(N:1,KV)} is within 1\% of \texttt{(N:1,BS)} when using the Ceph Jewel release on a HDD.
+For this specific hardware/software configuration, \texttt{(N:1,KV)} is a better implementation.}
 	\label{fig:jewel-hdd-128b}
@@
 then doing parameters sweeps, is insufficient for large general-purpose systems
-like Ceph. Had we built ZLog from the ground up (like CORFU) the story might be
+like Ceph. Had we built CORFU from the ground up the story might be
 different but to deal with the large state space we need a way to automatically

### 2016-05-25 paper.tex
asdf
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 \subsection{Motivating Application: CORFU}
+\label{sec:corfu}
 
@@
 But beyond the convience of writing less code, it is far easier for the
-programmer writing an interface such as CORFU to convinence herself of the
+programmer writing an interface such as CORFU to convenience herself of the
 correctness of the high-level details of the implementation without being
@@
 
-\subsection{Optimization}
-
-Choosing implementation details
-  - the refcount example is good. small data kept in ceph object bytestream
-until it becomes large then use omap.
-
-  - could use the dataflow graph to do interesting things like understand how
-  to do prefecting of things like keys or data that are often only known at
-  execution time
-
 \subsection{Other Interfaces}
@@
 express the semantics of the CORFU storage interface. We briefly sketch the
-implementation of an additional interface found in Ceph that used by the
+implementation of an additional interface found in Ceph that is used by the
 S3-compatible RADOS-Gateway service to perform object reference counting.
@@
 tag into the persistent \texttt{refs} collection.  Unlike the CORFU interface
-that took highlighted how collections are unified with persistent storage, the
+that highlighted how collections are unified with persistent storage, the
 reference counting interface requires access to RADOS runtime facilities that
@@
 
-\section{Evaluation}
+\subsection{Optimizations}
+
+The challenge of navigating the physical design space has served as the
+primary source of motivation for selection of a declarative langauge. While
+our implementation does not yet
+
+The Bloom language that we use as a basis for Brados produces a data flow
+graph that can be used in static analysis. We envision that this graph will be
+made available to the OSD and used to reorder and collesce requests based on
+optimization criteria available from a performance model combined with
+semantic information from the dataflow. For example today object classes are
+represented as black boxes from the point of view of the OSD execution engine.
+Understanding the behavior of an object class may allow intelligent
+prefetching. Another type of analysis that may be useful for optimization is
+optimistic execution combined with branch prediction where frequent paths
+through a dataflow are handled optimistically.
 
@@
 {\bf Active storage.}
-There is a wide variety of work related \emph{active storage} that seeks to
+There is a wide variety of work related to \emph{active storage} that seeks to
 increase performance and reduce data movement by exploiting remote storage
@@
 benefits of separating logical and physical specifications of component
-compositions, the ``dream’’ of data independence is to take the physical
+compositions, the ``dream'' of data independence is to take the physical
 design question completely off the table for programmers.  We are currently
@@
 brought to bear in the context of programmable storage as well.  Recall the
-sequencer logic of Corfu described in Section XX.  The Bloom specification
-(Figure YY) simply states that the current value of the sequencer is the
+sequencer logic of Corfu described in Section~\ref{sec:corfu}.  The Bloom specification
+(Listing~\ref{lst:write}) simply states that the current value of the sequencer is the
 result of evaluating an aggregate query over the log relation.  This means

### 2016-05-25 paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

--- 2016-05-25 paper.tex
@@
 primarily concerned with fixed interfaces, application-specific interfaces
-built using active storage techniques have not to the best of our knoweldge
+built using active storage techniques have not to the best of our knowledge
 addressed portability concerns that arise from defining interfaces
@@
 can result in poor performance compared to other designs. In the remainder of
-this section we will define the entire design space and use a set of targetted
+this section we will define the entire design space and use a set of targeted
 benchmarks to arrive at a final design. Finally we will show that these design
@@
 The fact that the appends per second are an order of magnitude slower on
-Firefly indicates that the the benchmark is bottlenecked by the Ceph software
+Firefly indicates that the benchmark is bottlenecked by the Ceph software
 instead of the external log implementation.
@@
 this choice had been made, then a future system upgrade could drop a
-significant number of iops on the floor.
+significant number of IOPS on the floor.
 
@@
 software defined storage strategy of Ceph because it gives the administrator
-the flexiblity to store the 3 componenets of BlueStore (e.g., data, RocksDB
+the flexibility to store the 3 components of BlueStore (e.g., data, RocksDB
 database, and RocksDB write-ahead log) on any partition on any device in the
@@
 \textbf{Takeaway}: choosing the best implementations is dependent on both the
-timing of the developemnt (Ceph Version) and the expertise of the administrator
+timing of the development (Ceph Version) and the expertise of the administrator
 (Ceph Features). Different versions and features of Ceph may lead the
@@
 and 95 of them focus on the OSD back end file system (i.e. its
-\texttt{filestore}). Ceph also has tunables for the subystems it uses, like
+\texttt{filestore}). Ceph also has tunables for the subsystems it uses, like
 LevelDB (10 tunables), RocksDB (5 tunables), its own key-value stores (5
 tunables), its object cache (6 tunables), its journals (24 tunables), and its
-other optional filestores like BlueStore (49 tunables).
+other optional object stores like BlueStore (49 tunables).
 
@@
 in~\cite{behzad:sc2013-autotuning} could go a long way. Regardless of the
-technique that we use, it is clear the the number of tunables increases the
+technique that we use, it is clear the number of tunables increases the
 physical design parameters to an unwieldy state space size.
@@
 different but to deal with the large state space we need a way to automatically
-and autonomically decide which parameters to choose.
+decide which parameters to choose.
 
@@
 As Ceph continues to develop into a target for building distributed services
-and applications, it is important to not restrict use to developers seasonsed
-in distributed programming and those with knowledge of the intracacies of Ceph
+and applications, it is important to not restrict use to developers seasoned
+in distributed programming and those with knowledge of the intricacies of Ceph
 and its performance model. The hard-coded imperative method of writing object
@@
 Trimming a log entry always succeeds. In Listing~\ref{lst:trim} the
-\texttt{<+-} operator simulatenously removes the log entry with the given 
+\texttt{<+-} operator simultaneously removes the log entry with the given 
 position and replaces it with an entry with its state set to \emph{trimmed}.
@@
 
-The write and fill interfaces are implemented similarily, and are both shown in
+The write and fill interfaces are implemented similarly, and are both shown in
 Listing~\ref{lst:write}.  A \texttt{valid\_write} collection is created if the
 operation position is not found in the log.  A \texttt{valid\_write} operation
-is then merged into log, otherwise a read only error is returend indicating
+is then merged into log, otherwise a read only error is returned indicating
 that the log position was already written to.  The fill operation is identical
-except a the fill state is set on the log entry.
+except the fill state is set on the log entry.
 
@@
 and does not have the filled or trimmed state set.  The data read from the log
-is returend in the case of a valid read operation, otherwise an error is
-returend through the output interface.
+is returned in the case of a valid read operation, otherwise an error is
+returned through the output interface.
 
@@
 
-But beyond the convience of writing less code, it is far easier for the
+But beyond the convenience of writing less code, it is far easier for the
 programmer writing an interface such as CORFU to convenience herself of the
@@
 The challenge of navigating the physical design space has served as the
-primary source of motivation for selection of a declarative langauge. While
+primary source of motivation for selection of a declarative language. While
 our implementation does not yet
@@
 graph that can be used in static analysis. We envision that this graph will be
-made available to the OSD and used to reorder and collesce requests based on
+made available to the OSD and used to reorder and coalesce requests based on
 optimization criteria available from a performance model combined with
@@
 (T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
-concerns have examined using managed lanuages, sandbox technologies, as well
+concerns have examined using managed languages, sandbox technologies, as well
 as restricting extension installation to
@@
 programming models in the context of active storage such as a stream-based
-model~\cite{acharya:asplos98}, a model that assisted in optmiizing balancing
+model~\cite{acharya:asplos98}, a model that assisted in optimizing balancing
 computation~\cite{wickremesinghe:hpdc022}, and optimization that took

### 2016-05-25 paper.tex
capitalized title
opt update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
-\title{brados: declarative, programmable object storage}
+\title{Brados: Declarative, Programmable Object Storage}
 
@@
 \subsection{Other Interfaces}
+\label{sec:oi}
 
@@
 
-The challenge of navigating the physical design space has served as the
-primary source of motivation for selection of a declarative language. While
-our implementation does not yet
+{\bf Physical design.} The challenge of navigating the physical design space
+has served as the primary source of motivation for selection of a declarative
+language. While our implementation does not yet make optimization decisions we
+can draw parallels between the challenges faced in physical design and the
+large body of mature work in query planning and optimization. Consider the
+simple interface reference counting interface described in
+Section~\ref{sec:oi}.  The C++ interface makes a strong assumption about a
+relatively small number of tags, and chooses to fully serialize and
+deserialize a C++ \texttt{std::map} for every request and store the marshalled
+data as an extended attribute.  As the number of tags grows the cost of false
+sharing will increase to the point that selection of an index-based interface
+will likely offer a performance advantage. While the monolithic version can
+outperform for small sets of tags, this type of optimization decision is
+precisely what can be achieved using a declarative interface definition that
+hides low-level evaluation aspects such as physical design.
+
+Other data structures.
 

### 2016-05-25 paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 We conclude by discussing the optimization opportunities that we can realize
-using analysis of interfaces.
+using analysis of interfaces. We hope you enjoy the show.
 
@@
 \subsection{Metadata Management}
+\label{sec:mm}
 
@@
 
-Other data structures.
-
-The Bloom language that we use as a basis for Brados produces a data flow
-graph that can be used in static analysis. We envision that this graph will be
-made available to the OSD and used to reorder and coalesce requests based on
-optimization criteria available from a performance model combined with
-semantic information from the dataflow. For example today object classes are
-represented as black boxes from the point of view of the OSD execution engine.
-Understanding the behavior of an object class may allow intelligent
-prefetching. Another type of analysis that may be useful for optimization is
-optimistic execution combined with branch prediction where frequent paths
-through a dataflow are handled optimistically.
+Looking beyond standard forms of optimization decisions that seek to select
+an appropriate mix of low-level I/O interfaces, data structure selection is an
+important point of optimization. For instance in Section~\ref{sec:mm} we
+showed that using the bytestream for metadata management as opposed to the
+key-value interface offered superior performance. However the unstructured
+nature of the bytestream data model imposes no restrictions on implementation
+or storage layout. Integration of common indexing techniques into an optimizer
+combined with a performance model will allow our CORFU interface to derive
+similar optimizations when appropriate.
+
+{\bf Static analysis.} The Bloom language that we use as a basis for Brados
+produces a data flow graph that can be used in static analysis. We envision
+that this graph will be made available to the OSD and used to reorder and
+coalesce requests based on optimization criteria available from a performance
+model combined with semantic information from the dataflow. For example today
+object classes are represented as black boxes from the point of view of the
+OSD execution engine.  Understanding the behavior of an object class may allow
+intelligent prefetching. Another type of analysis that may be useful for
+optimization is optimistic execution combined with branch prediction where
+frequent paths through a dataflow are handled optimistically.
 

### 2016-05-25 paper.tex
paper: minor edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-In this paper, we propose a declarative approach to subservice composition
-that allows programmers to focus on the high-level functional properties that
-are required by applications.  Choosing an implementation that is consistent
-with the declarative functional specification then can be posed as a search
-problem over the space of parameters such as block sizes, storage interfaces
-(e.g. key/value or block storage) and concurrency control mechanisms.  We
-present experimental evaluation of our prototype, (etc etc)
+In this paper, we propose a declarative approach to subservice composition that
+allows programmers to focus on the high-level functional properties that are
+required by applications.  Choosing an implementation that is consistent with
+the declarative functional specification then can be posed as a search problem
+over the space of parameters such as block sizes, storage interfaces (e.g.
+key/value or block storage) and concurrency control mechanisms. We base our
+observations and conclusions on data mining the git repository of the Ceph
+storage system and performance evaluating our own prototype implementations.
 \end{abstract}
@@
 relatively poor performance. Jewel with 1KB entry sizes sees a 2.5\(\times\)
-performance increase for the \texttt{(N:1,KV)} mapping. The large period of reduced
-throughput for corresponds to the OSD splitting file system directories in
-order to maintain a maximum directory size, and will occur in \texttt{(1:1,KV)}
-although the threshold number of objects is not reached in this example due to
-the reduced overall throughput of the \texttt{(1:1,KV)} strategy.
+performance increase for the \texttt{(N:1,KV)} mapping. This particular graph
+was extended to show an extra 30 minutes to demonstrate another disadvantage of
+\texttt{(1:1,BS)} - the large period of reduced throughput corresponds to
+the OSD splitting file system directories in order to maintain a maximum
+directory size, and will occur in \texttt{(1:1,KV)} although the threshold
+number of objects is not reached in this example due to the reduced overall
+throughput of the \texttt{(1:1,KV)} strategy.
 
@@
 
-\textbf{Ceph Features}: to show the affect of cutting-edge Ceph features, we
-now show results for BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
+\textbf{Ceph Features}: Ceph constantly adds new features and one particular
+feature that has the potential to greatly improve the performance of log
+appends is BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
 replacement for the FileStore file system in the OSD (traditionally XFS).
@@
 OSD.
+
+Unfortunately, due to space constraints, we do not show results for BlueStore.
+Anecdotally, performance for reads was a couple thousand IOPs better while
+writes are significantly worse. This is an example of a software feature that
+has a good architecture for a CORFU application on Ceph but is not
+production-ready yet.
 

### 2016-05-25 paper.tex
Tied up abstract
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@
 
+While our implementation does not yet map a declarative Brados specification
+to a particular physical design, the specification provides a powerful
+infrastructure for automating this mapping and achieving other optimizations.
+
 {\bf Physical design.} The challenge of navigating the physical design space
 has served as the primary source of motivation for selection of a declarative
-language. While our implementation does not yet make optimization decisions we
-can draw parallels between the challenges faced in physical design and the
+language. Given the declarative nature of the interfaces we have defined,
+we can draw parallels between the physical design challenges described in this
+paper and the
 large body of mature work in query planning and optimization. Consider the
